### 2017-07-18 paper.tex
asdfasdf

\documentclass[12pt]{report}

% This first part of the file is called the PREAMBLE. It includes
% customizations and command definitions. The preamble is everything
% between \documentclass and \begin{document}.

\usepackage[margin=1in]{geometry} % set the margins to 1in on all sides
\usepackage{graphicx} % to include figures
\usepackage{amsmath} % great math stuff
\usepackage{amsfonts} % for blackboard bold, etc
\usepackage{amsthm} % better theorem environments
\usepackage{xspace}

% various theorems, numbered by section

\newtheorem{thm}{Theorem} \newtheorem{claim}[thm]{Claim}
\newtheorem{remark}[thm]{Remark} \newtheorem{definition}[thm]{Definition}
\newtheorem{lem}[thm]{Lemma} \newtheorem{prop}[thm]{Proposition}
\newtheorem{cor}[thm]{Corollary} \newtheorem{conj}[thm]{Conjecture}

\DeclareMathOperator{\id}{id}

\newcommand{\reals}{\mathbb{R}}
\newcommand{\nats}{\mathbb{N}}

\newcommand{\logic}[1]{\text{\upshape #1}\xspace}
\newcommand{\FPC}{\logic{FPC}}
\newcommand{\FO}{\logic{FO}}
\newcommand{\FP}{\logic{FP}}
\newcommand{\FPR}{\logic{FPrk}}
\newcommand{\CPTC}{\logic{$\tilde{\text C}$PT(Card)}}

\newcommand{\PT}{\ensuremath{\mathrm{P}}\xspace}

\newcommand{\bd}[1]{\mathbf{#1}} % for bolding symbols
\newcommand{\RR}{\mathbb{R}} % for Real numbers
\newcommand{\ZZ}{\mathbb{Z}} % for Integers
\newcommand{\MB}{\mathbb{B}_{\matsym}} % for Integers
\newcommand{\SB}{\mathbb{B}_{\sym}} % for Integers
\newcommand{\col}[1]{\begin{matrix} #1 \end{matrix} \right]}
\newcommand{\comb}[2]{\binom{#1^2 + #2^2}{#1+#2}}
\newcommand{\stab}{\text{\textbf{Stab}}}
\newcommand{\setstab}{\text{\textbf{SetStab}}}
\newcommand{\matstab}{\text{\textbf{MatStab}}}
\newcommand{\inv}{\text{\textbf{inv}}}
\newcommand{\aut}{\text{\textbf{Aut}}}
\newcommand{\kernal}{\text{\textbf{ker}}}
\newcommand{\alt}{\text{\textbf{Alt}}}
\newcommand{\sym}{\text{\textbf{Sym}}}
\newcommand{\std}{\text{\textbf{std}}}
\newcommand{\rank}{\text{\textbf{rk}}}
\newcommand{\orb}{\text{\textbf{Orb}}}
\newcommand{\SP}{\text{\textbf{SP}}}
\newcommand{\SPs}{\text{\textbf{SPs}}}
\newcommand{\maj}{\text{\textbf{maj}}}
\newcommand{\dom}{\text{\textbf{Dom}}}
\newcommand{\img}{\text{\textbf{Img}}}
\newcommand{\child}{\text{child}}
\newcommand{\countgate}{\text{count}}
\newcommand{\agree}{\text{AGREE}}
\newcommand{\supp}{\text{SUPP}}
\newcommand{\rankparam}{\text{PARAM}}
\newcommand{\consp}{\text{sp}}
\newcommand{\consps}{\text{sps}}
\newcommand{\row}{\text{row}}
\newcommand{\column}{\text{col}}
\newcommand{\rowsp}{\text{r}}
\newcommand{\colsp}{\text{c}}
\newcommand{\EV}{\text{EV}}
\newcommand{\type}{\text{type}}
\newcommand{\types}{\text{types}}
\newcommand{\spstab}[1]{\stab (\consp (#1))}

\newcommand{\Alpha}{A}
\newcommand{\matsym}{\text{\textbf{MatSym}}}
\begin{document}

\title{Symmetric Circuits for Rank Gates}

\author{Anuj Dawar and Gregory Wilsenach \\ 
Computer Laboratory \\
University of Cambridge}

\maketitle

\tableofcontents

\chapter{Introduction}

A finite relational structure, such as a graph, can be represented as a binary
string in a standard way (for instance, by enumerating its adjacency matrix).
When we speak of the computational complexity of properties of graphs, we assume
such an encoding in order to present the structure as input to a machine or
circuit or other appropriate computational model. The properties of structures
we are interested in studying are invariant under isomorphisms. This requirement
translates into an invariance condition on sets of strings---say that a language
$L \subseteq \{0,1\}^\star$ is a \emph{graph property} if, whenever $x \in L$
for a string $x$ representing a graph $G$, we have $y\in L$ for any string $y$
obtained from $x$ by re-ordering the vertices of $G$. The central open question
in descriptive complexity---whether there is a logical characterisation of the
polynomial-time decidable properties of finite relational structures--- can then
be understood as the question of characterizing those languages $L$ that are in
$\PT$ and are graph properties.

Anderson and Dawar~\cite{AndersonD17} study this question through the lens of circuit
complexity, by considering \emph{symmetric circuits}. These are circuits that
necessarily accept graph properties, with this fact explicitly witnessed by
automorphisms of the circuit. To be precise, we can think of a property of
(directed) graphs on $n$ vertices as being given by a Boolean function $f:
\{0,1\}^{n^2} \rightarrow \{0,1\}$ which takes as input the $n^2$ potential
edges and outputs $1$ if the property is satisfied by the graph given on the
input. The function $f$ is \emph{graph-symmetric} in the sense that any permutation of
its inputs induced by a permutation of the $n$ vertices leaves $f$ unchanged
(note this is not the same thing as requiring that $f$ is invariant under all
permutations of its inputs). A symmetric circuit is a circuit computing a
Boolean function of $n^2$ inputs where each permutation of $[n]$ can be extended
to an automorphism of the circuit. Such a circuit necessarily computes an
invariant function.

The results of~\cite{AndersonD17} show a tight relationship between the decdability of
a graph property by means of a polynomial-size family of symmetric circuits and
its definability in extensions of fixed-point logic. In particular, they show
that a property is definable in fixed-point logic with counting ($\FPC$) if, and
only if, it is decidable by a polynomially-uniform family of symmetric circuits
using Boolean and majority gates (or, equivalently, threshold gates). Without
the symmetry restriction, polynomial-size families of circuits have the same
computational power whether we restrict ourselves to Boolean gates only or allow
richer bases such as threshold gates. One consequence of the results
in~\cite{AD14} is that, with the requirement of symmetry, these different bases
do yield different expressive power, suggesting some lack of robustness to the
model. One of the results in the present paper shows that the model with
majority circuits is robust in the sense that it is able to simulate, by means
of symmetric circuits, any symmetric circuit composed of gates computing Boolean
functions that are themselves symmetric (in the sense that they are invariant
under \emph{all} permutations of their inputs. Thus, any property decidable by a
polynomially-uniform family of such circuits is definable in $\FPC$.

This motivates the study of symmetric circuit families that involve gates which
themselves may compute non-symmetric functions. In particular, motivated by the
study of rank logic~\cite{DGHL09,GP15}, we consider gates that are
\emph{matrix-symmetric}. These are gates computing a Boolean function $f:
\{0,1\}^{A \times B} \rightarrow \{0,1\}$ which is invariant under arbitrary
permutations of $A$ and $B$. A typical example is the function computing the
rank of an $A \times B$ $0$-$1$ matrix. We show that the support theorem of
Anderson and Dawar can be shown to hold, even in the presence of
matrix-symmetric gates. This allows us to show that definability in the logic
fixed-point with rank ($\FPR$) is the same as decidability by
polynomially-uniform families of symmetric circuits with rank gates.

% Another extension of $\FPC$ that has received much attention in the descriptive
% complexity literature is the class $\CPTC$ of \emph{choiceless polynomial time
%   with counting}. The work of~\cite{AD14} leaves open the question of obtaining
% a circuit characterization of this class and it was suggested there that this
% would require a relaxation of the notion of symmetry used in defining symmetric
% circuits. In particular, while $\CPTC$ is based on a machine model designed to
% preserve symmetry by forbidding choices, the polynomial time and space
% restrictions apply int he context of the symmetries fo a \emph{single input
%   structure}. In this paper, we consider a relaxation of the notion of symmetric
% circuit intended to capture exactly this intuition. However, we are able to show
% that this does not yield a characterization of $\CPTC$. In particular, we show
% that no family in this class of relaxed symmetric circuits can distinguish odd
% from even Cai-F\"urer-Immerman graphs, something we know is do-able in
% $\CPTC$~\cite{DRR08}.

\chapter{Background}
In this section we present basic background on logics, first-order structures,
and symmetric circuits.

\section{Logic and Structures}
A \emph{relational vocabulary} (usually denoted by $\tau$) is a finite sequence
of relation symbols $R_1, \ldots, R_k$, such that for all $i \in [k]$, $R_i$ has
arity of $r_i \in \nats$.

A $\tau$-structure $\mathcal{A}$ is a tuple $\langle U, R^{\mathcal{A}}_1 ,
\ldots , R^{\mathcal{A}}_k \rangle$, where $U$ is a set and called the
\emph{universe} of $\mathcal{A}$ and for all $i \in [k]$ the relations
$R^{\mathcal{A}}_i \subseteq U^{r_i}$. The elements of $U$ are called the
elements of $\mathcal{A}$. A \emph{multi-sorted} structure is one whose universe
is a disjoint union, where each set in the union is called a \emph{sort}. The
size of a structure is the cardinality of its universe.

All structures in this paper are finite.

\section{First Order, Fixed-Point, Counting}

Let $\FO(\tau)$ denote \emph{first order logic} with respect to the vocabulary
$\tau$. \emph{Fixed-Point Logic} is the extension of $\FO(\tau)$ with an
inflationary fixed-point operator (see \cite{} for details). For logic
$\mathcal{L}$, a formula $\phi \in \mathcal{L}(\tau)$, $x$ a $k$-length tuple of
variables, $\mathcal{A}$ a (finite) structure, and $a \in U^k$, we have that
$\phi (x)$ denotes that the variables in $x$ are free in $\phi$ and we write
$\mathcal{A} \models_{\mathcal{L}} \phi[a]$ iff the the formula $\phi$ is true
in $\mathcal{A}$ with respect to $\mathcal{L}$ with the variables $x$ assigned
to $a$. The context usually makes the logic obvious, and in that case we usually
drop the subscript.

We also recall the definition of \emph{fixed-point logic with counting} $\FPC$,
an extension $\FP$ with counting quantifiers. For a structure $\mathcal{A}$,
Formulas of this logic are evaluated over a structure $\mathcal{A}$ by extending
that structure to $\mathcal{A}^\leq$, a two sorted extension of $\mathcal{A}$ by
an ordered number sort $\{1, \ldots, \vert U\vert \}$. Variables are taken to be
either point variables, ranging over $U$ or number variables ranging over
$[\vert U \vert]$. The counting quantifiers are of the form $\#_x$, where $x$ is
either a point or number variable and for $\phi(x) \in \FPC(\tau)$, $\#_x
\phi(x)$ is a term denoting the element of the number sort corresponding to
$\vert \{a \in \mathcal{A} | \mathcal{A} \models \phi[a]\}\vert$. Note
additionally that $k$-tuples of number variables can be used to encode numbers
in $\FPC$ of size up to $\vert U\vert^k$. For more details on $\FPC$ please see
\cite{}.

\section{Rank Logic}
This still needs to be filled in

\section{Symmetric Circuits}
This still needs to be filled in

\chapter{Symmetric Circuits}
In this section we discuss symmetric circuits, their limitations and extensions
of this model.

\section{Symmetric Circuits and Limitations}
Symmetric circuits, as developed by Anderson and Dawar, are, importantly defined
over (complete) Boolean basis consisting of symmetric functions. The following
result shows that once a majority gate has been included in this basis any
symmetric function can be computed using polynomial-size symmetric circuits, and
as such no additional symmetric function added to the Basis improves the power
of the model.

Recall from Anderson and Dawar \cite{AndersonD17} we have that $\mathbb{B}_{\std} = \{
\neg , \wedge , \lor \}$ and $\mathbb{B}_\maj = \{ \maj \} \cup \mathbb{B}$.

Let $F: \{0,1\}^* \rightarrow \{0,1\}$ be a symmetric Boolean function. Since
$F$ is symmetric we note that for a fixed size input the output of $F$ is
entirely determined by the number of 1's in its input. Then let
$c_{F}:\mathbb{N} \rightarrow 2^{\mathbb{N}}$ define a function where $c_{F}(n)$
is the set of all $m \leq n$ such that for all $\vec{x} \in \{ 0,1 \}^n$ with
$m$ 1's we have $F (\vec{x}) = 1$. Clearly any symmetric Boolean function $F$ is
entirely determined by $c_{F}$.

\begin{prop}
  \label{prop:fuctions-maj}
  There is a polynomial $p(k)$ such that for any symmetric function $F$ and a
  given $k \in \mathbb{N}$ there is a circuit $C_k$ on $k$ inputs over the basis
  $\mathbb{B}_\maj$ which is symmetric, constant depth and with width bounded by
  $p(k)$.
\end{prop}

\begin{proof}
  We define the circuit $C_k$ for inputs $\vec{x} = ( x_1, \ldots, x_k )$.

  For $a \in \mathbb{N}$ we define a get $\countgate_a$ by
  \begin{align*}
    &\countgate_a = \maj (x_1, \ldots, x_k, \underbrace {0, \ldots, 0}_{2a -
      k}) \land \neg \maj (x_1, \ldots , x_k, \underbrace{0, \ldots,  0}_{2a - k + 2})\text{ if $a \geq \frac{k}{2}$,} \\
    &\countgate_a = \maj (x_1, \ldots, x_k, \underbrace {1, \ldots, 1}_{x -
      2a}) \land \neg \maj (x_1, \ldots , x_k, \underbrace{1, \ldots,
      1}_{k - 2a -2}) \text{ if $a < \frac{k}{2}$. }
  \end{align*}

  Then let $g = \bigvee_{a \in c_{F_i}(k)}\countgate_a$ and let $C_{k}$ be the
  circuit with input gates labeled by $\vec{x}$ and output gate $g$.

  t is easy to see that $C_k$ is constant depth and it's width is a polynomial
  in $k$. We have that in each layer of $C_{g'}$ each gate is connected to all
  gates in the previous layer, and as such the circuit is symmetric.
\end{proof}

The above proposition has a straight forward application to circuit
characterisations.

\begin{thm}
  Let $F = \{F_i : i \in I \}$ be a family of symmetric Boolean functions where
  $F_i: \{0,1\}^* \rightarrow \{ 0,1 \}$.

  Let $(C_n)_{n \in \mathbb{N}}$ be family of symmetric circuits over the
  Boolean basis $\mathbb{B} \cup F$, where $C_n$ is a circuit on structures of
  size $n$, and the size of each circuit in the family is bounded by some
  function $f(n)$. Then there exists a polynomial $q(n)$ and a family of
  symmetric circuits $(C_n')_{n \in \mathbb{N}}$ over $\mathbb{B}_\maj$, where
  $C_n'$ is a circuit on structures of size $n$ and $\vert C_n' \vert \leq
  q(f(n))$.
\end{thm}

\begin{proof}
  From $C_n$ we construct $C_n'$ in the obvious way. For each gate $g \in C_n$
  of type $F_i$ we have a symmetric circuit $C_g$ from Proposition
  \ref{prop:function-maj} that computes the same function as $g$. Then let
  $C_n'$ be $C_n$ but with each gate $g$ replaced by $C_g$. It is easy to see
  that $C_n'$ is symmetric. We also have that each gate $g$ must have at most
  $f(n)$ inputs, and the size of $C_g$ is bounded by $p(f(n))$. Thus the size of
  $C_n'$ is bounded by $f(n)p(f(n))$.
\end{proof}

\section{G-Symmetric Functions}
In order to extend the notion of a circuit by extending the basis we need to
consider functions that are not symmetric in the full sense of the term but have
some weakened notion of symmetry.

\begin{definition}
  Let $f: \{0,1\}^X \rightarrow \{0,1\}$ be a function. Let $G \subseteq \sym_X$
  be a subgroup. We say that $f$ is \emph{$G$-symmetric} if for all $\sigma \in
  G$, $x \in X$, $f(\sigma x) = f(x)$.
\end{definition}

We can also introduce similar notions in the case of a two dimensional Boolean
string.

\begin{definition}
  Let $f: \{0,1\}^{A \times B} \rightarrow \{0,1\}$ be a function. Let $G
  \subseteq \sym_{A \times B}$ be a subgroup. We say that $f$ is
  \emph{matrix-symmetric} if it is $\sym_A \times \sym_B$-symmetric and we say
  that $f$ is \emph{graph-symmetric} if $A = B$ and it is $\sym_A$-symmetric
  (with $\sym_A$ acting on both $A$ and $B$).
\end{definition}

So if we think of $f: \{0,1\}^{A \times B} \rightarrow \{0,1\}$ as representing
a matrix, then matrix-symmetry corresponds to the property of the function being
invariant under row-column permutations. If, instead, we have that $A = B$ and
we then think of $f$ as encoding the adjacency matrix of a graph, then
graph-symmetry corresponds to the function being constant on isomorphism
classes.

Importantly, many natural functions of interest are matrix symmetric. For
example, the function that computes the rank of the matrix over $\mathbb{F}_2$.
or a thresholded rank function, for example the rank of the matrix over
$\mathbb{F}_p$ being larger then $r$, for some particular $(p, r) \in
\mathbb{N}$.

\section{Symmetric Circuits with Matrix-Symmetric Gates}

In this section we develop the corresponding notion of a symmetric circuit, ones
which allow for the inclusion of matrix-symmetric gates. In particular, these
circuits are defined using two sets of Boolean functions. The first, denoted by
$\mathbb{B}_\sym$, is taken to be the symmetric Boolean basis (either
$\mathbb{B}_\std$ or $\mathbb{B}_\maj$). The second, denoted by
$\mathbb{B}_\matsym$ is taken to be the matrix-symmetric basis, consisting of a
collection of matrix-symmetric functions.

\begin{definition}[Circuits on Structures]
  For $\mathbb{B}_{\sym}$ a basis of Boolean symmetric functions,
  $\mathbb{B}_{\matsym}$ a basis of Boolean matrix-symmetric functions and
  $\tau$ a set of relation symbols, we define a $(\mathbb{B}_\sym,
  \mathbb{B}_\sym, \tau)$-circuit $C_n$ computing a $q$-ary query $Q$ is a
  structure $\langle G, W, \Omega, \Sigma, \Lambda, L\rangle$.
  \begin{itemize}
    \setlength\itemsep{0mm}
  \item $G$ is called the set of gates of $C_n$ and $\vert C_n \vert := \vert G
    \vert$.
  \item $W \subseteq G \times G$, where $W$ is called the wires of the circuit.
    $(G,W)$ must be a directed acyclic graph. For $g \in G$ we $H_g := \{ h \in
    C_n : W(h,g)\}$ be the set of children of $g$.
  \item $\Omega$ is an injective function from $[n]^q$ to $G$. The gates in the
    image of $\Omega$ are called the output gates. When $q = 0$, $\Omega$ is a
    constant function mapping to a single output gate.
  \item $\Sigma$ is a function from $G$ to $\mathbb{B}_\sym \uplus
    \mathbb{B}_\matstab \uplus \tau \uplus \{0,1\} $ which maps input gates to
    $\tau \uplus \{0,1\}$ and where $\Sigma^{-1} (0) \leq 1$ $\Sigma^{-1} (1)
    \leq 1$ and the internal gates get mapped into $\mathbb{B}_\sym \uplus
    \mathbb{B}_\matstab$. Gates mapped to $\tau$ are called relational gates and
    gates mapped to 1 or 0 are called constant gates.
  \item $\Lambda$ is a sequence of injective functions $(\Lambda_R)_{R \in
      \tau}$ where for each $R \in \tau$, $\Lambda_R$ maps each relational gate
    $g$ with $R = \Sigma (g)$ to the tuple $\Lambda_R (g) \in [n]^r$, where $r$
    is the arity of the symbol $R$. When no ambiguity arises we write $\Lambda
    (g)$ for $\Lambda_R (g)$.
  \item $L$ assigns labels to the gates (and their inputs) in $C_n$. Let $g \in
    G$ be an internal gate and $H$ be the gates input to $g$. Then $L(g) : A
    \times B \rightarrow H$, where $L(g)$ must be surjective and $A, B \subseteq
    [n]$. We call this the \emph{matrix labelling} of $g$.
  \end{itemize}
\end{definition}

Note the additional requirement that each gate have an associated labelling in
order to facilitate the evaluation of the gate.

Given some finite $\tau$-structure $\mathcal{A}$ of size $n$ and some bijection
$\gamma: U \rightarrow [n]$, the evaluation of some $(\mathbb{B}_\sym,
\mathbb{B}_\sym, \tau)$-circuit) circuit $C$ proceeds by recursively evaluating
gates. Here the evaluation of the gate $g$ is denoted by $C[\gamma
\mathcal{A}](g)$. If $g$ is anything but a matrix-symmetric gate, then the
evaluation is as per the symmetric circuits of Anderson and Dawar \cite{}. If
$g$ is matrix-symmetric then the evaluation is given by applying the Boolean
operation $\Sigma(g)$ to the matrix $L^{\gamma}:\dom(L(g)) \rightarrow \{0,1\}$
defined by $L^{\gamma}(i,j) = C[\gamma \mathcal{A}](L(g)(i,j))$.

\begin{definition}[Invariant Circuit\cite{AndersonD17}]
  Let $C_n$ be a $(\mathbb{B}_\sym, \mathbb{B}_\sym, \tau)$-circuit, computing
  some $q$-ary query. We say $C_n$ is \emph{invariant} if for every
  $\tau$-structure $\mathcal{A}$ of size $n$, $a \in \mathcal{A}^q$, and
  bijections $\gamma_1, \gamma_2: U \rightarrow [n]$ we have that $C[\gamma_1
  \mathcal{A}](\Omega (\gamma_1 a)) = C[\gamma_2 \mathcal{A}](\Omega (\gamma_2
  a))$.
\end{definition}

\begin{definition}[Automorphism]
  let $C = \langle G, W, \Omega, \Sigma, \Lambda, L\rangle$ be a
  $(\mathbb{B}_\sym, \mathbb{B}_\matsym, \tau)$-circuit computing at $q$-ary
  query on structures of size $n$. Let $\sigma \in \sym_n$ and $\pi: G
  \rightarrow G$ be a bijection such that
  \begin{itemize}
    \setlength\itemsep{0mm}
  \item for all gates $g, h \in G$, $W(g,h)$ iff $W(\pi g, \pi h)$,
  \item for all output tuples $x \in [n]^q$, $\pi \Omega (x) = \Omega (\sigma
    x)$,
  \item for all gates $g \in G$, $\Sigma (g) = \Sigma (\pi g)$,
  \item for each relational gate $g \in G$, $\sigma \Lambda (g) = \Lambda (\pi
    g)$, and
  \item for each internal gate $g$ if $\Sigma (g) \in \mathbb(B)_\matsym$ then
    we have that $L(\pi g) \sim \pi \cdot L(g)$.
  \end{itemize}

  We call $\pi$ an \emph{automorphism} of $C$, and we say that $\sigma$
  \emph{induces the automorphism} $\pi$. The group of automorphisms of $C$ is
  called $\aut_n (C)$.
\end{definition}

\begin{definition}[Symmetric\cite{AndersonD17}]
  A circuit $C$ on structures of size $n$ is called \emph{symmetric} if every
  $\sigma \in \sym_n$ induces an automorphism on $C$.
\end{definition}

It follows for any symmetric circuit $C_n$ there is a mapping from $\sym_n$ to
$\aut_n(C)$, from $ \sigma$ to the induced automorphism. This homomorphism is
injective so long as a single element of $[n]$ appears in the in the label of
some input gate of $C$ (as then all elements appear by
symmetry)\cite{AndersonD17}. In this paper we always assume that there is always
one such element as otherwise all inputs are constant, and so the circuit just
computes a constant function. In order to assure this homomorphism is surjective
Anderson and Dawar \cite{AndersonD17} introduce the notion of a \emph{rigid}
circuit.

\begin{definition}[Rigidity]
  Let C be a $(\mathbb{B}_\sym, \mathbb{B}_\sym, \tau)$-circuit, where $C =
  \langle G, W, \Omega, \Sigma, \Lambda, L\rangle$. Say that $C$ is rigid if
  there are no internal gates $g, g' \in G$ such that $\Sigma(g) = \Sigma (g')$,
  $\Omega^{-1}(g) = \Omega^{-1}(g')$, and for every $g'' \in G$, $W(g'', g')$
  and $W(g,g'')$ iff $W(g', g'')$.
\end{definition}

Another property which simplifies our analysis is the property of having
\emph{bijective labels}.

\begin{definition}
  We say that a circuit $C$ has bijective labels if for each gate $g$ in $C$,
  $L(g)$ is a bijection.
\end{definition}

We prove in a later section that a circuit may be transformed in polynomial time
into an equivalent circuit that is both rigid and has bijective labelings. Hence
we may assume these two properties without a loss of generality.

With the assumption of rigidity in place, we abuse notation for permutations and
let $\sigma \in \sym_n$ also denote the induced automorphism.

We finally want to define symmetric circuits with rank gates.

\begin{definition}
  Let $\mathbb{B}_{\matsym}$ consist of functions of the form $f_{p,r}:
  \{0,1\}^{A \times B} \rightarrow \{0,1\}$ for $A, B$ initial segments of
  $\mathbb{N}$ and $f, p \in \mathbb{N}$, with $p$ prime, and such that for
  $f_{p,r}(M) = 1$ iff the matrix $M: A \times B \rightarrow \{0,1\}$ has rank
  at least $r$ over the field $\mathbb{F}_p$. A \emph{symmetric circuit with
    rank gates} is a $(\mathbb{B}_\maj, \mathbb{B}_\matsym, \tau)$-circuit, for some
  vocabulary $\tau$.
\end{definition}

The natural restriction to consider on families of circuits is uniformity.

\begin{definition}
  Let $(C_n)_{n \in \mathbb{N}}$ be a family of Boolean circuits. We say that
  $(C_n)_{n \in \mathbb{N}}$ is \emph{$P$-uniform} if the mapping $n \mapsto C_n$ is
  computable in polynomial time.
\end{definition}

We are now ready to state what will be the main theorem of this paper.

\begin{thm}[Main Theorem]
 A graph property is decidable by a $P$-uniform family of symmetric circuits
 with rank gates if, and only if, it is definable by an FPR sentence.
\end{thm}

% \section{Circuits to FPR}

% In this section I discuss how to determine the rank a matrix labelling at some
% gate $g$. This is a rough copy written for the purpose of discussion. In this
% section we fix some circuit $C_n$ and some rank gate $g$ in $C_n$ with a set
% of children $H$ and matrix labelling $L: [a] \times [b] \rightarrow H$ (which
% we assume WLOG is a bijection).

\chapter{Symmetry and Support}
In order to analyse families of symmetric circuits we will need to develop some
theory for analysing symmetries. In this section we define the terms needed
(e.g. the notion of a support) and prove basic results. We then prove the first
important theorem in this paper, the support theorem, using these tools. It is
worth noting that we prove the support theorem for arbitrary symmetric circuits
with matrix-symmetric gates, not just for rank gates.

\section{Supports and Supporting Partitions}
In this section we formally define a support and the related notion of a
supporting partition.

\begin{definition}
  Let $G$ be a subgroup of $\sym_n$ and let $S \subseteq [n]$. Then $S$ is a
  \emph{support} for $G$ if $\stab(S) \subset G$.
\end{definition}

\begin{definition}
  Let $G$ be a subgroup of $\sym_n$ and $\mathcal{P}$ be a partition of $[n]$.
  Then $\mathcal{P}$ is a \emph{supporting partition} for $G$ if
  $\stab(\mathcal{P}) \subseteq [n]$.
\end{definition}

Notice that if $\mathcal{P}$ is a supporting partition for $G$ and $P \in
\mathcal{P}$ then $\stab([n] \setminus P) \subseteq \stab(\mathcal{P}) \subset
G$, i.e. $[n] \setminus P$ is a support for $G$.

These supporting partitions admits a natural partial order in terms of
coarseness. If $\mathcal{P}, \mathcal{P}'$ we say that $\mathcal{P}'$ is as
course as $\mathcal{P}$ (denoted by $\mathcal{P}' \supseteq \mathcal{P}$) if for
all $x \in \mathcal{P}$ there exists $y \in \mathcal{P}'$ such that $x \subseteq
y$ \cite{AndersonD17}.

Anderson and Dawar \cite{AndersonD17} show that for any permutation group $G
\subseteq \sym_n$ a unique coarsest partition can always be found. We call this
partition the \emph{canonical supporting partition}, and denote it by $\SP (G)$.
This canonical supporting partition, along with the above observation about
supports, will allow us to construct for each group a canonical support.

\begin{definition}
  Let $G$ be a subgroup of $\sym_n$. Let $\| SP(G) \| = \min \{\vert [n]
  \setminus P \vert : P \in SP(G) \}$. We say that $G$ has small support if $\|
  SP(G) \| < \frac{n}{2}$.
\end{definition}

\begin{definition}
  Let $G$ be a subgroup of $\sym_n$ such that $G$ has small support. Let
  $\consp(G) = [n] \setminus P$, where $P$ is the maximal element of $\SP(G)$.
  We call $\consp(G)$ the canonical support of $G$.
\end{definition}

\begin{remark}
  Include a remark about the importance of asserting the size consideration. Use
  the alternating group as an example.
\end{remark}

We state the following two results from Anderson and Dawar \cite{AndersonD17}.

\begin{lem}
  \label{lem:SP_conjugation}
  Let $G \leq \sym_n$ and $\sigma \in \sym_n$ then $\sigma \SP (G) = \SP(\sigma
  G \sigma^{-1})$.
\end{lem}

\begin{lem}
  For any $G \leq \sym_n$ we have that $\stab (\SP (G)) \subseteq G
  \setstab(\SP(G))$.
\end{lem}

\section{Group Action on Supports}
In this paper, distinct the case of Anderson and Dawar \cite{AndersonD17}, we
will need to develop theory and terminology for dealing with group actions (and
supports) on more than just gates in the circuit. In particular, we use this
theory for understanding the actions on the rows and columns of a labelling of a
gate.

\begin{definition}
  Let $G$ be a subgroup of $\sym_n$ and $X$ be a set on which a left group
  action of $G$ on $X$ is defined. We denote the canonical supporting partition
  of $x\in X$ by $\SP (x) = \SP (\stab(x))$. We say that $x \in X$ has small
  support if $\stab(x)$ has small support. We say that $X$ has small supports if
  all $x \in X$, $x$ have small support.

  If $x \in X$ has small support, we denote the \emph{canonical support} of $x$
  by $\consp(x) = \consp (stab(x))$.
\end{definition}

\begin{definition}
  Let $G$ be a group and $X$ be a set on which a left group action of $G$ on $X$
  is defined. Then $\stab_G (x) = \{\pi \in G : \pi x = x\}$ and $\orb_G (x) =
  \{\pi x : \pi \in G\}$. In the event that the context makes the group obvious
  we omit the subscript.
\end{definition}

\begin{lem}
  \label{lem:stab_conjugation}
  Let $G$ be a subgroup of $\sym_n$, $X$ be a set on which a left group action
  of $G$ on $X$ is defined and $\sigma \in G$. Then for any $x \in X$ it follows
  that $\sigma \stab (x) \sigma^{-1} = \stab(\sigma x)$.
\end{lem}

\begin{proof}
  Let $\pi \in \stab(x)$, then $\sigma \pi \sigma^{-1}(\sigma x) = \sigma \pi
  (x) = \sigma (x)$, and so $\sigma \pi \sigma^{-1} \in \stab(\sigma x)$. Let
  $\pi \in \stab(\sigma (x))$ then $\pi \sigma (x) = \sigma (x)$ and so
  $\sigma^{-1} \pi \sigma (x) = x$. It follows that $\sigma^{-1} \pi \sigma \in
  \stab(x)$ and so $\pi = \sigma (\sigma^{-1} \pi \sigma) \sigma ^{-1} \in
  \sigma \stab(x) \sigma^{-1}$.
\end{proof}

\begin{lem}
  \label{lem:support_mapping}
  Let $G $ be a subgroup of $\sym_n$, $X$ be a set on which a left group action
  of $G$ on $X$ is defined and $\sigma \in G$. Then for any $x \in X$ it follows
  that $\sigma \SP (x) = \SP (\sigma x)$ and, if $x$ has small support, $\sigma
  \consp (x) = \consp (\sigma x)$.
\end{lem}
\begin{proof}
  From Lemma \ref{lem:SP_conjugation} and Lemma \ref{lem:stab_conjugation} have
  that $\sigma \SP (x) = \SP(\sigma \stab(h) \sigma^{-1}) = \SP (\stab(\sigma
  x)) = \SP (\sigma x)$. Thus proving the first part of the statement.

  From the fact that $\| \SP(x) \| < \frac{n}{2}$ it follows there exists a
  unique $P_1 \in \SP(x)$ such that $\vert P_1 \vert > \frac{n}{2}$ and
  $\consp(x) = [n] \setminus P_1$. But then $\sigma \consp (x) = \sigma([n]
  \setminus P_1) = [n] \setminus (\sigma P_1)$. We note that $\sigma P_1 \in
  \sigma \SP (x) = \SP(\sigma x)$ and $\vert \sigma P_1 \vert > \frac{n}{2}$.
  Thus $\sigma P_1$ is the unique largest partition in $\SP(\sigma x)$, and so
  $\consp(\sigma x) = [n]\setminus (\sigma P_1) = \sigma (consp(x))$. Thus
  proving the second part of the statement.
\end{proof}

\begin{lem}
  Let $G$ be a subgroup of $\sym_n$, $X$ be a set on which a left group action
  of $G$ on $X$ is defined. Then for any $x \in X$ with small support, and
  $\sigma, \sigma' \in G$ such that $\sigma (vec{\consp}(x)) = \sigma'
  (\vec{\consp}((x))$ we have that $\sigma (x) = \sigma'(x)$.
\end{lem}
\begin{proof}
  We have that $(\sigma')^{-1}\sigma(\vec{\consp}(x)) = \vec{\consp}(x)$ and so
  $(\sigma')^{-1} \sigma (x) = x$ and thus $\sigma (x) = \sigma' (x)$.
\end{proof}

% \begin{lem}
%   Let $G \leq \sym_n$, $X$ be a set on which a left group action of $G$ on $X$
%   is defined and $\sigma \in G$. Then if $y \in \orb_{G} (x)$ and $\consp (x)
%   = \consp (y)$ then $x = y$.
% \end{lem}

% \begin{proof}
%   Let $\sigma \in G$ be such that $\sigma x = y$. Then $\consp(x) = \consp (y)
%   = \consp (\sigma x) = \sigma \consp (x)$. Thus $\sigma (\sp (x))$
% \end{proof}

% \begin{definition}
%   Let $h \in H$. Let $\stab_r(h)$ be a subgroup of $\stab(\consp(g))$
%   consisting of exactly those permutations $\sigma$ such that $\row (\sigma h)
%   = \row (h)$. Let $\stab_c(h)$ be a subgroup of $\stab(\consp(g))$ consisting
%   of exactly those permutations $\sigma$ such that $\column (\sigma h) =
%   \column(h)$.
% \end{definition}

% \begin{definition}
%   A set $S \subseteq [n]$ is called a row-support for $h$ if $S$ is a support
%   for $\stab_r(h)$. A set $S \subseteq [n]$ is called a column-support for $h$
%   iff $S$ is a support for $\stab_c(h)$.
% \end{definition}

% \begin{definition}
%   A partition $\mathcal{P}_r$ is called a row-supporting partition iff
%   $\mathcal{P}_r$ is a support for $\stab_r(h)$. A partition $\mathcal{P}_c$
%   is called a column-supporting partition iff $\mathcal{P}_c$ is a support for
%   $\stab_c(h)$.
% \end{definition}

% \begin{definition}
%   Let $\SP^r(h)$ denote the conical supporting partition and $r_h$ the
%   canonical support of $\stab_r(h)$. Similarly, let $\SP^c(h)$ denote the
%   conical supporting partition and $c_h$ the canonical support of
%   $\stab_c(h)$.
% \end{definition}

% \begin{definition}
%   Let $G \subseteq \sym_n$ be a subgroup. Then let $\SPs(G)$ be the set of all
%   supporting partitions of $G$ and $\consps(G)$ be the set of all supports of
%   $G$.
% \end{definition}

% We now prove a few important lemmas on supports.

% \begin{lem}
%   Let $A, B \leq \sym_n$ then $\SP(A \cap B) \subseteq \SP (A) $.
% \end{lem}
% \begin{proof}
%   Let $\mathcal{P}$ be a supporting partition of $A \cap B$. Then
%   $\mathcal{P}$ is a supporting partition of $A$. Since $\SP(A)$ is coarser
%   than all such partitions the result follows.
% \end{proof}

% \begin{lem}
%   Let $h \in H$, then $\SP(\stab(h)) \subseteq \SP (\stab (\row (h)))$ and
%   $\SP (\stab(h)) \subseteq \SP (\stab (\column(h)))$.
% \end{lem}
% \begin{proof}
%   It can be shown that $\stab (h)$
% \end{proof}

% The following lemma will be used to prove the relationship between the support
% of a gate and the supports of its row and columns.
% % Use this lemma + a support type theorem in order to get the result.
% \begin{lem}
%   Let $G \leq \sym_n$, $X_1$ and $X_2$ be sets on which left group actions of
%   $G$ on $X_1$ and $X_2$ are defined. Let $x_1 \in X_1$ and $x_2 \in X_2$ and
%   suppose $\vert \consp(x_1) \vert < \frac{n}{2}$ and $ \vert \consp(x_2)
%   \vert < \frac{n}{2}$. Then if $\SP(\stab(x_1)) \subseteq \SP (\stab(x_2))$
%   then $\consp (x_2) \subseteq \consp (x_1)$.
% \end{lem}
% \begin{proof}
%   Since $\vert \consp(x_1) \vert = \| \SP (\stab(x_1)) \| < \frac{n}{2}$ it
%   follows that there a unique maximal $P_1 \in \SP (\stab(x_1))$ such that
%   $\vert P_1 \vert > \frac{n}{2}$. Similarly there is a unique maximal $P_2
%   \in \SP (\stab(x_2))$. Moreover, from the fact that $\SP (\stab(x_2))$ is
%   coarser than $\SP (\stab(x_1))$, it follows that there exists $P \in
%   \SP(\stab(x_2))$ such that $P_1 \subseteq P$. From the fact that
%   $\frac{n}{2} < \vert P_1 \vert \leq \vert P \vert$ and $P_2$ is the only
%   part such that $\frac{n}{2} < \vert P_2 \vert$, it follows that $P_1
%   \subseteq P = P_2$ and so $\consp(x_2) = [n] - P_2 \subseteq [n] - P_1 =
%   \consp(x_1)$.
% \end{proof}

\section{Support Theorems}
In this section we develop an analogous theorem to the support theorem of
Anderson and Dawar \cite{AndersonD17}.
%
%\begin{lem}[Support Lemma]
%  For any $\epsilon$ and $n$ such that $\frac{2}{3} \leq \epsilon \leq 1$ and
%  $n \geq \frac{128}{\epsilon^2}$, if $C$ is a symmetric, rigid circuit on
%  structures of size $n$ and $s := \max_{g \in C} \vert \orb (g)\vert \leq
%  2^{n^{1-\epsilon}}$, then, for all $g \in C$, $\vert \SP (g) \vert \leq
%  \frac{n}{2}$.
% \end{lem}

% \begin{thm}[Support Theorem]
%   For any $\epsilon$ and $n$ such that $\frac{2}{3} \leq \epsilon \leq 1$ and
%   $n \geq \frac{128}{\epsilon^2}$, if $C$ is a symmetric, rigid circuit on
%   structures of size $n$ and $s := \max_{g \in C} \vert \orb (g)\vert \leq
%   2^{n^{1-\epsilon}}$, then, $SP(C) \leq \frac{33}{\epsilon}\frac{\log s}{\log
%   n}$.
% \end{thm}

% \begin{lem}
%   For any $\epsilon$ and $n$ such that $\frac{2}{3} \leq \epsilon \leq 1$ and
%   $n \geq \frac{128}{\epsilon^2}$, if $C$ is a symmetric, rigid circuit on
%   structures of size $n$ and $s := \max_{g \in C} \vert \orb (g)\vert \leq
%   2^{n^{1-\epsilon}}$, then, for all $g \in C$, we have for all $(i,j) \in
%   \dom (L_g)$ that $\consp(i) \subseteq \consp (L_g(i,j))$ and $\consp(j)
%   \subseteq \consp(L_g (i,j)))$.
% \end{lem}

% \begin{lem}
%   Let $\mathcal{P}_1$ and $\mathcal{P}_2$ be partitions of $[n]$ such that
%   $\mathcal{P}_1 = A \cup S_1$ and $\mathcal{P}_2 = B \cup S_2$, and
%   \begin{enumerate}
%   \item for all $X \in A$ and $Y \in B$ we have that $X \cap Y = \emptyset$,
%     and
%   \item for all $x \in S_1 \cup S_2$, x is a singleton.
%   \end{enumerate}
%   It follows that $\varepsilon(\mathcal{P}_1, \mathcal{P}_2) = A \cup B \cup
%   S$, where $S$ is a set of singletons.
% \end{lem}
% \begin{proof}
% \end{proof}

% \begin{lem}
%   Let $G_1$ and $G_2$ be subgroups of $\sym_n$. Then $\SP^S(G_1) = \SP^S(G_2)$
%   iff $\consp^S(G_1) = \consp^S(G_2)$.
% \end{lem}
% \begin{proof}

% \end{proof}

% \begin{lem}
%   Let $G_1$ and $G_2$ be subgroups of $\sym_n$. Then $\consp^S(G_1) =
%   \consp^S(G_2)$ implies that $\consp (G_1) = \consp (G_2)$.
% \end{lem}
% \begin{proof}
% \end{proof}

% The following lemma asserts a few basic facts about the canonical row and
% column supports.

% \begin{lem}
%   Let $h \in H$, then $r_h \subset \consp(h)$, $c_h \subset \consp(h)$ and
%   $c_r \cup r_h = \consp(h)$.
% \end{lem}

\begin{definition}
  Let $\omega$ be a matrix labeling for $g$. Define the matrix stabilizer for
  $\omega$, denoted by $\matstab(\omega)$, to be the set of all $\sigma \in
  \sym_n$ such that $\sigma H = H$ and there exists $(\alpha, \beta) \in \sym_A
  \times \sym_B$ such that for all $(i,j) \in A \times B$ we have that $\omega
  (\alpha i, \beta j) = \sigma \omega (i,j)$.
\end{definition}

\begin{definition}
  Let $g$ be a gate with matrix labeling $\omega$, $h,h' \in H$ and $\sigma \in
  \sym_n$. We say that a pair $(h, h')$ is compatible with $(\sigma, \omega)$ if
  $\sigma h, \sigma h' \in H$ and there exists $(\gamma_1, \gamma_2),
  (\gamma_1', \gamma_2') \in \sym_A \times \sym_B$ s.t.
  \begin{align*}
    &(\gamma_1, \gamma_2) \cdot \omega^{-1} (h) = \omega^{-1}(\sigma h), \text{ and} \\ 
    &(\gamma_1', \gamma_2') \cdot \omega^{-1} (h') = \omega^{-1}(\sigma h'),
  \end{align*}
  and for all $(i,j) \in \omega^{-1}(h)$, $(i',j') \in \omega^{-1}(h')$ we have
  that
  \begin{align*}
    &i =i' \implies \gamma_1(i) = \gamma_1'(i'), \text{ and} \\
    &j =j' \implies \gamma_2(j) = \gamma_2'(j').
  \end{align*}

\end{definition}

\begin{lem}
  Let $g$ be a gate, $\omega$ be a matrix labeling of $g$, $\sigma \in \sym_n$.
  Then $\sigma \in \matstab(\omega)$ iff for all $h,h' \in H$ we have that
  $(h,h')$ is compatible with $(\sigma, \omega)$.
\end{lem}

\begin{proof}
  $`\Rightarrow'$: We have that $\sigma \in \matstab(\omega)$ and so there
  exists $(\alpha, \beta) \in \sym_A \times \sym_B$ such that for all $(i,j) \in
  A \times B$ we have $\sigma \omega (i,j) = \omega (\alpha i, \beta j)$. From
  compatibility we have that $h,h' \in H$ Let $h, h' \in H$ . Let $(\gamma_1,
  \gamma_2) := (\gamma_1', \gamma_2') := (\alpha, \beta)$. This assignment is
  sufficient to prove the direction.

  $`\Leftarrow'$: Suppose for all $h,h' \in H$ we have $(\gamma_1, \gamma_2),
  (\gamma_1, \gamma_2') \in \sym_A \times \sym_B$ satisfying the above
  requirements. Notice that for a given $i \in A$ and any $j, j' \in B$, let $h
  = \omega(i,j)$ and $h' = \omega(i,j')$, then we have that $\gamma_1 (i) =
  \gamma_1'(i)$. It follows that we can define a $\alpha \in \sym_A$ by
  $\alpha(i) = \gamma_1 (i)$. Similarly we can define $\beta \in \sym_B$ by
  $\beta (j) = \gamma_2 (j)$.

\end{proof}

\begin{definition}
  Let $g$ be a gate with matrix labeling $\omega$. Let $(\sigma, h, h') \in
  \sym([n]) \times H^2$. Say that $(\sigma, h, h')$ is useful if $(h,h')$ is
  incompatible with $(\sigma, \omega)$.

  Say that two distinct pairs $(\sigma_1, h_1, h_1'), (\sigma_2, h_2, h_2') \in
  \sym_n\times H^2$ are mutually independent if
  \begin{itemize}
    \setlength\itemsep{0mm}
  \item $\sigma_2 h_1 = h_1$,
  \item $\sigma_2 \sigma_1 h_1 = \sigma_1 h_1$,
  \item $\sigma_2 h_1' = h_1'$,
  \item $\sigma_2 \sigma_1 h_1' = \sigma h_1'$,
  \end{itemize}
  We say that a set $S \subseteq \sym_n \times H^2$ is useful if each pair in
  it is useful. We say that S is independent if each pair of distinct pairs in
  $S$ are mutually independent.
\end{definition}

We denote the usual equivalence relation on $\sym_n$ from the (right) cosets of
$\matstab(\omega)$ by $\sim_\omega$.

\begin{lem}
  Let $\sigma_1, \sigma_1 \in \sym_n$ and suppose $\sigma_1 \sim_\omega
  \sigma_2$ then for any $(h,h') \in H^2$ we have that $(h,h')$ is compatible
  with $(\sigma_1, \omega)$ iff $(h,h')$ is compatible with $(\sigma_2,
  \omega)$.
\end{lem}

\begin{proof}
  Suppose we have that $\sigma_1 \sim_\omega \sigma_2$. And additionally
  supposed $(h, h')$ compatible with $(\sigma_1, \omega)$.

  It follows there exists $(\alpha, \beta) \in \sym_A \times \sym_B$ such that
  $\sigma_1 \omega (i,j) = \sigma_2\omega (\alpha i, \beta j)$, and we have
  $(\gamma_1, \gamma_2), (\gamma_1', \gamma_2') \in \sym_A \times \sym_B$
  satisfying the requirements of compatibility. We note that $(\alpha, \beta)
  \cdot \omega^{-1}(\sigma_1 h) = \omega^{-1}(\sigma_2 h)$. Thus by composition
  we have that $(\alpha \gamma_1, \beta \gamma_2) \cdot \omega^{-1}(h) =
  \omega^{-1}(\sigma_2 h)$ and $(\alpha \gamma_1', \beta \gamma_2') \cdot
  \omega^{-1}(h') = \omega^{-1}(\sigma_2 h')$, and clearly the remaining
  requirement for compatibility follows from the bijectivity of $\alpha$ and
  $\beta$. The result follows. The other direction of the implication follows
  from symmetry.
\end{proof}

\begin{claim}
  \label{claim:useful-independant-set}
  Let $g$ be a rank gate with labeling $\omega$ and child set $H$. Let $S$ be a
  useful and independent. We then have that $\vert \sym_n: \matstab(\omega)
  \vert \leq 2^{\vert S \vert}$.
\end{claim}

\begin{proof}
  Let $R \subseteq S$ and define $\sigma_R = \Pi_{(\sigma, h, h') \in R} \sigma$
  (with some arbitrary order assumed on $S$). Let $R$ and $Q$ be distinct
  subsets of $S$ and WLOG let $\vert R \vert \geq \vert Q \vert$. We want to
  show that we don't have $\sigma_R \omega \sim_\omega \sigma_Q \omega$. Pick
  any $(\sigma, h, h') \in R/Q \neq \emptyset$. Given that $\sigma_R h = \sigma
  h$ and $\sigma_R h' = \sigma h'$ it is easy to see that the usefulness of
  $(\sigma, h,h')$ implies the incompatibility of $(h,h')$ with with $(\omega,
  \sigma_R)$. Moreover, the fact that $\sigma_Q h = h$ and $\sigma_Q h' = h'$
  makes it easy to see $(h,h')$ is compatible with $(\omega, \sigma_Q)$. From
  the above lemma we may conclude that that we do not have $\sigma_R \sim_\omega
  \sigma_Q$, and the result follows.
\end{proof}

The following two lemmas proved by Anderson and Dawar \cite{AndersonD17} are
both of use in proving the following theorem.

\begin{lem}
  \label{lem:big-or-small}
  For any $\epsilon$ and $n$ such that $0 < \epsilon < 1$ and $\log n \geq
  \frac{4}{\epsilon}$, if $\mathcal{P}$ is a partition of $[n]$ with $k$ parts,
  $s = [\sym_n : \setstab (\mathcal{P})]$ and $n \leq s leq 2^{n^{1-\epsilon}}$,
  then $\min \{k, n-k\} \leq \frac{8}{\epsilon} \frac{\log s}{\log n}$.
\end{lem}

\begin{lem}
  \label{lem:small-means-support}
  For any $\epsilon$ and $n$ such that $0 < \epsilon < 1$ and $\log n \geq
  \frac{8}{\epsilon^2}$, if $\mathcal{P}$ is a partition of $[n]$ with $\vert
  \mathcal{P} \vert \leq \frac{n}{2}$, $s:= [\sym_n : \setstab (\mathcal{P})]$
  and $n \leq s \leq 2^{n^{1-\epsilon}}$, then $\mathcal{P}$ contains a part $P$
  with at least $n - \frac{33}{\epsilon} \cdot \frac{\log s} {\log n}$.
\end{lem}

If $g$ is a symmetric gate (i.e. the usual gates in a circuit) we note that
$\orb(g) = [\sym_n : \stab (g)]$ by the orbit-stabilizer theorem.

If $g$ is a matrix-symmetric gate then $\orb(g) = [\sym_n:\matstab (\omega)]$,
where $\omega$ is the matrix labelling associated with $g$.

\begin{thm}
  For any $\epsilon$ and $n$ such that $\frac{2}{3} \leq \epsilon \leq 1$ and $n
  \geq \frac{128}{\epsilon^2}$, if $C$ is a symmetric, rigid circuit on
  structures of size $n$ and $s := \max_{g \in C} \vert \orb (g)\vert \leq
  2^{n^{1-\epsilon}}$, then, $SP(C) \leq \frac{33}{\epsilon}\frac{log s}{log
    n}$.
\end{thm}

\begin{proof}
  It is easy to see that if $g$ is a gate in $C$ then $\stab (g) \subseteq
  \setstab(\SP(g))$, and so $s \geq \orb(g) = [\sym_n : \stab(g)] \geq [\sym_n :
  setstab(\SP(g))]$. Thus if $\vert \SP(g) \vert \leq frac{n}{2}$, then from
  Lemma \ref{lem:small-means-support}, we have $\| SP (g) \| \leq
  \frac{33}{\epsilon} \cdot \frac{\log s} {\log n}$. The result thus follows
  from showing that for each $g$ in $C$ we have that $\vert \SP (g) \vert \leq
  \frac{n}{2}$.

  % Include some detail here for constant and relational gates
  The cases where $g$ is a constant or relational gate are easy to handle.

  We now consider the case for internal gates. Let $g$ be the topologically
  first internal gate with $\vert \SP(g) \vert > \frac{n}{2}$. If $g$ is not a
  matrix-symmetric gate then the result follows from the argument presented by
  Anderson and Dawar \cite{AndersonD17}. Suppose $g$ is a matrix-symmetric gate, and
  suppose $g$ has a labelling $\omega$. We now argue that this leads to a
  contradiction.

  Let $k' := \lceil \frac{8 \log s}{\epsilon \log n} \rceil$. From the
  assumptions on $s, n$ and $\epsilon$ we have that $k' \leq
  \frac{1}{4}n^{1-\epsilon} < \frac{n}{2}$. Lemma \ref{lem:big-or-small} implies
  that $n - \vert \SP(g) \vert \leq k'$

  From Claim \ref{claim:useful-independant-set} it remains to show that we can
  construct a sufficiently large useful and independent set of gate-automorphism
  pairs $S$. Divide $[n]$ into $\lfloor \frac{n}{k' + 2} \rfloor$ disjoint sets
  $S_i$ of size $k' + 2$ and ignore the elements left over. It follows that for
  each $i$ there is a permutation $\sigma_i$ which fixes $[n] / S_i$ pointwise
  but moves $g$. Suppose there was no such $\sigma_i$ it follows that every
  every permutation that fixes $[n]/S_i$ pointwise fixes $g$. Thus the partition
  of all the singletons in $[n]/S_i$ and $S_i$ is a supporting partition. As
  $\SP(g)$ is the coarsest partition it follows that $\vert \SP(g) \vert \leq n
  - (k'+2) + 1 = n - k' - 1$, which contradicts the inequality $n - \vert \SP(g)
  \vert \leq k'$.

  Since $g$ is moved by each $\sigma_i$ and $C$ is rigid it follows that we
  don't have $\sigma_i \notin \matstab(\omega)$. Thus there exists $(h_i, h_i')$
  that is inconsistent with $(\sigma_i, \omega)$, and so the triple $(\sigma_i,
  h_i, h_i')$ is useful.

  Let $\SP (h)^*$ be the union of all parts of $\SP(h)$ except for the largest
  part. Let $Q_i = \SP(h_i)^* \cup \SP(\sigma_i h_i)^* \cup \SP (h_i')^* \cup
  \SP (\sigma_i h_i')^*$. Then note that if $\sigma_j$ fixes $Q_i$ then by
  construction, we have that $\sigma_j \in \stab_n(\SP(h_i)) \cap
  \stab_n(\SP(\sigma_i h_i)) \cap \stab_n(\SP(h_i')) \cap \stab_n(\SP(\sigma_i
  h_i'))$

  Define a graph $K$ with vertices given by the sets $S_i$ and an edge from
  $S_i$ to $S_j$ (with $i \neq j$) if $Q_i \cap S_j \neq \emptyset$. It follows
  then that if there is no edge between $S_i$ and $S_j$ then $(\sigma_i, h_i,
  h_i')$ and $(\sigma_j, h_j, h_j')$ are mutually independent. It remains to
  argue that $K$ has a large independent set. This is possible as the out-degree
  of $S_i$ in $K$ is bounded by
  \begin{align*}
    \vert Q_i \vert \leq \|SP(h_i) \| + \|SP(\sigma_i h_i) \| + \|SP(h_i') \| + \|SP(\sigma_i h_i') \leq 4 \cdot \frac{33\log s}{\epsilon \log n}
  \end{align*}.

  This follows as the sets $S_i$ are disjoint and we may apply Lemma
  \ref{lem:small-means-support} to each of the child gates. It follows that the
  average total degree (in + out degree) of $K$ is at most $2 \cdot \vert Q_i
  \vert \leq 34 \cdot k'$. Now greedily select a maximal independent set in $K$
  by repeatedly selecting $S_i$ with the lowest total degree and eliminating it
  and its neighbours. This action does not affect the bound on the average total
  degree of $K$ and hence determines an independent set $I$ in $K$ of size at
  least
  \begin{align*}
    \frac{\lfloor \frac{n}{k' + 2} \rfloor}{34k' + 1} \geq \frac{n - (k'+2)}{34k'+1k'+2} \geq \frac{n\frac{7}{16}}{34k'^2 + 69k' +2} \geq \frac{n}{(16k')^2}.
  \end{align*}

  Take $S = \{(\sigma, h, h') : S_i \in I \}$. Then from the above argument we
  have that $S$ is useful and independent.

  Moreover, from Claim \ref{claim:useful-independant-set}, we have that $s \geq
  \vert \orb(g) \vert \geq 2^{\vert S \vert} \geq 2^{\frac{n}{(16k')^2}}$ then
  $n^{1-\epsilon} \geq \log s \geq n \cdot (\frac{128}{\epsilon}\frac{\log
    s}{\log n})^{-2} > n \cdot (n^{1-\epsilon})^{-2} = n^{2\epsilon -1} \geq
  n^{1-\epsilon}$. This is a contradiction.
\end{proof}

\chapter{Translating Circuits into Formulas}
In this section we develop the theory for defining a formula of FPR from a
family of $P$-uniform family of symmetric circuits with rank gates. We first
show prove results on the generality of rigid circuits and circuits with
bijective labelings. Second, we prove that there are polynomial time algorithms
for determining canonical supports of gates. Third, we develop a FPR definable
way of evaluating a rank gate in the circuit. Finally, we complete the result by
explicitly defining an FPR formula corresponding to a given family of symmetric
circuits with rank gates.

We note that most of these results are for the more general case of symmetric
circuits with matrix-symmetric gates.

\section{Rigid Circuits and Labelings}

\begin{definition}
  We say that a circuit $C$ has bijective labels if for each gate $g$ in $C$,
  $L(g)$ is a bijection.
\end{definition}

\begin{lem}
  \label{lem:bij_labels}
  There is an algorithm that runs in polynomial time that takes in a circuit $C$
  and outputs a circuit with unique gates $C'$. Moreover, if $C$ was symmetric
  then $C'$ is symmetric. If $C$ is rigid then $C'$ is rigid.
\end{lem}

\begin{proof}
  Let $S = 2*\vert C \vert$. Recurse through the gates of $C$ topologically and
  let $h$ be the next gate topologically. If for all $g \in W(h, \cdot)$ we have
  that $L\vert (g)^{-1}(h) \vert = 1$ continue on to the next gate. If not add
  in a tower of $S$ $\and$ gates such that $h \rightarrow \and^h_1 \rightarrow
  \ldots \rightarrow \and^h_S$ (i.e. we have a tower of $\and$ gates with $h$ as
  input to $\and^h_1$ and the output of each $\and^h_i$ connected to the input
  of each $\and^h_{i+1}$ for each $1 \leq i < S$). Now for each $g \in W(h,
  \cdot)$, if $L(g)^{-1}(h) = \{ s_0, \ldots, s_{r}\}$, for each $1 \leq i \leq
  r$ add in the wires $W(\and^h_i, g)$ and set $L(g)(s_{i}) = \and^h_i$. Now
  continue on to the next gate topologically and run the above algorithm.

  % First, notice that for a given gate $h$ and $k \in \mathbb{N}$, we can
  % define
  % a sub-circuit $h^k = 1 \and \cdots, 1 \and h$ (or rather a $k$-height tower
  % of
  % $k$ binary $\and$ gates with $h$ at the top and all other inputs set to
  % $1$).
  % We call $h$ the top of the sub-circuit $h^k$ and the bottom $and$ gate (the
  % output of the sub-circuit) the bottom gate. We may think of this as a
  % $k$-height copy of the gate $h$, in the sense that it has the same output as
  % $h$ and similar orbits. We now use different copies to distinguish gates
  % that
  % otherwise have the same labelling.

  % Let $h$ be the next gate topologically. Then let $r$ be maximal such that
  % $W(h,g)$ and $\omega_g^{-1}(h) = \{ s^g_1, s^g_2, \ldots, s^g_r \}$. Then
  % let
  % $k$ be the height of the highest tower with $h$ at the top and with it's
  % bottom connected to a gate in $W(h, \cdot)$. Then create a single
  % $k+r-1$-height tower by adding in the appropriate number of binary $\and$
  % gates below $h$.

  % For each $g \in W(h, \cdot)$ and for each $\omega_g^{-1}(h) = \{s^g_1,
  % s^g_2,\ldots , s^g_{r_g}\}$, set $L'(g)(s^g_1) = h$ each child starting with
  % the $and$-gate child to $h$ add in a wire from from the $i$th gate in the
  % tower to $g$ and set $L'(g)(s^g_{i+1}) = \and_i$, where $\and_i$ is the
  % $i$th
  % $\and$ gate in the tower starting with the $\and$ gate in the tower child to
  % $h$. There are enough gates in the tower as $r_g \leq r$.

  We call this updated circuit $C'$.

  Firstly, note that after running the above algorithm for each $g$ the
  labelling of $g$ will be a bijection. Moreover, it's easy to see that the
  output of each gate remains unchanged, and as such the output of the circuit
  is unchanged.

  Secondly, notice that the size of $C'$ is at most $2*\vert C \vert^2$, and
  note that the above algorithm runs in polynomial time.

  Now suppose that $C$ is symmetric. Let $\sigma$ be a permutation on the input
  universe and $\pi_C$ the induced automorphism on $C$. We now define $\pi$, the
  induced automorphism on $C'$. For each gate $g$ in $C'$, if $g$ is in $C$ then
  set $\pi(g) := pi_C(g)$. If $g$ is not in $C$ then $g$ must be some
  $\and^h_i$, for some $h$ in $C$. Then set $\pi(\and^h_i) :=
  \and^{\pi_C(h)}_i$. It is easy to see that $\pi$ is an automorphism, and $\pi$
  extends $\sigma$.

  Suppose that $C$ is rigid. It is easy to see that $C'$ will be rigid as well.

\end{proof}

\begin{lem}
  There is an algorithm that runs in polynomial time that takes in a circuit $C$
  and outputs a circuit $C'$ such that $C'$ is rigid and has bijective labels.
  Moreover, if $C$ was symmetric it follows that $C'$ will be symmetric.
\end{lem}

\begin{proof}
  First run the algorithm from Lemma \ref{lem:bij_labels} on $C$, and call the
  output circuit $C$.

  Recurse through the gates of $C$ topologically. For each internal gate $g$,
  for all $g'$ in $C$ such that $g \neq g'$, $W(\cdot, g) = W(\cdot, g')$, $W(g,
  \cdot) = W(g', \cdot)$, $\Sigma(g) = \Sigma(g')$, $L(g) \sim L(g')$ and
  $\Omega^{-1}(g) = \Omega^{-1}(g')$, delete $g'$ and for all $s \in L^{-1}(g')$
  set $L(s) := g$.

  Now re-run the algorithm from Lemma \ref{lem:bij_labels} on $C$ and output the
  result.
\end{proof}

  \begin{lem}
    Let $C = \langle G, W, \Omega, \Sigma, \Lambda, L \rangle$ be a $(\SB, \MB,
    \tau)-circuit$ on structures of size $n$. There is a deterministic algorithm
    which runs in Poly($\vert C \vert$) and outputs a rigid $(\SB, \MB, \tau)
    circuit$ $C'$ such that $G' = G$ and for any $g \in G$, and any input
    $\tau$-structure $\mathcal{A}$ and any bijection $\gamma$ from $A$ to $[n]$,
    $C[\gamma \mathcal{A}](g) = C'[\gamma \mathcal{A}](g)$ and if $C$ is
    symmetric then so is $C'$.
  \end{lem}

  \begin{proof}

  \end{proof}

  \section{Computing Supports}

 \begin{lem}
   Let $C$ be a rigid $(\SB, MB, \tau)$-circuit on structures of size $n$ and
   $\sigma \in \sym_n$. There is a deterministic algorithm which runs in time
   Poly($\vert C \vert$) and outputs for each gate $g$ its image under the
   automorphism $\pi$ induced by $\sigma$, if it exists.
 \end{lem}
 \begin{proof}
   The proof proceeds by recursively going through the circuit and building the
   mapping $\pi$ induced by $\sigma$.

   Suppose $g$ is a constant gate, then $\pi g := g$. Suppose $g$ is a
   relational gate, then there is at most one gate $g'$ such that $\Sigma (g) =
   \Sigma (g')$ and $\sigma\Lambda (g') = \Lambda (g)$. If such a $g'$ exists
   assign $\pi g := g'$, else terminate with failure.

   If $g$ is an symmetric internal gate then (from rigidity) there is at most
   one gate $g'$ such that $\Sigma (g) = \Sigma(g')$ and $W_{g'} = \pi W_g$.
   Assign $\pi g := g'$ if such a gate exists, or else terminate with failure.

   If $g$ is a matrix-symmetric internal gate then consider the set of gates
   $g'$ such that $g'$ has children $\pi W_g$ and $\Sigma(g) = \Sigma(g')$, and
   let $A \times B = \dom (Sigma(g))$. If no such gate $g'$ exists, terminate
   with failure. Define $\sigma_{\pi, g'}:A \times B \rightarrow A \times B$ by
   $\sigma_{\pi, g'} = \omega^{-1}_{g'} \pi \omega_{g}$. Then clearly
   $\omega_{g'} \sigma_{\pi, g'} = \pi \omega_{g}$, and it's easy to show that
   $\pi \omega_g \sim \omega_{g'}$ iff $\sigma_{\pi,g'} \in \sym_A \times
   \sym_B$. But this just involves checking that $\sigma$ acts as a bijection on
   $A$ and $B$ separately and, given that $\vert A \vert$ and $\vert B \vert$
   are both bounded by $\vert C \vert$, the algorithm which just iterates
   through $A$ and $B$ is sufficient. If for every $g'$ it is found that
   $\pi_{\sigma,g'}$ is not in $\sym_A \times \sym_B$ then terminate with
   failure. If there is a $g'$ for which $\pi_{\sigma, g'} \in \sym_A \times
   \sym_B$ then it is unique by rigidity and so set $\pi g := g'$.

   If $g$ is an output gate, then check that for all tuples in $[n]^{q}$ we have
   that $\pi \Omega (x) = \Omega (\sigma (x))$, and terminate with failure if
   the condition is not met.

   If the algorithm has not terminated with failure, output the automorphism.

   The algorithm clearly runs in Poly($\vert C \vert$)
 \end{proof}

\section {Evaluating Circuits}
Let $\mathcal{C} = (C_n)_{n \in \mathbb{N}}$ be a family of polynomial-size
rigid symmetric circuits with matrix-symmetric gates. that compute a $q$-ary
query. Let $n_0$ the constant in the hypothesis of the Support Theorem.

Let $\mathcal{A}$ be a structure of size $n$ over the universe $U$, and let $g$
be a matrix-symmetric gate in $C_n$. In this section we develop the theory which
will allow us to evaluate this gate in FPR and (using results from Anderson and
Dawar \cite{}) recursively evaluate the circuit.

We recall that in order to evaluate the gate $g$ we need to consider a bijection
$\gamma: U \rightarrow [n]$, with the evaluation of $g$ given by $C_n[\gamma
\mathcal{A}](g)$. We first show that the evaluation of $g$ depends only on which
elements of $U$ are mapped to $\consp(g)$.

\begin{definition}
  Let $f: X \rightarrow Y$ and $g : X' \rightarrow Y'$. We say that $f$ is
  \emph{compatible} with $g$, or $f \sim g$, if for all $a \in S_1 \cap S_2$,
  $f(a) = g(a)$ and for all $a \in S_1 \setminus S_2$ and $b \in S_2 \setminus
  S_1$, $f(a) \neq g(b)$.
\end{definition}

\begin{definition}
  Let $f: A \times B \rightarrow H$ and $p: A' \times B' \rightarrow H$. We say
  that $f$ and $p$ are \emph{row-column equivalent} if there exist bijections
  $\alpha: A \rightarrow A'$ and $\beta: B \rightarrow B'$ such that for all
  $(a, b) \in A \times B$, $f(a,b)g(\alpha(a), \beta(b))$. In this case we write
  $f \sim p$.
\end{definition}

For a given bijection $\gamma: U \rightarrow [n]$ we can define a matrix
$L^{\gamma} : \dom (g) \rightarrow \{0,1\}$ by $L^{\gamma} (a,b) := C[\gamma
\mathcal{A}](L(g)(a,b))$. In the following Lemma we show that the the matrix
symmetry of the gate $g$ ensures that for any two bijections $\gamma_1,
\gamma_2$, if they agree on the support of $g$ then $L^{\gamma_1}_g$ is
row-column equivalent to $L^{\gamma_2}_g$. This implies that the evaluation of
$g$ depends only on the assignment to its support.

\begin{lem}
  Let $g$ be a matrix-symmetric gate in $C_n$ with children $H$. Let $\eta:
  \consp(g) \rightarrow U$ be an injection, and suppose $\gamma_1, \gamma_2: U
  \rightarrow [n]$ with $\gamma^{-1}_1 \sim \eta$ and $\eta \sim \gamma^{-1}_2$.
  Let $A \times B = \dom (L(g))$. Then $L^{\gamma_1}$ and $L^{\gamma_2}$ are
  row-column equivalent.
\end{lem}

\begin{proof}
  We have that there exists a unique $\pi \in \sym_n$ such that $\pi \gamma_1 =
  \gamma_2$. Moreover, since $\gamma^{-1}_1$ and $\gamma^{-1}_2$ are both
  consistent with $\alpha$, it follows that $\pi$ must fix $\consp(g)$. Thus
  $L(g) \sim \pi \cdot L(g)$, and so there exists $(\sigma, \lambda)$ such that
  $\pi L(g) = L(g) (\sigma, \lambda)$.

  We then have that,
  \begin{align*}
    L^{\gamma_1} (a,b) &= C_n[\gamma_1 \mathcal{A}](L(g)(a,b))\\
                       & = C_n[\pi \gamma_1 \mathcal{A}][\pi L(g)(a,b)] \\
                       & = C_n[\gamma_2 \mathcal{A}][L(g)((\sigma, \lambda)(a,b))]\\
                       & = L^{\gamma_2} ((\sigma, \lambda) (a,b)),
  \end{align*}
  and it follows that $L^{\gamma_1} \sim L^{\gamma_2}$.
\end{proof}

For each gate we define the set of all assignments to support of that gate which
cause the evaluate to true: $\EV_g := \{ \alpha: U^{\underline{\consp(g)}} :
\exists \gamma \in U^{\underline{[n]}} C_n [\gamma \mathcal{A}](g) = 1 \wedge
\alpha \sim \gamma^{-1}\}$. We now construct $\EV_g$ using $\EV_h$ for each $h
\in H_g$. In doing so we entirely characterise exactly which bijections $\gamma:
U \rightarrow [n]$ cause $g$ to evaluate to true.

% \begin{remark}
%   Let $\sigma \in \spstab{g}$, then we know that there exists $(\sigma_r,
%   \sigma_c) \in \sym_A \times \sym_B$ and $\sigma_r (i) = \row (\sigma h)$,
%   for any $h \in H$ such that $\row (h) = i$ and similarly $\sigma_c (j) =
%   \column (\sigma h)$, for any $h \in H$ such that $\column (h) = j$.
% \end{remark}

We introduce here some useful notation.

\begin{definition}
  Let $S^r_h = \{ \sigma \in \stab (\consp(g)) : \sigma \vec{r_h} = \vec{r_h}
  \}$ and $G^r_h = \{ \sigma \in \stab(\consp(g)) : \row(\sigma h) = \row(h)\}$.
  Let $S^c_h = \{ \sigma \in \stab (\consp(g)) : \sigma \vec{c_h} = \vec{c_h}
  \}$ and $G^c_h = \{ \sigma \in \stab(\consp(g)) : \column(\sigma h) =
  \column(h)\}$.
\end{definition}

Clearly, by definition of a row and column support, $S^r_h \subseteq G^r_h$ and
$S^c_h \subseteq G^c_h$.

% \begin{remark}
%   In this section I assume for the moment that:
%   \begin{itemize}
%   \item $\consp (g) = \{\}$,
%   \item $S^r_h = G^r_h$ and $S^c_h = G^c_h$,
%   \item $\forall h,h' \in H$ we have $\vert r_h \vert = \vert r_{h'} \vert$
%     and $\vert c_h \vert = \vert c_{h'} \vert$,
%   \item For all $i,i' \in A$ $\exists \sigma \in \stab(\consp (g))$ such that
%     $\sigma_r i = i'$. Similarly for all $j, j' \in B$ $\sigma \exists \in
%     \spstab{g}$ such that $\sigma_c j = j'$.
%   \item $\forall h \in H$, $\orb (h) = H$
%   \end{itemize}
% \end{remark}

% Let $k_r = \vert r_h \vert$ and $k_c = \vert c_h \vert$. Let $\mathcal{G}_r =
% \{\sigma_r : \sigma \in \spstab{g}\}$ and $\mathcal{G}_c = \{sigma_c : \sigma
% \in \spstab{g}\}$.

% \begin{definition}
%   Let $\mathcal{G} \leq \sym_n$ and $x$ be an element of a set on which an
%   action of $\sym_n$ is defined. Then $\orb_{\mathcal{G}}(x) = \{x^\pi: \pi
%   \in \mathcal{G}\}$.

%   For $i \in [a]$ let $\orb_r(i) = \orb_{\mathcal{G}_r}(i)$ and for $j \in
%   [b]$ let $\orb_c (j) = \orb_{\mathcal{G}_c}(j)$.
% \end{definition}

% \begin{definition}
%   Let $A, B \subseteq [n]$ and let $t = (t_1, t_2, t_3) \in \mathbb{N}^3$. We
%   say that $(A,B)$ has \emph{type} $t$ if $\vert A \vert = t_1$, $\vert B
%   \vert = t_2$ and $\vert A \cap B \vert = t_{3}$. We say that $\type (A,B) =
%   t$.
% \end{definition}

For the sake of brevity, for sets $Z,X,Y$, where $f : Z \rightarrow X$ and $p :
Z \rightarrow Y$ are injections, let $f_p = f \cdot p^{-1}$.

\begin{definition}
  Let $\vec{x}, \vec{y} : Z \rightarrow X$ and $\vec{r}, \vec{c}: Z \rightarrow
  Y$ be injections. If $\vec{x} \cdot \vec{r}^{-1} \sim \vec{y} \cdot
  \vec{c}^{-1}$ we say that $(\vec{x}, \vec{y})$ and $(\vec{r} \vec{c})$ have
  the same \emph{type}.
\end{definition}

We note that if we have $f : X \rightarrow Y$ and $p: X' \rightarrow Y'$ and $f
\sim p$ then define $(f \vert p): X \cup X' \rightarrow Y \cup Y'$ by

\begin{align*}
  (f \vert p) (x) =
  \begin{cases}
    f (x) & x \in X \\
    p (x) & x \in Y.
  \end{cases}
\end{align*}

The compatibility of these two functions ensures that this function is well
defined.

% \begin{claim}
%   Let $A_1,B_1, A_2, B_2 \subseteq [n]$. Then $(A_1, B_1)$ and $(A_2, B_2)$
%   have the same type iff $\exists \pi \in \sym_n$ such that $A_1 = A^\pi_2$
%   and $B_1 = B^\pi_2$.
% \end{claim}

% \begin{definition}
%   Let $f: A \rightarrow S$ and $g: B \rightarrow S$ be injections, with $A$
%   and $B$ being finite sets. Then $(f,g)$ has \emph{type} $t= (t_1 , t_2, t_3)
%   \in \mathbb{N}^3$ if $\vert A \vert = t_1$ and $\vert B \vert = t_2$ $\vert
%   \{i \in A \cap B : f(i) = g(i)\} \vert = t_3$. We say that $\type (f,g) =
%   t$.
% \end{definition}

% \begin{definition}
%   Let $h \in H$ the \emph{type} of $h$ is the type of $(r_h, c_h)$. We denote
%   the type of $h$ by $\type (h)$.
% \end{definition}

We now fix an injection $\eta : \consp(g) \rightarrow U$ and describe an
approach for determining if $\eta \in EV_g$.

Let $X$ be a set on which the left group action on $\sym_n$ is defined and $s
\in X$. We denote $\vec{x}$ then $A_s = \{\vec{x} \in U^{\underline{[\vert
    \sp(s) \vert ]}} : \eta \sim \vec{x} \cdot \vec{\consp}(s)^{-1} \}$, where
$\vec{\consp}(s): [\vert \consp(s) \vert] \rightarrow \consp(s)$ maps $i \in
[\vert \consp(s) \vert]$ to the $i$th element of $\consp(s)$. For $\vec{x} \in
A_s$, we use $\vec{x}_s$ from here forward to denote $\vec{x} \cdot
\vec{\consp(s)^{-1}}$.

% Let $L(g)(i,j) = h \in H$, $\vec{x} \in A_i$ and $\vec{y} \in A_j$. Let
% $\vec{r} \in ^{\underline{\consp(i)}}$ and $\vec{c}
% \in\consp{j}^{\underline(\consp(j))}$. Suppose $(\vec{r}, \vec{c})$ has the
% type of $(\vec{x}, \vec{y})$. Then we can define $(\vec{x}\vert \vec{y}):
% \consp(i) \cup \consp(j) \rightarrow U$ by
% \begin{align*}
%   (\vec{x}_{\vec{r}} \vert \vec{y}_{\vec{c}}) (z) =
%   \begin{cases}
%     \vec{x} (\vec{r}^{-1}(z)) \text{ if } z \in \consp(i) \\
%     \vec{y} (\vec{r}^{-1}(z)) \text{ if } z \in \consp(j). \\
%   \end{cases}
% \end{align*}

This mapping allows us to combine two assignments into a single assignment. This
will be useful when combining the row and column supports of a gate.

% \begin{claim}
%   Let $f: A \rightarrow S$ and $g B \rightarrow S$ be injections. Then $(f,g)$
%   has the same type as $(A,B)$. It follows that if $h \in H$, and $(i,j) =
%   L^{-1}(h)$,with $f \in A^r_i$ and $ g \in A^c_j$ then $h$ has the same type
%   as $(f,g)$.
% \end{claim}

% For $(i,j) \in [a] \times [b]$ let $H_{i,j} = \{L(p,q): (p,q) \in \orb_r(i)
% \times \orb_c(j)]\}$.

% \begin{claim}
%   For all $(i', j') \in \orb_r(i) \times \orb_c(j)$ and $\sigma \in
%   \spstab{g}$, $(\sigma_r i', \sigma_cj') \in \orb_r(i) \times \orb_c(j)$.
% \end{claim}

% \begin{claim}
%   For all $(i,j) \in [a] \times [b]$, $H_{i,j}$ is a union of orbits.
% \end{claim}

% \begin{lem}
%   Let $(i,j) \in [a] \times [b]$. If $h, h' \in H_{i,j}$ and $h, h'$ are the
%   same type then $h' \in \orb(h)$.
% \end{lem}

We now define a matrix that we later be shown to be definable in FPR. We then
show that this matrix is row-column equivalent to $L^{\gamma}$, for any
bijection $\gamma: U \rightarrow [n]$, and hence allowing us to determine the
rank of the matrix input to $g$ in FPR.

Let $R^{\min} = \{\min (\orb(\row(h))) : h \in H\}$ and $C^{\min} = \{ \min
(\orb (\column(h))) : h \in H\}$ and let
\begin{align*}
  I = \{(i, \vec{x}): i \in R^{\min}, \vec{x} \in A_i\},
\end{align*}
and
\begin{align*}
  J = \{(j, \vec{y}): j \in C^{\min}, \vec{y} \in A_j \}.
\end{align*}

% \begin{align*}
%   M ((i, \vec{x}), (j, \vec{y})) := \bigvee_{t \in \types} ((\vec{x},
%   \vec{y}) \text{ has type } t) \land (\vec{x} | \vec{y}) \in \EV_{\mu_{i,j}(t)}.
% \end{align*}

Let $(i, \vec{x}) \in I$ and $(j, \vec{y}) \in J$.

Let $u_1 , \ldots , u_{2k}$ be the first $2k$ elements of $[n] \setminus
\consp(g)$ in the natural order on the structure. Then let
\[r = \eta^{-1} (\img(\vec{x}) \cap \eta (\consp(g))) \cup
  \{u_{\vec{x}^{-1}(a)}: a \in \img(x) \setminus \eta (\consp (g))\} \] and
\[s = \eta^{-1} (y \cap \eta (\consp(g))) \cup (x \cap y) \cup \{ u_{k +
    \vec{x}^{-1}(a)} : a \in y \setminus (x \cup \eta (\consp (g))) \}). \]
Define
\[
  \vec{r} (a) =
  \begin{cases}
    \eta^{-1} (\vec{x} (a)) & a \in \vec{x}^{-1} (x \cap \eta (\consp(g))) \\
    u_{a} & a \in \vec{x}^{-1} (x \setminus \eta(\consp(g)),
  \end{cases}
\]
and
\[
  \vec{c} (a) =
  \begin{cases}
    \eta^{-1} (\vec{y} (a)) & a \in \vec{y}^{-1}(y \cap \eta (\consp (g))) \\
    \vec{r} (a) & a \in \vec{y}^{-1} (x \cap y \setminus \eta (\consp (g))) \\
    u_{k+a} & \text{otherwise}.
  \end{cases}
\]

\begin{lem}
  \label{lem:permutation_row-column}
  There exists $\sigma_1, \sigma_2 \in \spstab{g}$ such that $\sigma_1 \cdot
  \vec{\consp(i)} = \vec{r}$ and $\sigma_2 \cdot \vec{\consp(j)} = \vec{c}$.
\end{lem}
\begin{proof}
  To be added from the book
\end{proof}

Let $h = L(g)(\sigma_1(i), \sigma_2 (j))$. Then we can define the matrix $M : I
\times J \rightarrow \{0,1\}$ by

\begin{align*}
  M((i , \vec{x}), (j, \vec{y})) := (\vec{x}_{\vec{r}} \vert \vec{y}_{\vec{c}}) \in EV_h.
\end{align*}

It remains to show that $M \sim L^{\gamma}$ for some (and so all) bijections
$\gamma: U \rightarrow [n]$.
% \begin{definition}
%   Let $M_1$ and $M_2$ be matrices over some field, with row and column indexes
%   given by $(A_1, B_1)$ and $(A_2, B_2)$, respectively. We say that $M_1$ and
%   $M_2$ are row-column equivalent iff there exists bijections $\alpha: A_1
%   \rightarrow A_2$ and $\beta: B_1 \rightarrow B_2$ such that for all $(a,b)
%   \in A_1 \times B_1$ we have that $M_1(a,b) = M_2 (\alpha (a), \beta (b))$.
%   If $M_1$ and $M_2$ are row-column equivalent we say that $M_1 \sim M_2$.
% \end{definition}

% Fix $\gamma: U \rightarrow [n]$ such that $\gamma^{-1} \sim \eta$. We define
% $L^\gamma: A \times B \rightarrow \{0,1\}$ by $L^\gamma(a,b) = C[\gamma
% \mathcal{A}](L (a,b))$.
% \\~\\
% We note that, in fact, it is the choice of the assignment to the support of
% $g$, i.e. $\eta$, that really matters in the sense that for any two global
% assignments that agree on the support of $g$ will produce row-column
% equivalent matrices. The following lemma formalises this observation.

% \begin{lem}
%   Let $g$ be a matrix-symmetric gate in $C_n$ with children $H$ and matrix
%   labelling $L(g)$. Let $\eta \in U^{\consp(g)}$, and suppose $\gamma_1,
%   \gamma_2: U \rightarrow [n]$ with $\gamma^{-1}_1 \sim \alpha$ and $
%   \gamma^{-1}_2 \sim \alpha$. Let $A \times B = \dom (L(g))$.

%   Then for an input structure $\mathcal{A}$, $L^{\gamma_1} \sim L^{\gamma_2}$.
% \end{lem}
% \begin{proof}
%   We have that there exists a unique $\pi \in \sym_n$ such that $\pi \gamma_1
%   = \gamma_2$. Moreover, since $\gamma^{-1}_1$ and $\gamma^{-1}_2$ are both
%   consistent with $\eta$, it follows that $\pi$ must fix $\consp(g)$. Thus
%   $L(g) \sim \pi L(g)$, and so there exists $(\alpha, \beta)$ such that $\pi
%   L(g) = L(g) \cdot (\alpha, \beta)$.

%   We then have that,
%   \begin{align*}
%     L^{\gamma_1} (a,b) &= C_n[\gamma_1 \mathcal{A}](L(g)(a,b))\\
%                        & = C_n[\pi \gamma_1 \mathcal{A}][\pi L(g)(a,b)] \\
%                        & = C_n[\gamma_2 \mathcal{A}][L(g)((\alpha, \beta)(a,b))]\\
%                        & = L^{\gamma_2} ((\alpha, \beta) (a,b)),
%   \end{align*}
%   and it follows that $L^{\gamma_1} \sim L^{\gamma_2}$.
% \end{proof}

\begin{claim}
  Let $\sigma \in \spstab{g}$, $\eta \in U^{\underline{\consp(g)}}$, $\gamma: U
  \rightarrow [n]$ a bijection such that $\gamma^{-1} \sim \eta$. Then
  $\gamma^{-1} \cdot \sigma \sim \eta$.
\end{claim}
\begin{proof}
  Suppose $a \in \consp(g)$, then $\sigma (a) = a$ and so $\gamma^{-1} (\sigma
  (a)) = \gamma^{-1} (a) = \eta (a)$.
\end{proof}

% \begin{claim}
%   For $h \in H$ and $\sigma \in \stab(\consp (g))$, we have that
%   $\vec{r}_{\sigma h_1} = \sigma \vec{r}_{h_1}$.
% \end{claim}
% \begin{proof}
%   Proof in book
% \end{proof}

The following result shows that the action on the support of an object
determines the action on that object.

\begin{lem}
  \label{lem:support_determine_action}
  Let $\sigma, \sigma' \in \spstab{g}$, $a$ be an object on which on the action
  of these two permutations is defined, and $\consp(a) \subset [n]$ be a support
  of $a$. Then if $\sigma (\vec{\consp}(a)) = \sigma' (\vec{\consp}(a))$ then
  $\sigma (a) = \sigma' (a)$.
\end{lem}
\begin{proof}
  From $\sigma (\vec{\consp(a)}) = \sigma' (\vec{\consp(a)})$, it follows that
  $\pi = (\sigma')^{-1} \sigma$ fixes $\vec{\consp}(a)$. Thus $\sigma (a) =
  \sigma' (\pi (a)) = \sigma' (a)$.
\end{proof}

% \begin{lem}
%   \label{lem:map_same_support_same_row}
%   Let $\sigma, \sigma' \in \spstab{g}$ and suppose for some $i \in [a]$ we
%   have that $\sigma (\vec{r}_i) = \sigma (\vec{r}_i)$. It follows that
%   $\sigma_r (i) = \sigma'_r (i)$.
% \end{lem}
% \begin{proof}
%   Let $pi = (\sigma^{-1} \sigma')_r \in G^r_i$. Then $ (\sigma^{-1} \sigma')_r
%   (i) = \sigma^{-1}_r \sigma'_r (i) = i$, and so $\sigma_r (i) = \sigma'_r
%   (i)$.
% \end{proof}

% Let $X$ be a set on which the left group action on which the left group action
% of $\sym_n$ is defined and let $a \in X$. Let $\vec{x} \in A_i$ and $f \in
% U^{\underline{[\vert \consp{a} \vert]}}$. Then $\Pi^{\gamma}_{\vec{x}_f} (a)$
% is the action on $a$ defined by $\Pi^{\gamma}_{\vec{x}_f} (z) =
% \gamma(\vec{x_f}(z))$, for all $z \in \consp (a)$. Lemma
% \ref{lem:support_determine_action} tells us that this action is well defined.

% Let $(i, j) \in A \times B$, let $\vec{x} \in A^r_i$ and $\vec{y} \in A^c_j$.
% Define permutations $\Pi^{\gamma}_{\vec{x}}$ and $\Pi^{{\gamma},c}_{\vec{y}}$
% such that $\Pi^{\gamma,r}_{\vec{x}}(\vec{r}_i) = \gamma (\vec{x})$ and
% $\Pi^{\gamma,c}_{\vec{y}}(\vec{c}_j) = \gamma (\vec{y})$. From Lemma
% \ref{lem:support_determine_action} we have that the choice of $\vec{x}$ and
% $\vec{y}$ uniquely determine the mappings $\Pi^{\gamma,r}_{\vec{x}}(i)$ and
% $\Pi^{\gamma,c}_{\vec{y}}(j)$

Let $\alpha^{\gamma}: I \rightarrow A$ and $\beta^{\gamma}: J \rightarrow B$ be
defined by $\alpha^{\gamma} (i, \vec{x}) = \Pi^{\gamma}_{\vec{x}_{i}}(i)$ and
$\beta^{\gamma} (j, \vec{y}) = \Pi^{\gamma}_{\vec{y}_{j}}(j)$.

We now show that $\alpha^{\gamma}$ and $\beta^{\gamma}$ act as witnesses to the
row-column of equivalence of $M$ and $L^{\gamma}$. The following lemma proves
surjectivity.

\begin{lem} 
  For any bijection $\gamma : U \rightarrow [n]$ both $\alpha^{\delta}$ and
  $\beta^{\delta}$ are surjective.
\end{lem}
\begin{proof}
  We show that $\alpha$ is surjective, with the same result for $\beta$
  following similarly. Let $q \in A$ and let $i = \min (\orb (q))$. Then there
  exists $\sigma \in \spstab{g}$ such that $\sigma i = q$. Let $\vec{x} =
  \gamma^{-1} (\sigma(\vec{\consp}(i)))$. Notice that for $a \in \consp(i)$ we
  have that $\vec{x}_i(a) = \gamma^{-1} (\sigma (a))$, and since $\gamma^{-1}
  \cdot \sigma \sim \eta$, it follows that $\vec{x} \in A_i$.

  For $a \in \consp(i)$ we have $\Pi^{\gamma}_{\vec{x}_i} (a) = \gamma
  (\vec{x}_i(a)) = \gamma \cdot \gamma^{-1} \sigma (a) = \sigma (a)$. From Lemma
  \ref{lem:support_determine_action} it follows that $\alpha(i, \vec{x}) = q$.
\end{proof}

The following lemma allows us to factor through permutations.
\begin{lem}
  \label{lem:alpha_and_gamma}
  Let $(i,\vec{x}) \in I$ and $(j, \vec{y}) \in J$. Let $\gamma: U \rightarrow
  [n]$ be a bijection such that $\gamma^{-1} \sim \eta$ and $\pi \in spstab{g}$.
  Then $\pi \alpha^{\gamma}(i, \vec{x}) = \alpha^{\pi \gamma}(i, \vec{x})$ and
  $\pi \beta^{\gamma}(j, \vec{y}) = \beta^{\pi \gamma}(j, \vec{y})$.
\end{lem}
\begin{proof}
  We have that $\pi \alpha^{\gamma}(i, \vec{x}) = \pi
  \Pi^{\gamma}_{\vec{x}_i}(i)$ and $(\pi
  \Pi^{\gamma}_{\vec{x}_{i}}(\vec{\consp}(i)) = \pi \cdot \gamma (\vec{x}) =
  \Pi^{\pi \gamma}_{\vec{x}_i}(\vec{\consp}(i))$. Since
  $\Pi^{\gamma}_{\vec{x}_i}$ and $\Pi^{\pi \gamma}_{\vec{x}_i}$ are in
  $\spstab{g}$, it follows from Lemma \ref{lem:support_determine_action}, that
  $\pi \alpha^{\gamma}(i, \vec{x}) = \pi \Pi^{\gamma}_{\vec{x}_i} (i) = \Pi^{\pi
    \gamma}_{\vec{x}_i}(i) = \alpha^{\pi \gamma}(i, \vec{x})$. Similarly, $\pi
  \beta^{\gamma}(j, \vec{y}) = \beta^{\pi \gamma} (j, \vec{y})$.
\end{proof}

\begin{lem}
  \label{lem:alpha_ind_gamma}
  Let $(i,\vec{x}) \in I$ and $(j, \vec{y}) \in J$. Let $\gamma_1, \gamma_2: U
  \rightarrow [n]$ be bijections such that $\gamma^{-1}_1 \sim \eta$ and
  $\gamma^{-1}_2 \sim \eta$. Let $\mathcal{A}$ be a structure. Then
  $C_n[\gamma_1 \mathcal{A}] (L(\alpha^{\gamma_1}(i, \vec{x}),
  \beta^{\gamma_1}(j, \vec{y}))) = C_n[\gamma_2 \mathcal{A}]
  (L(\alpha^{\gamma_2}(i, \vec{x}), \beta^{\gamma_2}(j, \vec{y})))$.
\end{lem}
\begin{proof}
  We note that there exists $\pi \in \sym_n$ such that $\gamma_1 = \pi
  \gamma_2$. Moreover, since $\gamma^{-1}_1$ and $\gamma^{-1}_2$ are both
  consistent with $\eta$, it follows that $\pi \in \spstab{g}$. We then have
  that
  \begin{align*}
    C_n[\gamma_1 \mathcal{A}](L(\alpha^{\gamma_1}(i, \vec{x}), \beta^{\gamma_1}(j,
    \vec{y})) &= C_n[\pi \gamma_1 \mathcal{A}](\pi L(\alpha^{\gamma_1}(i, \vec{x}),
                \beta^{\gamma_1}(j, \vec{y})) \\
              &= C_n[\pi \gamma_1 \mathcal{A}](L(\pi
                \alpha^{\gamma_1}(i, \vec{x}), \pi \beta^{\gamma_1}(j, \vec{y}))\\
              &= C_n[\pi
                \gamma_1 \mathcal{A}](L(\alpha^{\pi \gamma_1}(i, \vec{x}), \pi \beta^{\pi
                \gamma_1}(j, \vec{y})\\
              &= C_n[\gamma_2 \mathcal] (L(\alpha^{\gamma_2}(i,
                \vec{x}), \beta^{\gamma_2}(j, \vec{y})))\\
  \end{align*}The third equality follows from Lemma \ref{lem:alpha_and_gamma}.
\end{proof}

\begin{lem}
  \label{lem:defining_h_from_IJ}
  Let $(i, \vec{x}) \in I$ and $j, \vec{y} \in J$. Let $\vec{r}$ and $\vec{c}$
  be described as above. Let $\sigma_1$ and $\sigma_2$ be as from Lemma
  \ref{lem:permutation_row-colum}. Let $h = L(g) (\sigma_1 (i), \sigma_2 (j))$.
  Then $\alpha^{\gamma'} (i, \vec{x}) = \row (h)$ and $\beta^{\gamma'}
  (i,\vec{y}) = \column(h)$.
\end{lem}
\begin{proof}
  $\alpha^{\gamma'}(i, \vec{x}) = \Pi^{\gamma'}_{\vec{x}_i} (i)$. It is
  sufficient to show that for all $a \in sp(g)$, $\Pi^{\gamma}_{\vec{x}_i} (a) =
  \sigma_1 (a)$. Note that $\Pi^{\gamma'}_{\vec{x}_i} (a) =
  (\Pi^{\gamma}_{(\vec{x}_{\vec{r}} \vert \vec{y}_{\vec{c}})})^{-1} \gamma
  (\vec{x}_i (a))$, and if $b = \sigma_1 (a)$ we have that $\gamma (\vec{x}_i
  (a)) = \gamma (\vec{x} (\vec{\consp}(i)^{-1} (a))) = \gamma (\vec{x}
  (\vec{\consp}(i)^{-1} (\sigma^{-1}_1 b))) = \gamma (\vec{x} ((\sigma_1 \cdot
  \vec{\consp}(i))^{-1} (b))) = \gamma (\vec{x} (\vec{r}^{-1}(b))) =
  \Pi^{\gamma}_{(\vec{x}_{\vec{r}} \vert \vec{y}_{\vec{c}})}(b)$. So
  $\Pi^{\gamma'}_{\vec{x}} (a) = b = \sigma_1 (a)$. Similarly for
  $\beta^{\gamma'}$.
\end{proof}

\begin{thm}
  Let $\delta: U \rightarrow [n]$, $M$ is row-column equivalent to $L^{\gamma}$.
  This equivalence is witnessed by $\alpha^{\gamma}$ and $\beta^{\gamma}$.
\end{thm}
\begin{proof}
  We have that $\alpha^{\gamma}$ and $\beta^{\gamma}$ are surjective. Let
  $\gamma' = (\Pi^{\gamma}_{(\vec{x}_{\vec{r}} \vert \vec{y}_{\vec{c}})})^{-1}
  \gamma$.

  \begin{align*}
    M((i, \vec{x}), (j, \vec{y}))
    &= (\vec{x}_{\vec{r}} \vert \vec{y}_{\vec{c}}) \in \EV_h \\
    &= C_n[\gamma \mathcal{A}] (\Pi^{\gamma}_{(\vec{x}_{\vec{r}} \vert \vec{y}_{\vec{c}})} (h)) \\
    &= C_n[\gamma' \mathcal{A}] (h) \\
    &= C_n[\gamma' \mathcal{A}](L(\row(h), \column (h)))\\
    &= C_n[\gamma' \mathcal{A}](L(\alpha^{\gamma'}(i, \vec{x}), \beta^{\gamma'}(j, \vec{y})))\\
    &= C_n[\gamma \mathcal{A}](L(\alpha^{\gamma}(i, \vec{x}), \beta^{\gamma}(j, \vec{y}))y)\\
    &= L^{\gamma}(\alpha^{\gamma}(i, \vec{x}), \beta^{\gamma}(j, \vec{y}))
  \end{align*}
  The second equality follows from Lemma \ref{}. The fifth equality follows from
  Lemma \ref{lem:defining_h_from_IJ}. The sixth equality follows from Lemma
  \ref{lem:alpha_ind_gamma}.

\end{proof}

\section{FPR Formulas}

% \begin{lem}
%   Let $(i, \vec{x}) \in I$ and $(j, \vec{y}) \in J$. Let $h \in H$ such that
%   $\row (h) \in \orb_r (i)$, $\column (h) \in \orb_c(j)$, and $\type(h) =
%   \type(\vec{x}, \vec{y})$. Let $\gamma: U \rightarrow [n]$ be a bijection and
%   $\mathcal{A}$ a structure. Then $C_n[\gamma
%   \mathcal{A}](\Pi^{\gamma}_{\vec{x} \vert \vec{y}} (h)) = C_n[\gamma
%   \mathcal{A}](L(\alpha^{\gamma}(i, \vec{x}), \beta^{\gamma}(j, \vec{y})))$.
% \end{lem}
% \begin{proof}
%   Let $\gamma' = (\Pi^{\gamma}_{\vec{x} \vert \vec{y}})^{-1} \gamma$. First we
%   show that $h = L(\alpha^{\gamma'}(i, \vec{x}), \beta^{\gamma'}(j,
%   \vec{y}))$. Notice that $\Pi^{\gamma'}_{\vec{x}}(\vec{r}_i) =
%   \gamma'(\vec{x}) = (\Pi^{\gamma}_{\vec{x} \vert \vec{y}})^{-1} \gamma
%   (\vec{x}) = \vec{r}_h$. But $\row(h) \in \orb_r(i)$ and so there exists
%   $\vec{x}' \in A^r_i$ such that $\Pi^{\gamma'}_{vec{x}'}(i) = \row{h}$ From
%   Lemma \ref{lem:support_determine_action} it follows that $\row (h) =
%   \Pi^{\gamma'}_{\vec{x}}(i) = \alpha^{\gamma'}(i, \vec{x})$. Similarly, we
%   can show that $\column(h) = \beta^{\gamma'}(j, \vec{y})$. It follows that $h
%   = L(\alpha^{\gamma'}(i, \vec{x}), \beta^{\gamma'}(j, \vec{y}))$, and so
%   \begin{align*}
%     C_n[\gamma \mathcal{A}](\Pi^{\gamma}_{\vec{x}
%     \vert \vec{y}} (h))
%     &= C_n[\gamma' \mathcal{A}](h)\\
%     &= C_n[\gamma' \mathcal{A}](L(\alpha^{\gamma'}(i, \vec{x}), \beta^{\gamma'}(j, \vec{y})))\\
%     &= C_n[\gamma \mathcal{A}](L(\alpha^{\gamma}(i, \vec{x}), \beta^{\gamma}(j, \vec{y}))).
%   \end{align*}
%   The final equality follows from Lemma \ref{lem:alpha_ind_gamma}.
% \end{proof}

% \chapter{New Stuff}
% Let $x,y \subset U$ such that $\vert x \vert = \vert y \vert = k \in
% \mathbb{N}$. Let $\vec{x}: [k] \rightarrow x$ and $\vec{y}: [k] \rightarrow y$
% be bijections.

% We need to define $r, c \subset [n]$ such that $\vert r \vert = \vert c \vert
% = k$ and there exists bijections $\vec {r}: [k] \rightarrow r$ and $\vec{c}:
% [k] \rightarrow c$.

% Let $u_1 , \ldots , u_{2k}$ be the first $2k$ elements of $[n] \setminus
% \SP(g)$ in order. Then let
% \[r = \eta^{-1} (x \cap \eta (\SP(g))) \cup \{u_{\vec{x}^{-1}(a)}: a \in x
%   \setminus \alpha (\SP (g))\} \] and
% \[s = \eta^{-1} (y \cap \eta (\SP(g))) \cup (x \cap y) \cup \{ u_{k +
%   \vec{x}^{-1}(a)} : a \in y \setminus (x \cup \alpha (\SP (g))) \}). \]
% Define
% \[
%   \vec{r} (a) =
%   \begin{cases}
%     \eta^{-1} (\vec{x} (a)) & a \in \vec{x}^{-1} (x \cap \eta (\SP(g))) \\
%     u_{a} & a \in \vec{x}^{-1} (x \setminus \eta(\SP(g)),
%   \end{cases}
% \]
% and

% \[
%   \vec{c} (a) =
%   \begin{cases}
%     \eta^{-1} (\vec{y} (a)) & a \in \vec{y}^{-1}(y \cap \eta (\SP (g))) \\
%     \vec{r} (a) & a \in \vec{y}^{-1} (x \cap y \setminus \eta (\SP (g))) \\
%     u_{k+a} & \text{otherwise}.
%   \end{cases}
% \]

% \begin{lem}
%   $x_r \sim \eta$ and $x_c \sim \eta$
% \end{lem}

% \begin{lem}
%   $\SP(g) \cap \SP(i) = \SP(g) \cap r$ and $\SP (g) \cap SP (j) = \SP (g) \cap
%   c$.
% \end{lem}

% Let $(\vec{x}, \vec{y})_{\vec{r}, \vec{c}}: r \cup c \rightarrow U$ be defined
% by
% \[
%   (\vec{x}, \vec{y})_{\vec{r}, \vec{c}}(a) =
%   \begin{cases}
%     \vec{x}(\vec{r}^{-1}(a)) & a \in r \\
%     \vec{y}(\vec{c}^{-1}(a)) & a \in c
%   \end{cases}
% \]

% This function is well defined.

% % \begin{lem}
% %   Let $r, c \subset U$ and let $r_1, r_2 : [k_1] \rightarrow r$ and $c_1, c_2 :
% %   [k_2] \rightarrow c$ then $\type(\vec{r}_1, \vec{c}_1) = \type (\vec{r}_2 ,
% %   \vec{c}_2)$.
% % \end{lem}

% \begin{lem}
%   $(\vec{x}_{\vec{r}} \vert \vec{y}_{\vec{y}}) \sim \eta$
% \end{lem}

% \begin{lem}
%   There exists $\sigma_1, \sigma_2 \in \spstab{g}$ such that $\sigma_1 \cdot
%   \vec{\consp(i)} = \vec{r}$ and $\sigma_2 \cdot \vec{\consp(j)} = \vec{c}$.
% \end{lem}

% \begin{lem}
%   Let $(i, \vec{x}) \in I$ and $j, \vec{y} \in J$. Let $\vec{r}$ and $\vec{c}$
%   be described as above. Let $\sigma_1$ and $\sigma_2$ be as from Lemma
%   \ref{}. Let $h = L(g) (\sigma_1 (i), \sigma_2 (j))$. Then $\alpha^{\gamma'}
%   (i, \vec{x}) = \row (h)$ and $\beta^{\gamma'} (i,\vec{y}) = \column(h)$.
% \end{lem}
% \begin{proof}
%   $\alpha^{\gamma'}(i, \vec{x}) = \Pi^{\gamma'}_{\vec{x}} (i)$. It is
%   sufficient to show that for all $a \in sp(g)$, $\Pi^{\gamma}_{\vec{x}} (a) =
%   \sigma_1 (a)$. Note that $\Pi^{\gamma'}_{\vec{x}_i} (a) =
%   (\Pi^{\gamma}_{(\vec{x}, \vec{y})_{\vec{r}, \vec{c}}})^{-1} \gamma
%   (\vec{x}_i (a))$, and if $b = \sigma_1 (a)$ we have that $\gamma (\vec{x}_i
%   (a)) = \gamma (\vec{x} (\vec{\consp}(i)^{-1} (a))) = \gamma (\vec{x}
%   (\vec{\consp}(i)^{-1} (\sigma^{-1}_1 b))) = \gamma (\vec{x} ((\sigma_1 \cdot
%   \vec{\consp}(i))^{-1} (b))) = \gamma (\vec{x} (\vec{r}^{-1}(b))) =
%   \Pi^{\gamma}_{(\vec{x}, \vec{y})_{\vec{r}, \vec{c}}}(b)$. So
%   $\Pi^{\gamma'}_{\vec{x}} (a) = b = \sigma_1 (a)$. Similarly for
%   $\beta^{\gamma'}$.
% \end{proof}

% \begin{lem}
%   Let $(i, \vec{x}) \in I$ and $(j, \vec{y}) \in J$. There exists $h \in H$
%   such that $h$ has type $(\vec{x}, \vec{y})$ and for any bijection $\gamma: U
%   \rightarrow [n]$ and structure $\mathcal{A}$ then $C_n[\gamma \mathcal{A}]
%   (\Pi^{\gamma}_{(\vec{x} \vert \vec{y})_{h}}) = C_n [\gamma
%   \mathcal{A}](L(\alpha^{\gamma}(i, \vec{x}), \beta^{\gamma}(j, \vec{y})))$.
% \end{lem}

% \begin{lem}
%   Let $h \in H$ and $\vec{z} \in A_h$ and $\Pi^{\delta}_h$ be a permutation
%   such that $\Pi^{\delta}_{\vec{z}}(a) = \delta(\vec{z}(a))$ for all $a \in
%   \sp(h)$. Then $C_n[\gamma \mathcal{A}](\Pi^{delta}_{\vec{z}} (h))$ iff
%   $\vec{z} \in \EV_h$.
% \end{lem}

% \begin{thm}
%   The matrix $M$ is equivalent to $L$.
% \end{thm}

\chapter{Concluding Remarks}
\bibliographystyle{plain} \bibliography{references}
\end{document}