### 2017-07-26 sections/introduction.tex

\documentclass[../paper.tex]{subfiles}
\begin{document}
% A finite relational structure, such as a graph, can be represented as a binary
% string in a standard way (for instance, by enumerating its adjacency matrix).
% When we speak of the computational complexity of properties of graphs, we assume
% such an encoding in order to present the structure as input to a machine or
% circuit or other appropriate computational model. The properties of structures
% we are interested in studying are invariant under isomorphisms. This requirement
% translates into an invariance condition on sets of strings---say that a language
% $L \subseteq \{0,1\}^\star$ is a \emph{graph property} if, whenever $x \in L$
% for a string $x$ representing a graph $G$, we have $y\in L$ for any string $y$
% obtained from $x$ by re-ordering the vertices of $G$. The central open question
% in descriptive complexity---whether there is a logical characterisation of the
% polynomial-time decidable properties of finite relational structures--- can then
% be understood as the question of characterizing those languages $L$ that are in
% $\PT$ and are graph properties.

% Anderson and Dawar~\cite{AndersonD17} study this question through the lens of
% circuit complexity, by considering \emph{symmetric circuits}. These are circuits
% that necessarily accept graph properties, with this fact explicitly witnessed by
% automorphisms of the circuit. To be precise, we can think of a property of
% (directed) graphs on $n$ vertices as being given by a Boolean function $f:
% \{0,1\}^{n^2} \rightarrow \{0,1\}$ which takes as input the $n^2$ potential
% edges and outputs $1$ if the property is satisfied by the graph given on the
% input. The function $f$ is \emph{graph-symmetric} in the sense that any
% permutation of its inputs induced by a permutation of the $n$ vertices leaves
% $f$ unchanged (note this is not the same thing as requiring that $f$ is
% invariant under all permutations of its inputs). A symmetric circuit is a
% circuit computing a Boolean function of $n^2$ inputs where each permutation of
% $[n]$ can be extended to an automorphism of the circuit. Such a circuit
% necessarily computes an invariant function.

% The results of~\cite{AndersonD17} show a tight relationship between the
% decdability of a graph property by means of a polynomial-size family of
% symmetric circuits and its definability in extensions of fixed-point logic. In
% particular, they show that a property is definable in fixed-point logic with
% counting ($\FPC$) if, and only if, it is decidable by a polynomially-uniform
% family of symmetric circuits using Boolean and majority gates (or, equivalently,
% threshold gates). Without the symmetry restriction, polynomial-size families of
% circuits have the same computational power whether we restrict ourselves to
% Boolean gates only or allow richer bases such as threshold gates. One
% consequence of the results in~\cite{AD14} is that, with the requirement of
% symmetry, these different bases do yield different expressive power, suggesting
% some lack of robustness to the model. One of the results in the present paper
% shows that the model with majority circuits is robust in the sense that it is
% able to simulate, by means of symmetric circuits, any symmetric circuit composed
% of gates computing Boolean functions that are themselves symmetric (in the sense
% that they are invariant under \emph{all} permutations of their inputs. Thus, any
% property decidable by a polynomially-uniform family of such circuits is
% definable in $\FPC$.

% This motivates the study of symmetric circuit families that involve gates which
% themselves may compute non-symmetric functions. In particular, motivated by the
% study of rank logic~\cite{DGHL09,GP15}, we consider gates that are
% \emph{matrix-symmetric}. These are gates computing a Boolean function $f:
% \{0,1\}^{A \times B} \rightarrow \{0,1\}$ which is invariant under arbitrary
% permutations of $A$ and $B$. A typical example is the function computing the
% rank of an $A \times B$ $0$-$1$ matrix. We show that the support theorem of
% Anderson and Dawar can be shown to hold, even in the presence of
% matrix-symmetric gates. This allows us to show that definability in the logic
% fixed-point with rank ($\FPR$) is the same as decidability by
% polynomially-uniform families of symmetric circuits with rank gates.
% % Another extension of $\FPC$ that has received much attention in the
% % descriptive complexity literature is the class $\CPTC$ of \emph{choiceless
% % polynomial time with counting}. The work of~\cite{AD14} leaves open the
% % question of obtaining a circuit characterization of this class and it was
% % suggested there that this would require a relaxation of the notion of symmetry
% % used in defining symmetric circuits. In particular, while $\CPTC$ is based on
% % a machine model designed to preserve symmetry by forbidding choices, the
% % polynomial time and space restrictions apply int he context of the symmetries
% % fo a \emph{single input structure}. In this paper, we consider a relaxation of
% % the notion of symmetric circuit intended to capture exactly this intuition.
% % However, we are able to show that this does not yield a characterization of
% % $\CPTC$. In particular, we show that no family in this class of relaxed
% % symmetric circuits can distinguish odd from even Cai-F\"urer-Immerman graphs,
% % something we know is do-able in $\CPTC$~\cite{DRR08}.

\end{document}