### 2017-11-13 sections/symmetric-circuits.tex

\documentclass[../paper.tex]{subfiles}
%\usepackage{../mymacros}
\begin{document}
% Note: include circuits in background

%AD: re-wrote the introductory paragraphs to be a bit more substantive.

A Boolean circuit $C$ computing a function from $\{0,1\}^n$ to
$\{0,1\}$ is usually described as a directed acyclic graph, with each
gate $g$ that has $m$ incoming edges labelled with a Boolean function
$f_g : \{0,1\}^m \rightarrow \{0,1\}$ from a basis $\mathbb{B}$ and
exactly $n$ input gates of $C$ are labelled by variables.  If the
function $f_g$ can be arbitrary, then the circuit $C$ must impose an
order on the gates that provide the inputs to $g$.  When we say that
$C$ is a directed acyclic graph, without further ordering the nodes,
we implicitly assume that the functions in the basis $\mathbb{B}$ are
\emph{symmetric}.  That is to say $f_g$ is invariant under all
permutations of its $m$ inputs.  This unstated assumption is pervasive
in circuit complexity.  In particular, the standard Boolean basis of
AND, OR and NOT gates as well as bases with majority or threshold
gates only contain symmetric functions.

We are interested in Boolean circuits that compute queries on
structures, such as graphs.  In the case of (undirected, loop-free)
graphs on $n$ vertices, such a circuit might compute a function from
$\{0,1\}^{n \choose 2}$ to $\{0,1\}$.  In this case, the function
computed by the circuit is not necessarily invariant under all
permutations of the inputs, but it is invariant under all permutations
of the $n \choose 2$ inputs induced by permutations of $[n]$.
We are especially concerned with \emph{symmetric} circuits, that is
those where the permutations of $[n]$ extend to automorphisms of the
circuit.  Note that this use of the word ``symmetric'' is distinct to
that its use when applied to Boolean functions.  For such circuits, Anderson and Dawar~\cite{AndersonD17}
consider $\PT$-uniform circuits in the standard Boolean basis as
well one with majority gates.  It is a consequence of their results
that the latter are strictly more powerful than the former.  In
particular, they are shown to be equivalent
to the logic $\FPC$.  In contrast, we show, in this section, that adding any
further symmetric functions to the basis does not allow us to extend
the power of $\PT$-uniform symmetric circuits beyond that of $\FPC$.

As our aim is a circuit characterisation of $\FPR$, whose expressive
power is strictly greater than that of $\FPC$, we need to
consider gates corresponding to non-symmetric Boolean functions.  In
particular, we consider \emph{rank} gates whose inputs are a matrix
and whose result is invariant under row and column permutations.  To
lead up to this, we first develop a general framework of structured
Boolean functions.  These are functions whose inputs naturally encode
$\tau$-structures and where the output is invariant under the natural
symmetries of such structures.  The\emph{matrix-invariant} functions
are a natural partiruclar case.   We therefore define symmetric circuits in a
general form where gates can be labelled by \emph{isomorphism-invariant} structured
functions.

Ultimately, our aim is replicate the results Anderson and
Dawar~\cite{AndersonD17} in this more general setting.  A difficulty
we face is that the proof methods in that paper heavily rely on the
assumption that the Boolean functions computed by individual gates are
symmetric functions.  We address this difficulty in the next section.

% Boolean circuits have long acted as an alternative model from which to study
% algorithmic complexity. A circuit is evaluated for some input binary sequence by
% recursively evaluating each gate in the circuit, starting from the input gates.
% Suppose then we in the process of evaluating the circuit $C$ and are at the
% stage of evaluating the gate $g$. We take the input sequence formed from the
% evaluations of the gates input to $g$ and applying the element of the basis that
% labels $g$ to that sequence. It follows that in order to evaluate $g$ we need to
% suppose some order on the evaluations of the input gates. In many cases this
% order is simply induced by an assumed order on the whole of the circuit; an
% assumption common in most parts of complexity theory, where the encoding of
% structures for computation induces such orders, but problematic in finite model
% theory.

% Circuit models have also been studied in finite model theory, often as a means
% for characterising the expressive power of logics. In particular, research has
% focused on symmetric circuits defined over bases of symmetric Boolean functions
% (see \cite{}). The fact that the basis consist of only symmetric functions is a
% crucial requirement, ensuring that the evaluation of a gate does not depend on
% an external order. However, in this paper we will need to consider circuits with
% gates that compute rank, a function that is not symmetric.

% In this section we extend this circuit model so as to allow for the inclusion of
% non-symmetric functions in the basis. We allow each gate $g$ to index its input
% gate in accordance with the basis element labeling $g$. We introduce the notion
% of a \emph{structured function}, a Boolean function whose input is indexed in
% accordance with a many-sorted vocabulary $\tau$. The input vector to a
% structured function can then be thought of as encoding a $\tau$-structure. In
% particular, we consider structured functions that compute properties of
% $\tau$-structures, which we call \emph{isomorphism-invariant} structured
% functions. We then define what it means for a circuit over a basis of
% isomorphism-invariant structured functions to be symmetric. This generalises the
% notion of a symmetric circuit. We also show that there are limitations on the
% power of symmetric circuits over bases of symmetric functions, justifying the
% importance of this generalisation.

% We incorporate this labelling in the circuit model by including for each gate
% $g$ labelled by a structured function $$an indexing on its children that
% matches the index for the function that labels $g$. We allow these circuits to
% be defined over bases of structured functions with this more general symmetry
% condition

% In order then to include a gate computing rank in our symmetric circuit model
% we need to rec We are interested in circuit with gates that can compute rank
% and so, since rank is not a symmetric function, we develop a symmetric circuit
% model capable of incorporating non-symmetric functions in its basis. Moreover,
% we note that

% In particular, we are interested in Boolean functions that compute rank, i.e.
% functions that take in a binary matrix and compute the rank of that matrix
% over some prime field and compare the result with some threshold. But clearly
% this rank function is not symmetric. Furthermore, its evaluation depends on
% how the input set is labelled to form a matrix.

% In this section we develop the

% allowing us to develop natural symmetry conditions on the functions in terms
% of symmetry conditions on the structure. This indexing allows us to generalise
% the notion of a symmetric function by relativising the symmetry condition in
% accord with the structure on the input string. Incorporating this indexing
% into the circuit and

% ore recently, they have played a role in finite model theory, with many
% authors (e.g. ) studying circuit models that take in some representation of a
% finite algebraic structure and which obey certain natural symmetry conditions.
% However, there are some subtleties in the definition and evaluation of
% circuits which must be considered carefully and have important implications
% for the notion of a symmetric circuit studied in this paper. Let $C$ be a
% Boolean circuit $g$ be an internal gate labelled by some function $f_g:
% \{0,1\}^{A} \rightarrow \{0,1\}$, for some index set $A$ with $\vert A \vert =
% \vert H_g \vert$. Formally, evaluating $g$ requires a (usually bijective)
% indexing function $f_i: A \mapsto H_g$, which organises the inputs of $g$ into
% an appropriate sequence. The evaluation of $g$ is then given by $C[\vec{x}](g)
% = f_g (a \mapsto C[\vec{x}](f_i(a)))$. As such the evaluation of a gate in a
% circuit may depend on how the inputs to the gate are indexed.

% In most contexts in complexity theory, circuits may be taken to be implicitly
% ordered objects and so this indexing function can be defined as taking the
% $i$th element of $A$ to the $i$th element of $H_g$. However, in general this
% renders the evaluation of a circuit for a given input a property both of the
% circuit and an arbitrary order. In order to address this problem this problem
% many authors only define circuits over Boolean bases of symmetric functions
% (e.g. the standard basis or the majority basis). In this case the evaluation
% of a gate is independent of the choice of indexing function.

% However, in this paper we require gates labelled by Boolean rank functions,
% i.e. a function that takes in a Boolean sequence indexed by `row' and `column'
% sets, outputs 1 iff that matrix, when interpreted as having entries in
% $\mathbb{F}_p$, has rank at least $r$. These rank functions clearly depend on
% how the input sequence is labelled, and so are not symmetric. However, it is
% worth noting that while such rank functions may be symmetric, they are
% symmetric in the weaker sense of being constant under permutations on the rows
% and columns of the input matrix.

% Motivated by the desire to study circuits with gates that compute rank, We
% develop a general framework for circuits that allows gates to be labelled by
% Boolean functions of the form $F: \{0,1\}^{X} \rightarrow \{0, 1\}$, where the
% labelling $X$ is the universe of a many-sorted structure over some vocabulary
% $\tau$. The Moreover, we generalise the notion of a circuit for structures
% (see Anderson and Dawar \cite{AndersonD17}), incorporate an appropriate
% labelling for the inputs of each gate and allowing for a Boolean basis
% including non-symmetric functions. Furthermore, the vocabulary $\tau$ will
% allow us to impose natural symmetry conditions on $F$, and so allowing us to
% develop a useful notion of a symmetric circuit.

% We also prove that no family symmetric functions defined over a basis of
% symmetric functions can improve on the expressive power of the circuit model
% of Anderson and Dawar \cite{AndersonD17}. Together with the main theorem of
% this paper this implies that these more general symmetric circuits are
% strictly more powerful

% consider the problem of explicitly labelling the input of the function both so
% that the semantics of the function can be reasonably defined and so as to
% ensure that appropriate symmetry conditions on the function can be enforced.

% labelled by functions that are not symmetric, we need to think explicitly
% about how the input is to be labelled and how that structure should be
% mirrored in the circuit definition. In particular, we should like to consider
% functions which invariant under some action and so choose labellings which
% appropriately reflect the required symmetry conditions of the function.

% In this section we first discuss how to structure the input to a Boolean
% function appropriately and how symmetries on that input structure induce
% useful symmetries on the function.

\subsection{Structured Functions and Symmetry}
Let $X$ be a finite set and $F: \{0,1\}^{X} \rightarrow \{0,1\}$. We call $X$
the \emph{index} of $F$, and denote it by $\ind(F)$. We are often interested in
classes of functions characterised by certain symmetry conditions. Perhaps the
simplest example is the class of symmetric functions, i.e.\ those functions
invariant under the action of the symmetric group on their index. However, in
many cases this symmetry condition may be too strong, or otherwise
inappropriate. For example, suppose $X$ consists of all two element subsets of
some set $V$. We may think of $X$ as containing all potential edges in a graph,
and the input sequence $ \vec{x} \in \{0,1\}^X$ as encoding a graph. In this
case we might require instead that $F$ be invariant under the action of $\sym_V$
on $\vec{x}$. This condition guarantees that $F$ determines a graph property,
and we call such a function \emph{graph-invariant}. Another example of interest,
especially in this paper, is the case where $X := A \times B$, where $A$ and $B$
are non-empty sets, and we think of $\vec{x} \in \{0,1\}^X$ as encoding an
(unordered) matrix. It is natural then to consider the case where $F$ is
invariant under the action of $\sym_A \times \sym_B$, and we say such a function
is \emph{matrix-invariant}. In this subsection we develop a general framework
for dealing with functions that take in structures.

%AD - rewrote the paragraph to reduce the use of subscripts and superscripts.
Let $\tau = (R, S, \nu)$ be a many-sorted signature.
% with $R = \{R_1 \ldots,
% R_t\}$, $S = \{s_1, \ldots , s_p\}$, and $\nu(R_i) = (a^i_1, \ldots ,
% a^i_{r_i)})$, where $r_i = \arty(R_i)$ and $a^i_q \in S$ for all $q \in
% [r_i]$.
Let $D := \biguplus_{s \in S} D_{s} = \{(s,d) : d \in D_s\}$, be a
disjoint union of non-empty sets, indexed by $S$, i.e.\ $D$ is
potentially the universe of a $\tau$-structure. We call $D$ a \emph{$\tau$-sorted set}.
For each $P \in R$, let $P^D$ denote the trivial relation over the
many-sorted universe $D$, i.e.\ $P^D := D_{a_1} \times \ldots \times
D_{a_{r}}$, where $\nu(P) = (a_1, \ldots , a_{r)})$.   Let the \emph{index defined by the pair $(\tau, D)$} be
the set $\biguplus_{P\in R} P^{D}$.  We use this set to index the input to a Boolean
function, as such we abuse notation and denote the index defined by
$(\tau, D)$ by $\ind(\tau, D)$. We call $\tau$ the vocabulary and $D$ the universe of $\ind(\tau, D)$.

For the remainder of this subsection we fix a vocabulary $\tau$ and a
many-sorted set $D=\biguplus_{s \in S} D_{s}$
and let $X := \ind (\tau, D)$. Let $G \leq \sym_{X}$, and let
$\sortsym(D) := \prod_{s \in S} \sym_{D_{s}}$.  We call the elements of $\sortsym(D)$ \emph{sorted permutations}. Note that
$\sortsym(D)$ can be identified with a subgroup of $\sym_X$. There is an
obvious action of $G$ on $X$.  We can extend this to an action on the set of functions
of the form $f: X \rightarrow X$, for any set $H$, defined by $\sigma \cdot f(x)
= f(\sigma \cdot x)$ for all $x \in X$, $\sigma \in G$.

% If $H$ is surjective, we say that \emph{$f$ indexes $H$ by $X$} or \emph{$f$
% indexes $H$ by $\mathcal{D}$}.
We say that two functions $f, g : X \rightarrow H$ are \emph{$G$-equivalent} if
there exists $\sigma \in G$ such that $f = \sigma \cdot g$. We say $f$ and $g$
are \emph{isomorphism-equivalent} if they are $\sortsym(D)$-equivalent.
We are interested in the case where $H = \{0,1\}$, i.e.\ $f : X
\rightarrow \{0,1\}$. In this case $f$ defines for each relation $R^D_i$ a subset of that relation, and thus determines a $\tau$-structure over the universe $D$. We note that two functions $f, g: X \rightarrow
\{0,1\}$ are isomorphism-equivalent if, and only if, the two $\tau$-structures
they determine are isomorphic.

We call a function of the form $F:\{0,1\}^{X} \rightarrow \{0,1\}$ a
\emph{$(\tau, D)$-structured function}, or just a \emph{structured function}
when there is no need to emphasise the particular vocabulary.  We note that
$\ind(F) = \ind(\tau, D)$, justifying our choice of notation.  We let
\emph{the vocabulary of $F$} and \emph{the universe of $F$} denote
$\tau$ and $D$ respectively.

% It is easy to see that if $\mathcal{C}_A$ is the category with objects given
% by $\tau$-structures over $A$ and morphisms by the action of $G$, and
% $\mathcal{C}_X$ is the category of subsets of $X^{\tau}_A$ with morphisms
% similarly given by the action of $G$, then these two categories are
% equivalent. More informally, the subsets of $X^{\tau}_A$ encode the
% $\tau$-structures over $A$. This relationship is bijective and the action of
% $G$ factors through this bijection.

% \begin{remark}
%   Is the categorical language unnecessary? It seemed the quickest way of
%   saying what I wanted to say formally.
% \end{remark}

% \begin{remark}
%   In the above section (and just below) we talk about the action of
%   $\sym_{A_{s_1}} \times \ldots \times \sym_{A_{s_p}}$ on elements of
%   $X^\tau_A$ and structured functions. In the proof of the support theorem,
%   when defining a definable matrix for converting circuits into formulas, and
%   later on in this section, it is useful to speak more generally and instead
%   look at bijections from $A_1, \ldots A_{s_p}$ to $B_1, \ldots, B_{s_p}$,
%   thus allowing us to map between $X^\tau_A$ and $X^\tau_B$ and so between the
%   associated structured functions. I've added this in later on in the section,
%   but I still need to incorporate it into the original definition (it's a more
%   general notion in a sense, and so I think it should be incorporated). I'll
%   wait for feedback before doing this as I'm not sure this formulation will
%   survive.
% \end{remark}

Structured functions give us a general framework for indexing the input sequence to a Boolean function and associating with that function an appropriate symmetry condition. Let $F$ be a
$(\tau, D)$-structured function. We say $F$ is \emph{$G$-invariant} if it is
constant on $G$-equivalent input vectors, i.e.\ for all $G$-equivalent $\vec{x},
\vec{y} \in \{0,1\}^{X}$, $F(\vec{x}) = F(\vec{y})$. We say $F$ is
\emph{isomorphism-invariant} if it is $\sortsym (D)$-invariant. Note that
a structured function with vocabulary $\tau$ is isomorphism-invariant if, and
only if, it decides a property of $\tau$-structures.

At the beginning of this subsection we defined three Boolean functions with
differently indexed inputs and corresponding symmetry conditions. We now look at
how these functions might be thought of as structured functions, and how the
symmetry conditions we discussed are naturally reproduced. Suppose $\tau$ is
single-sorted and consists of a single unary relation. Then any structured
function $F$ with vocabulary $\tau$ is isomorphism-invariant if, and only if, it
is symmetric. If instead the vocabulary of $F$ is the graph vocabulary, then $F$
is isomorphism-invariant if, and only if, $F$ decides a property of directed
graphs. Lastly, if the vocabulary of $F$ has just two sorts $(s_1, s_2)$ and one
binary relation with type $(s_1, s_2)$, then $F$ is isomorphism-invariant if,
and only if, it is matrix-invariant.

In each case a natural choice of the structure indexes the input to the
structured function appropriately and produces the appropriate symmetry
condition. It is worth mentioning that the graph-invariant case is not strictly
reproduced above, as the structured functions are rather defined over directed
graphs.

From the above observations, we may think of symmetric functions as
isomorphism-invariant structured functions over the single-sorted vocabulary with
a single unary relation. We can similarly think of matrix-invariant functions as
isomorphism-invariant structured functions over the appropriate vocabulary. More
formally, if the vocabulary of $F$ consists of $p$ sort symbols $s_1, \ldots ,
s_p$ and only one relation symbol with type $(s_1, \ldots, s_p)$ and $F$ is
isomorphism-invariant, then we say $F$ is \emph{sort-invariant}. If $p = 2$ and
$F$ is sort-invariant then we say that $F$ is \emph{matrix-invariant}.

In keeping with convention for Boolean functions, and for introducing simplicity
without a loss of generality, we assume that the universe of any structured
function is always a disjoint union of initial segments of the natural
numbers, i.e.\ if $F$ is structured function with universe $D_{s_1} \uplus \ldots
\uplus D_{s_p}$, then for all $s \in \{s_1 , \ldots , s_p\}$ there exists $d_s \in \mathbb{N}$ such
that $D_{s} = [d_s]$. We also assume that every $(\tau, D)$-structured function has the property that every $a \in D$ appears in at least one element of $\ind(\tau, D)$.

% Let $F$ be a $(\tau, A)$-structured function with $\tau = (\{R\}, \{1\},
% \nu)$, where $R$ is a binary relation. Let $G \leq \sym_A$ such that for any
% $v, w \in A$ We say $F$ is \emph{graph-symmetric} if $F$ is $(\tau, A)$simple
% and matrix-symmetric.

% If $\tau$ is single-sorted we call $F$ \emph{simple} and if it consists of a
% single unary relation, we call $F$ \emph{unary-relational}.

% We briefly introduce notions of equivalence useful for comparing functions.
% \begin{definition}
%   Let $\tau = (R, S, \nu)$ be a many-sorted signature and let $A = A_1 \times
%   \ldots A_s$ and $B = B_1 \times \ldots \times B_s$ be a product of non-empty
%   sets. A \emph{sorted bijection} between $A$ and $B$ is a function bijections
%   $f: A \rightarrow B$ such that $f (A_i) = B_i$ for all $i \in S$. There is
%   an obvious action of $f$ that maps $X^\tau_A$ to $X^\tau_B$ and
%   $\{0,1\}^{X^\tau_A}$ to $\{0,1\}^{X^\tau_B}$.

%   Let $L: X^\tau_A \rightarrow H$ and $L': X^\tau_B \rightarrow H$ be two
%   functions for some finite set $H$. We say that $L \sim L'$, or \emph{$L$ is
%   equivalent to $L'$}, if there is a sorted bijection $f: A \rightarrow B$
%   such that for all $x \in X^\tau_A$, $L(x) = L'(f (x))$.

%   Let $F$ be a $(\tau, A)$-structured function and $G$ be a $(\tau,
%   B)$-structured function. Then we say that $F \sim G$, or \emph{$F$ is
%   equivalent to $G$}, if there is a sorted bijection $f: A rightarrow B$ such
%   that for all $\vec{i} \in \{0,1\}^{X^\tau_A}$, $F (\vec{i}) =
%   G(f(\vec{i}))$.
% \end{definition}

% \begin{remark}
%   There is an obvious connection between $\tau$-symmetric functions and
%   generalised quantifiers (or closed classes of structures). Should I include
%   details about this connection? I also feel a lot of dirtiness might be
%   avoidable if we instead recast everything in terms of generalised
%   quantifiers. For one, generalised quantifiers give a natural way of defining
%   a Boolean function for each input universe.
% \end{remark}

\subsection{Symmetric Circuits}
% We define a circuit over a basis of structured functions that are not
% necessarily symmetric. As such, the circuit must include for each gate $g$ an
% index on its children corresponding to the index of the function labelling
% $g$.
In this section we extend the circuit definitions of Anderson and Dawar
\cite{AndersonD17}, introducing a circuit for structures that incorporates an
indexing for each gate in accordance with the Boolean function labelling that
gate. This allows us to define (and evaluate) circuits with gates labelled by
non-symmetric functions.

% Having developed the notion of a function that accepts input structured in
% accordance with some vocabulary, we now develop a circuit model which
% incorporates this structure on the inputs of a gate. We use this general
% framework to define the notion of a matrix-symmetric circuit, a circuit with
% gates labelled by matrix-symmetric Boolean functions, and finally we develop
% the notion of a matrix-symmetric circuit with rank.

% Importantly, many natural functions of interest are matrix symmetric. For
% example, the function that computes the rank of the matrix over
% $\mathbb{F}_2$. or a thresholded rank function, for example the rank of the
% matrix over $\mathbb{F}_p$ being larger then $r$, for some particular $(p, r)
% \in \mathbb{N}$.

We now define a \emph{$(\mathbb{B}, \tau)$-circuit} . This is a circuit that
computes a function on structures in the relational vocabulary $\tau$ and where
the gates are labelled by elements of the basis $\mathbb{B}$. Each element of
the basis is a $(\tau,D)$-structured function for some $\tau$ and $D$. Note
that, in the interests of keeping our definitions from getting excessively
complicated, we do not consider circuits computing functions over multi-sorted
vocabularies.
\begin{definition}[Circuits on Structures]
  Let $\mathbb{B}$ be a basis and $\tau$ be a relational vocabulary, we define a
  \emph{$(\mathbb{B}, \tau)$-circuit} $C_n$ computing a $q$-ary query $Q$ as a
  structure $\langle G, \Omega, \Sigma, \Lambda, L \rangle$.
  \begin{itemize}
    \setlength\itemsep{0mm}
  \item $G$ is called the set of gates of $C_n$ and $\vert C_n \vert := \vert G
    \vert$.
  \item $\Omega$ is an injective function from $[n]^q$ to $G$. The gates in the
    image of $\Omega$ are called the output gates. When $q = 0$, $\Omega$ is a
    constant function mapping to a single output gate.
  \item $\Sigma$ is a function from $G$ to $\mathbb{B} \uplus \tau \uplus
    \{0,1\} $ such that $\vert \Sigma^{-1} (0) \vert \leq 1$ and $\vert
    \Sigma^{-1} (1) \vert \leq 1$. Those gates mapped to $\tau \uplus \{0,1\}$
    are called input gates, with those mapped to $\tau$ called relational gates
    and those mapped $\{0,1\}$ called constant gates. Those gates mapped to
    $\mathbb{B}$ are called internal gates.
  \item $\Lambda$ is a sequence of injective functions $(\Lambda_{R_i})_{R_i \in
      R}$ such that $\Lambda_{R_i}$ maps each relational gate $g$ with $\Sigma
    (g) = R_i$ to the tuple $\Lambda_{R_i} (g) \in [n]^{r}$. When no ambiguity
    arises we write $\Lambda (g)$ for $\Lambda_{R_i} (g)$.
  \item $L$ associates with each internal gate $g$ a function $L(g):
    \ind(\Sigma(g)) \rightarrow G$ such that if we define a relation $W
    \subseteq G^{2}$ by $W(h_1,h_2)$ iff $h_2$ is an internal gate and $h_1$ is
    in the image of $L(h_2)$, then $(G, W)$ is a directed acyclic graph.
  \end{itemize}
\end{definition}

% \begin{definition}[Circuits on Structures]
%   Let $\mathbb{B}$ be a basis of structured functions and $\tau := (R, \{s_1,
%   \ldots, s_p\}, \nu)$ be a many-sorted vocabulary and let $\vec{n} :=
%   (n_{s_1}, \ldots , n_{s_p}) \in \mathbb{N}^{S}$, we define a
%   \emph{$(\mathbb{B}, \tau)$-circuit} $C_{\vec{n}}$ computing a $q$-ary query
%   $Q$ of type $(s^Q_{1}, \ldots , s^Q_{q})$ is a structure $\langle G, W,
%   \Omega, \Sigma, \Lambda, L\rangle$.
%   \begin{itemize}
%     \setlength\itemsep{0mm}
%   \item $G$ is called the set of gates of $C_{\vec{n}}$ and $\vert C_{\vec{n}}
%     \vert := \vert G \vert$.
%   \item $W \subseteq G \times G$, where $W$ is called the wires of the
%     circuit. $(G,W)$ must be a directed acyclic graph. For $g \in G$ we $H_g
%     := \{ h \in G: W(h,g)\}$ be the set of children of $g$.
%   \item $\Omega$ is an injective function from $[n_{i_1}] \times \ldots \times
%     [n_{i_q}]$ to $G$. The gates in the image of $\Omega$ are called the
%     output gates. When $q = 0$, $\Omega$ is a constant function mapping to a
%     single output gate.
%   \item $\Sigma$ is a function from $G$ to $\mathbb{B} \uplus \tau \uplus
%     \{0,1\} $ which maps input gates to $\tau \uplus \{0,1\}$ and where $\vert
%     \Sigma^{-1} (0) \vert \leq 1$ $\vert \Sigma^{-1} (1) \vert \leq 1$ and the
%     internal gates get mapped into $\mathbb{B}$. Gates mapped to $R$ are
%     called relational gates and gates mapped to 1 or 0 are called constant
%     gates.
%   \item $\Lambda$ is a sequence of injective functions $(\Lambda_{R'})_{R' \in
%     R}$ where for each $R' \in R$ with arity $r$ and type $(s^{R'}_{1},
%     \ldots, s^{R'}_{r})$, $\Lambda_{R'}$ maps each relational gate $g$ with
%     $R' = \Sigma (g)$ to the tuple $\Lambda_{R'} (g) \in [n_{s^{R'}_1}] \times
%     \ldots \times [n_{S^{R'}_r}]$. When no ambiguity arises we write $\Lambda
%     (g)$ for $\Lambda_{R'} (g)$.
%   \item $L$ maps to each internal gate a labelling on its children. Let $g$ be
%     an internal gate. We have that $L(g)$ is a surjection from $\ind (\Sigma
%     (g))$ to $H_g$. We call $L(g)$ the \emph{child-labelling} for $g$.
%   \end{itemize}
% \end{definition}

Let $C_n = \langle G, \Omega, \Sigma, \Lambda, L \rangle$ be a $(\mathbb{B},
\tau)$-circuit. The $n$ here denotes that $C_n$ accepts structures of size $n$.
For a gate $g \in C_n$, we define the \emph{index} of $g$, denoted by $\ind(g)$,
as $\ind(\Sigma(g))$ and we let $H_g$ denote the set of all gates in the image
of $L(g)$. We also refer to the vocabulary and universe of $g$ to abbreviate
referring to the vocabulary and universe of $\Sigma(g)$.

% It is worth noting that in general computing the syntactic equivalence
% relation in polynomial time may not be possible. In fact, given two bipartite
% graphs it is possible to construct (in time polynomial in the size of the
% graphs) a circuit with gates labelled by matrix-symmetric functions such that
% two specified gates are syntacticly equivalent if,and only if, the two
% bipartite graphs are isomorphic. Since there is a polynomial time reduction
% from the graph isomorphism problem to the graph isomorphism problem on
% bipartite graphs, it follows that there is a polynomial time reduction from
% the graph isomorphism to the problem of computing syntactic equivalence. We
% formalise this result in the following proposition, and give the explicit
% construction referenced.

% \begin{prop}
%   There is a a polynomial time reduction from Graph Isomorphism to the problem
%   of deciding if two specified gates in a matrix-circuit are syntacticly
%   equivalent.
% \end{prop}
% \begin{proof}
% \end{proof}

% In contrast, in the restricted case where the gates of the circuit are
% labelled only by symmetric Boolean functions then the syntactic equivalence
% relation can be computed in polynomial time.

% \begin{prop}
%   Let $C_n$ be a circuit with symmetric gates. Then the problem of syntactic
%   equivalence relation on the gates of $C_n$ can be computed in polynomial
%   time.
% \end{prop}
% \begin{proof}
% \end{proof}

Let $\tau$ be a relational vocabulary, % = (R_1, \ldots , R_t)$
$\mathcal{A}$ a $\tau$-structure over an $n$-element universe $U$ and $\gamma
\in [n]^{\underline{U}}$.
% AD - need to define the notation [n]^{\underline{U}} in the background section
% -- DONE

Let $\gamma \mathcal{A}$ be the structure on universe $[n]$ formed by mapping
$U$ in accordance with $\gamma$. The evaluation of a $(\mathbb{B},
\tau)$-circuit $C_n$ computing a $q$-ary query $Q$ proceeds by recursively
evaluating the gates in the circuit. The evaluation of the gate $g$ for the
bijection $\gamma$ and input $\mathcal{A}$ is denoted by $C[\gamma
\mathcal{A}](g)$, and is given as follows:
\begin{enumerate}
  \setlength\itemsep{0mm}
\item If $g$ is a constant gate then it evaluates to the bit given by
  $\Sigma(g)$.
\item If $g$ is a relational gate then $g$ evaluates to true iff $\gamma
  \mathcal{A} \models \Sigma(g)(\Lambda (g))$.
\item If $g$ is an internal gate such that $\Sigma (g)$ let $L^{\gamma}_g:
  \ind(g) \rightarrow \{0,1\}$ be defined by $L^{\gamma}_g(x) = C[\gamma
  \mathcal{A}](L(g)(x))$, for all $x \in \ind(g)$. Then $g$ evaluates to true
  iff $\Sigma(g) (L^{\gamma}_g) = 1$.
\end{enumerate}
We say that $C_n$ defines the $q$-ary query $Q \subseteq U^q$ under $\gamma$
where $\vec{a} \in Q$ iff $C_n[\gamma \mathcal{A}](\Omega (\gamma \vec{a})) =
1$.

% The circuit $C_n$ The following definition is an important circuit property
% introduced by Anderson and Dawar \cite{AndersonD17}. We introduce it for the
% sake of comparison.

The following circuit condition is from Anderson and Dawar \cite{AndersonD17}.

\begin{definition}[Invariant Circuit]
  Let $C_n$ be a $(\mathbb{B}, \tau)$-circuit, computing some $q$-ary query. We
  say $C_n$ is \emph{invariant} if for every $\tau$-structure $\mathcal{A}$ of
  size $n$, $\vec{a} \in U^{q}$, and $\gamma_1, \gamma_2: [n]^{\underline{U}}$
  we have that $C[\gamma_1 \mathcal{A}](\Omega (\gamma_1 \vec{a})) = C[\gamma_2
  \mathcal{A}](\Omega (\gamma_2 \vec{a}))$.
\end{definition}

In other words $C_n$ is invariant if the evaluation of the circuit on any
$n$-element structure $\mathcal{A}$ does not depend on the particular map
$\gamma$ that takes the universe of the structure to $[n]$. In general, the
evaluation of a circuit depends on the presentation of the structure (i.e.\ the
bijection $\gamma$). If a family of $(\mathcal{B}, \tau)$-circuits $\mathcal{C}$
is invariant it follows that the query computed is a $q$-query on
$\tau$-structures. Thus if $q = 0$ then $\mathcal{C}$ computes a property of
$\tau$-structures. The following lemma allows us to recast this notion in terms
of the language developed in this paper.

\begin{lem}
  Let $C_n$ be a $(\mathbb{B}, \tau)$-circuit. The function computed by $C_n$ is
  $\tau$-invariant iff $C_n$ is an invariant circuit.
\end{lem}
% AD - does this require a few lines of proof? To be added

% \begin{definition}[Automorphism]
%   let $C = \langle G, \Omega, \Sigma, \Lambda, L\rangle$ be a
%   $(\mathbb{B},\tau)$-circuit computing a $q$-ary query on structures of size
%   $n$, and where $\mathbb{B}$ is a basis of isomorphism-invariant structured
%   functions. Let $\sigma \in \sym_n$ and $\pi: G \rightarrow G$ be a bijection
%   such that
%   \begin{itemize}
%     \setlength\itemsep{0mm}
%   \item for all output tuples $x \in [n]^q$, $\pi \Omega (x) = \Omega (\sigma
%     x)$,
%   \item for all gates $g \in G$, let $\Sigma (g) = \Sigma (\pi g)$,
%   \item for each relational gate $g \in G$, $\sigma \Lambda (g) = \Lambda (\pi
%     g)$, and
%   \item For each pair of gates $g, h \in G$ $W(h,g)$ iff $W(\pi h, \pi g)$ and
%     for each internal gate $g$ we have that $L(\pi g)$ and $ \pi \cdot L(g)$
%     are isomorphism-equivalent.
%   \end{itemize}
%   We call $\pi$ an \emph{automorphism} of $C$, and we say that $\sigma$
%   \emph{extends to an automorphism} $\pi$. The group of automorphisms of $C$
%   is called $\aut (C)$.
% \end{definition}

\begin{definition}[Automorphism]
  let $C = \langle G, \Omega, \Sigma, \Lambda, L\rangle$ be a
  $(\mathbb{B},\tau)$-circuit computing a $q$-ary query on structures of size
  $n$, and where $\mathbb{B}$ is a basis of isomorphism-invariant structured
  functions. Let $\sigma \in \sym_n$ and $\pi: G \rightarrow G$ be a bijection
  such that
  \begin{itemize}
    \setlength\itemsep{0mm}
  \item for all output tuples $x \in [n]^q$, $\pi \Omega (x) = \Omega (\sigma
    x)$,
  \item for all gates $g \in G$, let $\Sigma (g) = \Sigma (\pi g)$,
  \item for each relational gate $g \in G$, $\sigma \Lambda (g) = \Lambda (\pi
    g)$, and
  \item For each pair of gates $g, h \in G$ $W(h,g)$ if and only if $W(\pi h,
    \pi g)$ and for each internal gate $g$ we have that $L(\pi g)$ and $ \pi
    \cdot L(g)$ are isomorphism-equivalent.
  \end{itemize}
  We call $\pi$ an \emph{automorphism} of $C$, and we say that $\sigma$
  \emph{extends to an automorphism} $\pi$. The group of automorphisms of $C$ is
  called $\aut (C)$.
\end{definition}

\begin{definition}[Symmetry]
  A circuit $C$ on structures of size $n$ is called \emph{symmetric} if every
  $\sigma \in \sym_n$ extends to an automorphism on $C$.
\end{definition}

It follows that for any symmetric circuit $C_n$ there is a homomorphism $h$ that
maps $\sym_n$ to $\aut(C)$, taking any $\sigma \in \sym_n$ to an automorphism
that extends it. Such a homomorphism is injective so long as some element of
$[n]$ appears in a label of some input gate of $C$ (as then all elements appear
by symmetry)~\cite{AndersonD17} and therefore no two distinct elements of
$\sym_n$ agree on all input gates. In this paper we always assume that this is
the case, as otherwise all inputs are constant, and so the circuit just computes
a constant function. In order to develop a great many technical tools needed in
this paper we should like to be able to, without a loss of generality, assume
that that this homomorphism is in fact an isomorphism as well. In order to do so
we make use of the rigidity condition.

We define rigidity in terms of an equivalence relation on the gates of a
circuit. This recursively defined equivalence relation holds between two gates
if the circuit underneath these two gates are in some sense the same. Before we
define this relation we need to introduce some notation.

% When analysing circuits we are often interested in identifying families of
% gates that We are interested not just in the case where the labelling function
% $L(g)$ is an injection but when there are no two gates in its image that are
% in some broader sense the same. Of course we cannot efficiently decide if two
% gates in a circuit compute the same function (unless SAT is in $\P$). We may,
% however, consider a weaker relation on gates -- a syntactic condition which
% encodes the idea that the circuits underneath two gates are in some sense the
% same. We define this notion formally below. Before we do we define what it
% means for two functions to be isomorphism-equivalent up to an equivalence
% relation.

% We will often be interested in the case where two functions are
% isomorphism-equivalent up to syntactic equivalence. We define this notion (for
% an arbitrary equivalence relation) below.

\begin{definition}
  Let $\tau$ be a many-sorted vocabulary and $D$ be a $\tau$-sorted set. Let X
  be a set and $\sim$ be an equivalence relation defined on $X$. Let $Y, Z
  \subset X$. We say two functions $f : \ind(\tau,D) \rightarrow X$ and $g :
  \ind(\tau, D) \rightarrow Y$ are \emph{isomorphism-equivalent up to $\sim$} if
  $Y/{\sim} = X/{\sim}$ and $f/{\sim}$ is isomorphism-equivalent to $g/{\sim}$.
\end{definition}

\begin{definition}
  Let $C_n$ be a $(\mathbb{B}, \tau)$-circuit such that $C_n = \langle G, W,
  \Omega, \Sigma, \Lambda, L\rangle$. We recursively define the equivalence
  relation \emph{syntactic equivalence}, which we denote using the symbol
  `$\equiv$', on $G$ as follows. We say $g \equiv h$ if (i) $\Sigma(g) =
  \Sigma(h)$, (ii) if $g$ and $h$ are constant gates then they are equal, (iii)
  if $g$ and $h$ are relational gates then $\Lambda(g) = \Lambda(h)$, (iv) if
  $g$ and $h$ are internal gates then the functions $L(g)$ and $L(h)$ are
  isomorphism-equivalent up to syntactic equivalence, and (v) if $g$ and $h$ are
  output gates then $\Omega^{-1}(g) = \Omega^{-1}(h)$.
\end{definition}

We have already required that for each gate $g$ in a circuit must have an
injective labelling function. The following condition ensures that $L(g)/{\sim}$
is also an injection.

\begin{definition}
  Let $C_n$ be a circuit and $g$ be a gate in $C_n$. We say $g$ has \emph{unique
    labels} if $\vert H_g \vert = \vert H_g \setminus \equiv \vert $. We say
  that the circuit $C_n$ has \emph{unique labels} if every gate in $C_n$ has
  unique labels.
\end{definition}

We are also ready to define what it means for a circuit to be rigid.

\begin{definition}
  We say that a circuit $C_n$ is \emph{rigid} if every equivalence class of
  gates under the syntactic equivalence relation is a singleton.
\end{definition}

We prove in Proposition~\ref{} that there is a polynomial time algorithm that
takes in a symmetric circuit with unique labels and outputs an equivalent rigid
circuit. It is worth noting that the requirement that the circuit have unique
labels is very important (see Section \ref{} for a detailed discussion of this
point). Since we are interested in $P$-uniform families of circuits, this result
allows us to assume the rigidity of a circuit without a loss of generality.
Moreover, as we will show in Proposition \ref{}, if a circuit is rigid then
every $\sigma \in \sym_n$ has exactly one an extension to an automorphism of
$C_n$, and so the above defined homomorphism mapping $\sym_n$ to $\aut(C_n)$ is
an isomorphism. Since we may assume rigidity without a loss of generality this
property may also be assumed without a loss of generality. As such we abuse
notation and let $\sigma \in \sym_n$ also denote the induced automorphism of the
circuit.

% \begin{definition}[Rigidity]
%   Let $C_n$ be a $(\mathbb{B}, \tau)$-circuit, where $C_n = \langle G, W,
%   \Omega, \Sigma, \Lambda, L\rangle$. Say that $C_n$ is \emph{rigid} if there
%   are no distinct internal gates $g, g' \in G$ such that $\Sigma(g) = \Sigma
%   (g')$, $\Omega^{-1}(g) = \Omega^{-1}(g')$, $H_g = H_{g'}$, and $L(g')$ and
%   $L(g)$ are isomorphism-equivalent.
% \end{definition}

Let $\MB$ be the set of all structured functions that are symmetric or
matrix-invariant. Let $r, p \in \nats$, with $p$ prime, and let $D_1$ and $D_2$
be initial segments of the natural numbers. Then there is a matrix-invariant
structured function $F^{\rank}_{p,r}: \{0,1\}^{D_1 \times D_2}$, with universe
$D_1 \uplus D_2$, such that $F^{\rank}_{p,r}(M) = 1$ if, and only if, the matrix
$M \in \{0,1\}^{D_1 \times D_2}$ has rank at most $r$ when its entries are
understood as elements of the finite field $\mathbb{F}_p$. Let
$\mathbb{B}_\rank$ be the union of the set of all such functions with
$\mathbb{B}_{\maj}$.

\begin{definition}
  Let $\mathbb{B} \subseteq \MB$. We call a symmetric $(\mathbb{B},
  \tau)$-circuit with unique labels a \emph{symmetric matrix-circuit}.
\end{definition}

\begin{definition}
  Let $\mathbb{B} \subseteq \MB$. We call a symmetric $(\mathbb{B}_\rank,
  \tau)$-circuit with unique labels a \emph{symmetric rank-circuit}.
\end{definition}

% The natural restriction to consider on families of circuits is uniformity.

% \begin{definition}
%   Let $(C_n)_{n \in \mathbb{N}}$ be a family of Boolean circuits. We say that
%   $(C_n)_{n \in \mathbb{N}}$ is \emph{$P$-uniform} if the mapping $n \mapsto
%   C_n$ is computable in polynomial time.
% \end{definition}
We are now ready to state the main theorem of this paper.

\begin{thm}[Main Theorem]
  A graph property is decidable by a $P$-uniform family of symmetric
  rank-circuits if, and only if, it is definable by an $\FPR$ sentence.
\end{thm}

% Anderson and Dawar use the above proposition (or, rather, an analogous
% relation) in order to

% an automorphism of a circuit $C_n$ from the elements to $\aut(C_n)$. In the
% following sections this fact

% It is worth noting that in general computing the syntactic equivalence
% relation in polynomial time may not be possible. In fact, given two bipartite
% graphs it is possible to construct (in time polynomial in the size of the
% graphs) a circuit with gates labelled by matrix-symmetric functions such that
% two specified gates are syntacticly equivalent if,and only if, the two
% bipartite graphs are isomorphic. Since there is a polynomial time reduction
% from the graph isomorphism problem to the graph isomorphism problem on
% bipartite graphs, it follows that there is a polynomial time reduction from
% the graph isomorphism to the problem of computing syntactic equivalence. We
% formalise this result in the following proposition, and give the explicit
% construction referenced.

\subsection{Limitations of Symmetric Bases}
In this section we show that any family of symmetric circuits over an arbitrary
basis of symmetric functions can be transformed in polynomial time into a family of
symmetric circuits that decide the same language but are defined over the basis
$\mathbb{B}_\maj$.  In other words, it makes no difference to the
expressive power of such circuits whether we include \emph{all}
symmetric functions in the basis or just $\mathbb{B}_\maj$.
It follows that in order to construct $P$-uniform families of
symmetric circuits that define queires not in $\FPC$, we need to
consider bases with non-symmetric functions.

% Recall from Anderson and Dawar \cite{AndersonD17} we have that
% $\mathbb{B}_{\std} = \{ \neg , \wedge , \lor \}$ and $\mathbb{B}_\maj = \{
% \maj \} \cup \mathbb{B}_{\std}$.

Let $F : \{0,1\}^n\rightarrow \{0,1\}$ be a symmetric Boolean function. Recall
that the output of $F$ is entirely determined by the number of $1$s in its
input. Let $c_{F} \subseteq [n]$ be the set of all $m \leq n$ such that for
all $\vec{x} \in \{ 0,1 \}^n$ with $m$ $1$s we have $F (\vec{x}) = 1$. Clearly
any symmetric function $F$ is entirely determined by $c_{F}$. As such, $F$
may be encoded by a tuple $f \in \{0,1\}^{n}$, where $f (i) = 1 $ iff $i
\in c_{F}$. We assume this encoding below.

\begin{prop}
  \label{prop:fuctions-maj}
  There is a deterministic algorithm that outputs for each symmetric function
  $F: \{0,1\}^n \rightarrow \{0,1\}$ (encoded as a binary $n$-tuple as above) a symmetric circuit $C$ defined over
  the basis $\mathbb{B}_\maj$ that computes $F$. Moreover, this function is
  computable in time polynomial in $n$ and the circuit $C$ has depth at most
  $5$, width at most $2n+2$ and size at most $5n + 3$.
\end{prop}

\begin{proof}
  We have $c_{F}$ from the input. We now define $C$. We define the set of
  gates and wires of $C$ layer by layer as follows. The first layer consists
  of just the $n$ input gates labelled by the variables $x_1, \ldots , x_n$. The
  second layer consists of two `majority' gates for each $a \in c_{F}$, which
  we denote by $\maj_a$ and $\maj^{\neg}_a$. For each $a \in c_{F}$ there is
  one wire from each of $x_1, \ldots , x_n$ to $\maj_a$ and $\maj^{\neg}_a$. For
  all $a \geq \frac{n}{2}$, there are $2a - n$ wires from $0$ to $\maj_a$ and
  $2a - n + 2$ wires from $0$ to $\maj^{\neg}_a$. For all $a < \frac{n}{2}$
  there are $n - 2a$ wires from $1$ to $\maj_a$ and $n - 2a - 2$ wires from $1$
  to $maj^{\neg}_a$. The third layer consists of one `not' gate for each $a \in
  c_{F}$, which we denote by $\neg_a$. For each $a \in c_{F}$ there is a
  wire from $\maj^\neg_a$ to $\neg_a$. The fourth layer consists of one `and'
  gate for each $a \in c_{F}$, which we denote by $\countgate_a$. For each $a
  \in c_{F}$ there is a wire from each of $\maj_a$ and $\neg_a$ to
  $\countgate_a$. The fifth layer consists of just a single `or' gate,
  designated as the output gate, and for each $a \in c_{F}$ there is a wire
  from $\countgate_a$ to the output gate.

  We summarise the circuit up to the fourth layer as follows:
  \[
    \countgate_a = \begin{cases} \land ( \maj ( x_1, \ldots, x_n, \underbrace
      {0, \ldots, 0}_{2a - n} ), \neg ( \maj (x_1, \ldots , x_n, \underbrace{0,
        \ldots,
        0}_{2a - n + 2} ) )) &  a \geq \frac{n}{2} \\
      \land ( \maj ( x_1, \ldots, x_n, \underbrace {1, \ldots, 1}_{n - 2a} ),
      \neg ( \maj ( x_1, \ldots , x_n, \underbrace{1, \ldots, 1}_{n - 2a -2} )
      )) & a < \frac{n}{2}.
    \end{cases}
  \]
  We note that for an input vector $\vec{x}$, $count_a$ evaluates to 1 if, and
  only if, the number of 1's in $\vec{x}$ equals $a$. Thus we have that $C$
  evaluates to 1 if, and only if, there exists $a \in c_{F}$ such that the
  number of $1$'s in $\vec{x}$ equals $a$ if, and only if, $F (\vec{x}) = 1$.
  Moreover, it is easy to see that $C$ is symmetric and the algorithm defined
  runs in time polynomial in $n$.
%AD - perhaps include a brief argument for why C is symmetric?

  Furthermore, notice that the first layer contains $n$ gates and the second
  layer contains at most $2 \vert c_{F} \vert \leq 2n$ gates. The third and
  fourth layer each contain at most $n$ gates. As such $C$ has size at most $n
  + 2n + 2n + 1 + 2 = 5n +3$ (the additional 2 is for the constant gates). The
  width of $C$ is at most $2n + 2$, and the depth is at most $5$.
\end{proof}

\begin{thm}
  Let $\mathbb{B}$ be a basis of symmetric functions and let $(C_n)_{n \in
    \mathbb{N}}$ be a family of symmetric circuits defined over the basis $\mathbb{B}$. Then there exists a family
  of symmetric circuits $(C_n')_{n \in \mathbb{N}}$ defined over
  $\mathbb{B}_\maj$ that decides the same language. Moreover, the map $C_n
  \mapsto C_n'$ is polynomial-time computable and for each $n \in \nats$, $\vert
  C_n' \vert \leq (5 \cdot \vert C_n \vert + 3) \cdot \vert C_n \vert$.
\end{thm}

\begin{proof}
  From $C_n$ we construct $C_n'$ as follows. For each gate $g \in C_n$ 
  labelled by a member of $\mathbb{B}$ we have a symmetric circuit $C_g$ from
  Proposition~\ref{prop:function-maj} that computes the same function as $g$.
  Then let $C_n'$ be as $C_n$ but with each gate $g \in C_n$ replaced by $C_g$.
  It is easy to see that $C_n'$ is symmetric. We also have that each gate $g$
  must have at most $\vert C_n \vert$ inputs, and so the size of $C_g$ is
  bounded by $5 \vert C_n \vert + 3$. Thus the size of $C_n'$ is bounded by
  $(5f(n)+3) f(n)$. This algorithm clearly runs in polynomial-time.
\end{proof}

This result gives us that for any family of circuits over an arbitrary basis of
symmetric functions we can construct another symmetric circuit family
computing the same function over the majority basis without a blowup in size.
Moreover, we also have that if the first family was uniform the second family
will be as well. As such, we cannot extend the power of the $P$-uniform
symmetric circuits families studied by Anderson and Dawar \cite{AndersonD17} by
simply considering alternative bases of symmetric functions, thus motivating the
generalisation developed in the previous subsection to non-symmetric functions.
\end{document}