### 2018-02-10 sections/conclusion.tex

\documentclass[../paper.tex]{subfiles}
\begin{document}
In the quest for a logic that captures polynomial time

At the time of writing there are two candidate logics which 
Recent work on the problem of capturing $\PT$ has produced two `candidate'
logics for capturing polynomial-time.
In this paper we extend Anderson and Dawar's symmetric circuit characterisation
of $\FPC$, showing that $P$-uniform families of transparent symmetric circuits
with rank gates have the same expressive power as $\FPR$. Since $\FPR$ remains
one of the two c
the candidate logics for capturing $\PT$, along with choiceless polynomial time, 
\end{document}