### 2018-02-13 sections/translating-formulas-to-cicuits.tex
FEB 13

\documentclass[../paper.tex]{subfiles}
\begin{document}
In this section we first introduce a number of useful polynomial-time computable
functions which define transformations on circuits. In the second subsection we
use these functions to define a transformation that takes $\FPR$-formulas to
equivalent families of transparent symmetric rank-circuits.

\subsection{Algorithms for Circuits}
In order to simplify the presentation of the translation between $P$-uniform
families of circuits and formulas of $\FPR$, as well as the proofs of other
results (see Lemma~\ref{lem:transparent-unique}), we define a number of
often-used polynomial-time computable circuit transformations.

We first define the function $\op{merge}$, which maps a circuit and a set of
syntactically-equivalent gates to a circuit formed by combining those gates into
a single gate.

\begin{definition}
  Let $C := \langle G, \Omega, \Sigma , \Lambda, L \rangle$ be a circuit and $H
  := \{h_1, \ldots, h_p\}$ be a set of syntactically equivalent gates. We let
  $\op{merge}(C, H) := C' := \langle G', \Omega', \Sigma', \Lambda', L'
  \rangle$, where
  \begin{itemize}
    \setlength\itemsep{0mm}
  \item $G' = G \setminus \{h_2, \ldots, h_p\}$,
  \item $\Omega' = \Omega$,
  \item $\Sigma' = \restr{\Sigma}{G'}$,
  \item $\Lambda' = \Lambda$, and
  \item For all $g \in G'$ and $x \in Ind(g)$ if $L(g) (x) \in H$ then $L'(g)(x)
    = h_1$ and if $L(g)(x) \notin H$ then $L'(g)(x) = L(g)(x)$.
  \end{itemize}
\end{definition}

It can be shown that $\op{merge}$ can be computed in time polynomial in the size
of the input circuit. We note that the input and output circuits compute the
same function. Moreover, if the input circuit is symmetric (resp. has unique
labels, is transparent) then the output circuit will also be symmetric (resp.
have unique labels, transparent).

We now define the function $\op{merge-all}$ which maps a circuit and a set of
gates to a new circuit formed by applying $\op{merge}$ to the
syntactic-equivalence classes of the set of gates.

\begin{definition}
  \label{def:merge-all}
  The function $\op{merge-all}$ maps a circuit $C$ and a set of gates $H$ to a
  circuit $\op{merge-all}(C, H) := C'$, where $C'$ is defined as follows. Let
  $H_1, \ldots, H_p$ be the syntactic equivalence classes of $H$, and let $C'$
  be defined from $C$ by successively applying $\op{merge}$ to each of these
  equivalence classes.
\end{definition}

We can compute $\op{merge-all}$ using the following algorithm. Let $(C, H)$ be a
tuple in the domain of $\op{merge-all}$. We assume a reasonable encoding of the
circuit, and note that any such encoding induces a linear order on the
structure. Let $H_1, \ldots, H_p$ be the syntactic-equivalence classes of $H$
linearly ordered such that for all $i, j \in [p]$ if $i \leq j$ the height of
the gates in $H_i$ is less than or equal to the height of the gates in $H_j$.
Let $C'$ be the result of successively executing the algorithm for $\op{merge}$
on each such equivalence class in order.

It is worth noting that for an input tuple $(C, H)$, if follows from
Lemma~\cite{lem:unique-labels-syntactic-equiv} that if $H$ has unique labels
then the syntactic-equivalence relation on $H$ can be computed in time
polynomial in the size of $C$. Thus, if each gate in $H$ has unique labels, it
can be shown that $\op{merge-all}$ can be computed in time polynomial in the
size of the input circuit.

As in the case of $\op{merge}$, the function $\op{merge-all}$ preserves
symmetry, unique labels and the ensures that the input and output circuits
compute the same function. Let be a circuit $C$ a circuit and $H$ a set of gates
in $C$ with unique labels. Then if $g$ is a gate in both $C$ and $C' :=
\op{merge-all}(C, H)$, and all of the children of $g$ in $C$ are in $H$, then no
two distinct children of $g$ in $C'$ are syntactically-equivalent. However, we
have no guarantee that $g$ will have unique labels in $C'$ as we have no
guarantee that $L'(g)$ is injective.

The now define the function $\op{make-injective}$ that maps a circuit and a
symmetric gate to an equivalent circuit containing all the same gates as the
input circuit but such that the specified gate now has injective labels.

\begin{definition}
  \label{def:make-injective}
  The function $\op{make-injective}$ maps a circuit $C := \langle G, \Omega,
  \Sigma , \Lambda, L \rangle$ and a gate $h \in G$ to the circuit
  $\op{make-injective}(C, h) := C' := \langle G', \Omega', \Sigma', \Lambda', L'
  \rangle$, defined as follows. Let $c = \max \{\vert L(g)^{-1}(h)\vert : g \in
  W(h, \cdot)\}$. If $c = 1$ then $C' := C$. If $c > 1$ then $C'$ is defined
  such that
  \begin{itemizens}
  \item $G' = G \uplus \{h_2, \ldots, h_{c-1}\}$,
  \item $\Omega' = \Omega$,
  \item For each $g \in G$ let $\Sigma' (g) = \Sigma(g)$, otherwise let
    $\Sigma'(g)$ be a single-input $\AND$-gate.
  \item $\Lambda'= \Lambda$
  \item For each $g \in G'$ if $g \in G \setminus W(h, \cdot)$ then $L'(g) =
    L(g)$. If $g \in W(h, \cdot)$ let $\{x_1, \ldots, x_k\} := L(g)^{-1}(h)$ and
    for each $x \in \ind(g)$ if $L(g)(x) \neq h$ then $L'(g)(x) = L(g)(x)$ else
    if $x = x_i \in L(g)^{-1}(h)$ then let $L'(g)(x) = h_i$.
  \end{itemizens}
\end{definition}

It can be shown that $\op{make-injective}$ can be computed by an algorithm
running in time polynomial in the size of the input circuit. Moreover, the input
and output circuits of this operation compute the same function.

We now define the function $\op{make-injective-all}$ that maps a circuit to an
equivalent circuit such that every symmetric gate now has injective labels.

\begin{definition}
  The function $\op{make-injective-all}$ maps a circuit $C := \langle G, \Omega,
  \Sigma , \Lambda, L \rangle$ to a circuit $\op{make-injective}(C) := C' :=
  \langle G', \Omega', \Sigma', \Lambda', L' \rangle$ defined as follows.

  Let $F$ be a function that maps an ordered circuit $C*$ to the circuit
  $\op{make-injective}(C*, h)$, where $h$ is the topologically first gate such
  that there exists a gate $g$ in $C*$ where $h$ is a child of $g$ and $\vert
  L*(g)^{-1}(h) \vert > 1$. Let $(C^0, C^1, \ldots)$ be a sequence of circuits
  where $C^0 := C$ and for $i > 1$, $C^i := F(C^{i-1})$. It can be shown that
  this sequence converges after at most $\vert C \vert$ many steps. Let $C'$ be
  the limit of this sequence.
\end{definition}

Since $\op{make-injective}$ can be computed in polynomial time, and the number
of iterations required to converge to the limit in the above definition is
bounded by the size of the circuit, it can be shown that
$\op{make-injective-all}$ can be computed in time polynomial in the size of the
input circuit. Furthermore, we note that $\op{make-injective-all}$ preserves
symmetry, rigidity and unique labels. Moreover, the input and output circuits
compute the same function.

We now define the function $\op{insert}$ that maps two circuits to a circuit
formed by inserting the gates of the first circuit into the second circuit.

\begin{definition}
  The function $\op{insert}$ maps a $(\mathbb{B}_1, \tau_1)$-circuit $C^1 :=
  \langle G_1, \Omega_1, \Sigma_1 , \Lambda_1, L_1 \rangle$ and a
  $(\mathbb{B}_2, \tau_2)$-circuit $C^2 := \langle G_2, \Omega_2, \Sigma_2 ,
  \Lambda_2, L_2 \rangle$ to a $(\mathbb{B}_1 \cup \mathbb{B}_2, \tau_1 \uplus
  \tau_2)$-circuit $\op{insert}(C^1, C^2) := C := \langle G, \Omega, \Sigma ,
  \Lambda, L \rangle$, where
  \begin{itemize}
    \setlength\itemsep{0mm}
  \item $G = G_1 \uplus G_2$,
  \item $\Omega = \Omega_2$,
  \item $\Sigma = \Sigma_1 \uplus \Sigma_2$,
  \item $\Lambda = \Lambda_1 \uplus \Lambda_2$, and
  \item $L = L_1 \uplus L_2$.
  \end{itemize}
\end{definition}

It can be shown that $\op{insert}$ can be be computed in time polynomial in the
size of the input circuits. Moreover, if the input circuits are symmetric then
the output circuit is symmetric.

We now define the function $\op{replace}$ that maps a circuit $C$ and a pair of
gates $(g_1, g_2)$ in $C$ to a circuit $C'$ formed by replacing $g_1$ with $g_2$
in $C$.

\begin{definition}
  The function $\op{replace}$ maps a circuit $C = \langle G, \Omega, \Sigma,
  \Lambda, L\rangle$ and a pair of gates $(g_1, g_2)$ in $C$ to a circuit
  $\op{replace}(C, g_1, g_2) := C' := \langle G', \Omega', \Sigma', \Lambda', L'
  \rangle$, where
  \begin{itemize}
    \setlength\itemsep{0mm}
  \item $G' = G \setminus \{g_1\}$,
  \item $\Omega'$ is an function from the domain of $\Omega$ to $G'$ such that
    for each $\vec{a} \in \dom(\Omega)$ if $\Omega(\vec{a}) = g_1$ then
    $\Omega'(\vec{a}) = g_2$ and otherwise $\Omega'(\vec{a}) = \Omega(\vec{a})$.
  \item $\Sigma' = \restr{\Sigma}{G'}$
  \item $\Lambda' = \restr{\Lambda}{G'}$
  \item For each $g \in G'$ and each $x \in \ind(g)$ if $L(g)(x) = g_1$ then
    $L'(g)(x) = g_2$ and if not then $L'(g)(x) = L(g)(x)$.
  \end{itemize}
\end{definition}

It can be shown that the function $\op{replace}$ can be computed in time
polynomial in the size of the input circuit.

We now define the function $\op{tag}$ that maps a circuit, a relation symbol, a
gate in that circuit, and a tuple to a circuit computing the same function as
the input circuit but with an additional gadget, constructed using the gates
labelled by the given relation symbol and tuple, marking the designated gate.
The intuition is that this marking helps to distinguish a gate, and is used as a
step to ensure that a circuit has unique labels.

\begin{definition}
  The function $\op{tag}$ maps a $(\mathbb{B}, \tau)$-circuit $C := \langle G,
  \Omega, \Sigma, \Lambda, L \rangle$ of order $n$, a relation symbol $R \in
  \tau$, a gate $g \in G$, and a tuple $\vec{a} \in [n]^k$ for some $k$ to a
  circuit $\op{tag}(C, R, g, \vec{a}) := C' := \langle G', \Omega', \Sigma',
  \Lambda', L'\rangle$ of order $n$, such that $G' = G \uplus \{\land_g,
  \land_{g'}, \land_1, \ldots , \land_k\}$, $\Omega' = \Omega$, and $\Lambda' =
  \Lambda$. For each $g \in G$ $\Sigma'(g) = \Sigma(g)$. For each $g \in
  \{\land_0, \land_1, \ldots , \land_k \}$, $\Sigma(g)$ is an $\AND$-gate that
  takes in two inputs. $\Sigma(\land_g)$ is an $\AND$-gate that takes in a
  single input, $g$ is the lone input to $\land_g$, and $\land_0$ has as input
  $\land_g$ and $\land_1$. For each $i \in [k-1]$ the inputs of $\land_i$ are
  $\land_{i+1}$ and the gate $g_i \in G'$ such that $\Sigma(g_i) = R$ and
  $\Lambda^{-1}(g_i) = \vec{a}_i$, where $\vec{a}_i$ is the $k$-tuple
  $(\vec{a}(i), \ldots, \vec{a}(i))$. The inputs of $\land_k$ are the constant
  gate $1$ and the gate $g_k$ such that $\Sigma(g) = R$ and $\Lambda^{-1}(g_k) =
  \vec{a}_k$. For each gate $h \in G \setminus \{g\}$ and $x \in \ind(h)$, if
  $L(h)(x) = g$ then $L'(h)(x) = \land_0$, otherwise $L'(h)(x) = L(h)(x)$.
\end{definition}

It can be shown that $\op{tag}$ can be computed in time polynomial in the size
of its input. The input and output circuits compute the same function. Moreover,
if the input circuit has unique labels then the output circuit will have unique
labels. If the input relation symbol is obvious from the context we omit it when
invoking $\op{tag}$.

\begin{definition}
  The function $\op{tag-all}$ maps a $(\mathbb{B}, \tau)$-circuit $C := \langle
  G, \Omega, \Sigma, \Lambda, L \rangle$ of order $n$, a relation symbol $R \in
  \tau$, a set of gates $H \subseteq G$, and $T := (\vec{a}_h)_{h \in H}
  \subseteq [n]^{k}$ a sequence of tuples indexed by $H$ for some fixed $k$, to
  a circuit $\op{tag-all}(C, R, H, T) := C' := \langle G', \Omega', \Sigma',
  \Lambda', L'\rangle$ of order $n$ defined as follows.

  Suppose there is a linear order on $H$, and write $H = \{h_1, \ldots, h_t\}$.
  Let $C_0 := C$ and for all $i \in [t]$ let $C_i := \op{tag}(C_{i-1}, R, h_i,
  T(h_i))$. Then $C' = C_{t}$.
\end{definition}

It is easy to see that, since $\op{tag}$ can be computed in time polynomial in
the combined size of its inputs, $\op{tag-all}$ can also be computed in time
polynomial in the combined size of its inputs.

We now define the function $\op{compose}$ that maps a circuit and a sequence of
circuits and to a circuit formed by replacing the input gates of the designated
circuit with the output gates of the circuits in the sequence.

\begin{definition}
  The function $\op{compose}$ maps a $(\mathbb{B}, \tau)$-circuit $C := \langle
  G, \Omega, \Sigma, \Lambda, L \rangle$, a sequence of relation sybmols $\tau_R
  := (R_1, \ldots, R_t)$ such that $\tau_R \subseteq \tau$, a sequence of
  circuits $\mathcal{C} := ( C^1, \ldots, C^t)$, where for each $i \in [t]$,
  $C^i = \langle G_i, \Omega_i, \Sigma_i, \Lambda_i, L_i \rangle$ is a
  $(\mathbb{B}_i, \tau_i)$-circuit, and $\tau_i \not\subset \tau_R$ and $C^i$
  computes a query of arity $\arty(R_i)$ to a circuit $\op{compose}(C,
  \mathcal{C}, \tau_R) := C' := \langle G', \Omega', \Sigma', \Lambda', L'
  \rangle$. Note that all circuits take in structures of size $n$. The circuit
  $C'$ is defined as follows.

  Let ${C^0}' := \langle G_0', \Omega_0', \Sigma_0', \Lambda_0', L_0' \rangle :=
  C$, and let $i \in [t]$. Let $D^{i-1}:= \op{insert} (C^i, {C^{i-1}}')$ and
  $G^{i-1}_{R_i}$ be the set of all relational gates in $D^{i-1}$ labelled by
  the relational symbol in $R_i$. Let ${C^{i-1}}' := \langle G_i', \Omega_i',
  \Sigma_i', \Lambda_i', L_i' \rangle$ be the circuit formed from $D^{i-1}$ by
  replacing each gate in $g \in G^{i-1}_{R_i}$ with the gate
  $\Omega_{i}(\Lambda_{i-1}'(g))$ in $D^{i-1}$ (by repeated calls to the
  $\op{replace}$ function). Let $C' = {C^{t}}'$.
\end{definition}

It can be shown that $\op{compose}$ can be computed in time polynomial in the
sum of the sizes of the input circuits. We also note that the size of $C'$ is
less than or equal to the sum of the sizes of the input circuits. Moreover, if
the input circuits are all symmetric then the output circuit will be symmetric.
Also, for an input tuple $(C, \mathcal{C}, \tau_R)$, we have that if $C$ has
symmetric gates and transparent, and all the circuits in $\mathcal{C}$ are
transparent, then $C' := \op{compose}(C, \mathcal{C}, \tau_R)$ is transparent.
If $C$ contains a non-symmetric gates then, as the output gates of the circuits
in $\mathcal{C}$ are no longer labelled as output gates in $C'$, it may be that
some of these gates are syntactically-equivalent in $C'$, resulting in $C'$ not
being transparent.

In certain contexts we are interested composing gates such that if the input
circuits are transparent (resp. have unique labels) then the output circuits
will be transparent (resp. have unique labels). In this case, while composing
the circuits we apply the $\op{tag}$ function to the output gates of the
circuits in the input sequence of circuits. We now define the function
$\op{compose-and-tag}$ that implements this operation.

\begin{definition}
  The function $\op{compose-and-tag}$ maps a $(\mathbb{B}, \tau)$-circuit $C :=
  \langle G, \Omega, \Sigma, \Lambda, L \rangle$, a sequence of relation symbols
  $\tau_R := (R_1, \ldots, R_t)$ such that $\tau_R \subseteq \tau$, a sequence
  of circuits $\mathcal{C} := ( C^1, \ldots, C^t)$, where for each $i \in [t]$,
  $C^i = \langle G_i, \Omega_i, \Sigma_i, \Lambda_i, L_i \rangle$ is a
  $(\mathbb{B}_i, \tau_i)$-circuit, and $\tau_i \not\subset \tau_R$ and $C^i$
  computes a query of arity $\arty(R_i)$ to a circuit $\op{compose-and-tag}(C,
  \mathcal{C}, \tau_R, R) := C'$. The circuit $C'$ is defined as follows. Note
  that all circuits take in structures of size $n$.

  Let $D^0 := \op{compose}(C, \mathcal{C}, \tau_R)$. For each $i \in [t]$, let
  $D^{i} := \op{tag-all}(D^{i-1}, R, \img(\Omega^{-1}_i), \Omega^{-1}_i)$. Let
  $C' := D^t$.
\end{definition}

It is easy to see that $\op{compose-and-tag}$ can be computed in time polynomial
in the sun of the sizes of the input circuits. We also have that if the input
circuits are symmetric then the output circuits are symmetric. Furthermore, if
the input sequences of circuits and relations are both singletons, then if the
input circuits are all transparent (resp. have unique labels) then the output
circuit is transparent (resp. have unique labels). We also note that the output
circuit

\begin{definition}
  The function $\op{tag-inputs}$ maps a circuit $C := \langle G, \Omega, \Sigma,
  \Lambda, L \rangle$, a relation symbol $R \in \tau$, and $H \subseteq G$
\end{definition}

\subsection{Translating FPR to Symmetric Matrix Circuits}
% In this subsection we translate $\FPR$-formulas into families of transparent
% rank-circuits. In particular, we now show that for $\theta(\vec{x}) \in
% \FPR[\tau]$ there exists a $P$-uniform family of symmetric rank-circuits
% $(C_n)_{n \in \nats}$ such that the query computed by $(C_n)_{n \in \nats}$ is
% equal to the query defined by $\theta (\vec{x})$.

We say that a circuit $C$ \emph{translates} a formula $\psi(\vec{x})$ for $n \in
\nats$ if $C$ is a transparent symmetric rank-circuit of order $n$, $C$ is
defined over the same vocabulary as $\psi$, and $C$ computes the query defined
by $\psi(\vec{x})$ on structures of size $n$. We say a family of circuits
$(C_n)_{n \in \nats}$ \emph{translates} a formula $\psi (\vec{x})$ if $(C_n)_{n
  \in \nats}$ is $\PT$-uniform and for all $n \in \nats$, $C_n$ translates
$\psi(\vec{x})$ for $n$.

We now prove that for every $\FOrk$-formula $\theta(\vec{x})$ and $n \in \nats$
there exists a circuit $C$ that translates $\theta(\vec{x})$ for $n$. We
construct $C$ by including for each sub-formula $\psi \in \cl{\theta}$ and
assignment $\alpha \in [n]^{\free{\psi}}$ a gate $g_{\psi[\alpha]}$. The
children of $g_{\psi[\alpha]}$ are the gates of the form $g_{\phi[\beta]}$,
where $\phi$ is an immediate sub-formula of $\psi$ and $\beta$ is an assignment
to the free variables of $\psi$ that is compatible with the assignment $\alpha$.
The gate $g_{\psi[\alpha]}$ for an input structure over the universe $[n]$
evaluates to one if, and only if, $\psi$ holds for the input structure with the
assignment $\alpha$. We now construct the translation of $\theta$ for $n$, and
prove that the construction can be completed in a time bound that depends on $n$
and certain properties of the formula $\theta$.

\begin{lem}
  There is an algorithm that takes in a $\FOrk$-formula $\theta(\vec{x})$ and $n
  \in \nats$ and outputs a rank-circuit $C$ that translates $\theta(\vec{x})$
  for $n$. Moreover, there is a polynomial in $p$ such that for an input
  $(\theta(\vec{x}), n)$ the algorithm terminates in at most
  $p(n^{\width(\theta)}\vert \cl{\theta} \vert)$ many steps.
  \label{lem:translating-FOrk}
\end{lem}
\begin{proof}
  Let $\tau$ be the vocabulary of $\theta(\vec{x})$. Let $k := \width(\theta)$.
  Suppose $\tau$ is empty. Since the vocabulary of $\tau$ is empty, all of the
  atomic formula in $\theta(\vec{x})$ are of the form $y_1 = y_2$, where $y_1$
  and $y_2$ are variables. It can be shown that, for an assignment $\vec{a}$ to
  $\vec{x}$, we can evaluate $[n] \models \theta[\vec{a}]$ in time polynomial in
  $n^{k} \vert \cl{\theta} \vert$. Let $f : [n]^k \rightarrow \{0,1\}$ be
  defined such that $f (\vec{a}) = 1$ if, and only if, $[n] \models
  \theta[\vec{a}]$. Let $C := \langle G, \Sigma, \Omega, \Lambda, L \rangle$,
  where
  \begin{itemizens}
  \item $G = \{g_0, g_1\} \cup \{g_{\vec{a}} : \vec{a} \in [n]^{k}\}$,
  \item $\Sigma (g_0) = 0$, $\Sigma (g_1) = 1$ and $\Sigma (g_{\vec{a}}) =
    \land[1]$,
  \item $\Omega : [n]^k \rightarrow G$ is defined by $\Omega(\vec{a}) =
    g_{\vec{a}}$ for all $\vec{a} \in [n]^k$,
  \item $\Lambda = \emptyset$, and
  \item for $\vec{a} \in [n]^k$, $L(g_{\vec{a}}) : [1] \rightarrow G$ is defined
    by $L(g_{\vec{a}})(1) = g_{f(\vec{a})}$.
  \end{itemizens}

  Since $C$ has symmetric gates, $C$ is transparent. Notice that, since
  $\theta(\vec{x})$ defines a query on structures over the empty vocabulary, we
  have for all $\vec{a} \in [n]^k$, $\sigma \in \sym_n$, $f(\sigma \vec{a}) =
  f(\vec{a})$. For $\sigma \in \sym_n$, let $\pi_\sigma : G \rightarrow G$ be a
  bijection that fixes the constant gates and such that $\pi_{\sigma}
  g_{\vec{a}} = g_{\sigma \vec{a}}$ for all $\vec{a} \in [n]^k$. Then
  $\pi_\sigma \Omega (\vec{a}) = \pi_\sigma g_{\vec{a}} = g_{\sigma \vec{a}} =
  \Omega (\sigma \vec{a})$. It follows that $\pi_\sigma$ is an automorphism of
  the circuit extending $\sigma$, and $C$ is symmetric. We thus have that $C$
  translates $\theta(\vec{x})$ for $n$.

  Suppose $\tau$ is not empty. Let $R \in \tau$. We first define two helper
  formulas. For a variable $y$ let $\op{no-op}(y) := (R(y, y) \land (\neg R(y,
  y)))$. For a sequence of variables $\vec{y} = (y_1, \ldots, y_m)$ let
  $\op{tag} (\vec{y}) := \op{no-op}(y_1) \land (\op{no-op}(y_2) \land (
  \op{no-op}(y_2) \land ( \ldots \land (\op{no-op}(y_m)) \ldots)))$. Let
  $\lambda (\vec{x})$ be constructed from $\theta(\vec{x})$ by replacing each
  sub-formula $\psi(\vec{y})$ of the form $\rank^r_p \vec{w} \vec{z} . \phi$
  with the formula $\rank^r_p \vec{w}\vec{z} . ((\forall u . u = u) \land \phi)
  \land \op{tag}(\vec{w} \cup \vec{z})$. Since the sub-formulas added are
  tautologies, the matrix defined at this point remains unchanged, and so
  $\lambda (\vec{x})$ and $\theta (\vec{x})$ are equivalent. The idea is that
  $\op{tag}$ adds a tower of conjunctions of tautologies each referencing
  exactly one variable from $\vec{w} \cup \vec{z}$, and so encoding the order of
  the sequence of variables $\vec{w} \cup \vec{z}$ in the syntax of the formula.
  When we define a circuit from this formula, each input to the rank gate will
  be `tagged' with a unique assignment to $\vec{w} \cup \vec{z}$, preventing two
  inputs from being syntactically-equivalent, and so ensuring the circuit is
  transparent.

  % We also relabel the bound variables in $\lambda(\vec{x})$ such that there
  % are
  % no two sub-formulas $\psi$ and $\psi'$ such that $\psi$ and $\psi'$ are
  % equal
  % up to a renaming of those variables that are bound in $\psi$ or $\psi'$.

  % Let $\psi (\vec{y}) \in \FOrk$ and $\alpha \in [n]^{\vec{y}}$. Let
  % $\psi[\alpha]$ be the result of substituting $\vec{y}$ for $\alpha(y)$ in
  % $\psi(\vec{y})$. If $\psi$ is of the form $y_1 = y_2$. Then $\psi [\alpha] =
  % 1$ if $\alpha(y_1) = \alpha (y_2)$ and $\psi[\alpha] = 0$ otherwise.

  We note that $\lambda(\vec{x})$ has the same width as $\theta(\vec{x})$, and
  that the number of sub-formulas in $\lambda(\vec{x})$ is at most $\vert
  \cl{\theta} \vert (5 + 4\width(\theta))$. We now construct a circuit $C$ and
  show that $C$ translates $\lambda(\vec{x})$ (and hence $\theta(\vec{x})$) at
  $n$.

  % We first
  % pre-proccess the formula $\theta(\vec{x})$

  % Let $\op{no-op}(y) := (R(y, y) \land (\neg R(y, y)))$. Let $\vec{y} := (y_1,
  % \ldots, y_k)$ be a vector of variables and $\op{add-tag} (\vec{y}) :=
  % \op{no-op}(y_1) \land (\op{no-op}(y_2) \land ( \op{no-op}(y_2) \land (
  % \ldots
  % \land (\op{no-op}(y_k)) \ldots)))$. From $\theta (\vec{x})$ we construct a
  % new
  % formula $\theta' (\vec{x})$ by replacing each sub-formula $\phi(\vec{y})$ of
  % the form $\rank^r_p \vec{w} \vec{z} . \psi$ by the sub-formula $\rank^r_p
  % \vec{w}\vec{z} . ((\forall u . u = u) \land \phi) \land \op{add-tag}(\vec{w}
  % \cup \vec{z})$.

  Let $\vec{y}$ be a sequence of variables. We abuse notation and write $\alpha
  :\vec{y} \rightarrow A$ to indicate that $\alpha$ maps the variables in the
  sequence $\vec{y}$ to elements of the set $A$.

  Fix $n \in \nats$. Let $\psi \in \FOrk$ and $\alpha \in [n]^{\free {\psi}}$.
  We let $\psi[\alpha]$ be the result of substituting each occurrence of the
  free variable $y \in \free{\psi}$ in $\psi$ with $\alpha(y)$. If $\psi$ is of
  the form $y_1 = y_2$. Then $\psi [\alpha] = 1$ if $\alpha(y_1) = \alpha (y_2)$
  and $\psi[\alpha] = 0$ otherwise. We call $\psi[\alpha]$ a \emph{ground
    formula}. For each $\psi \in \cl{\theta}$ let $G_\psi := \{g_{\phi[\alpha]}
  : \alpha \in [n]^{\free{\psi}} \}$. Let $G := \bigcup_{\psi \in \cl{\lambda}}
  G_{\psi}$. We define $\Sigma$, $\Lambda$ and $L$ as follows. Let $\Sigma (g_1)
  = 1$ and $\Sigma (g_0) = 0$. Let $g := g_{\psi[\alpha]} \in G$.

  \begin{itemizens}
  \item If $\psi[\alpha] = T(\vec{a})$ for some $T \in \tau$ and $\vec{a} \in
    \arty(T)$, then $\Sigma (g) = T$ and $\Lambda_T (g) = \vec{a}$.
  \item If $\psi[\alpha]$ is $0$ or $1$, then $\Sigma (g) = \psi[\alpha]$.
  \item If $\psi$ is of the form $\rank^r_p \vec{w} \vec{z}. \phi$, let $A =
    [n]^{\vert \vec{w}\vert}$ and $B = [n]^{\vert \vec{z}\vert}$, and $\Sigma
    (g_{\psi, \alpha}) = \rank[A, B]$. Let $L(g) : A \times B \rightarrow G$ be
    defined by $L(g)(\vec{a},\vec{b}) = g_{\phi[\beta]}$, where $\beta \in
    [n]^{\free{\phi}}$, and where $\alpha$, and the assignments
    $\alpha^{\vec{w}}_{\vec{a}}$ and $\alpha^{\vec{z}}_{\vec{b}}$, are
    compatible with $\beta$.
  \item If $\psi$ is of the form $Q z . \phi(\vec{y}, z)$, for $Q \in \{\forall,
    \exists\}$, then if $Q = \forall$ we have $\Sigma (g) = \land [n]$,
    otherwise $\Sigma (g) = \lor [n]$. Let $L(g) : [n] \rightarrow G$ be defined
    for $i \in [n]$ by $L(g)(i) = g_{\phi[\beta_i]}$, where $\beta_i \in
    [n]^{\free{\psi}}$, and where both $\alpha$ and $\alpha^{z}_{i}$ are
    compatible with $\beta_i$.
  \item If $\psi$ is of the form $\phi_1 \land \phi_2$, then $\Sigma(g) =
    \land[2]$ and $L(g) : [2] \rightarrow G$ is defined for $i \in [2]$ by
    $L(g)(i) = g_{\phi_i[\beta_i]}$, where $\beta_i \in [n]^{\free{\psi}}$ and
    $\beta_i$ is compatible with $\alpha$. The same approach is used in the
    disjunctive case.
  \item if $\psi$ is of the form $\neg \phi$ then $\Sigma (g) = \neg$ and $L(g):
    [1] \rightarrow G$, where $L(g)(1) = g_{\phi[\alpha]}$.
  \end{itemizens}

  We let $\Omega : [n]^k \rightarrow G$ be defined by $\Omega (\vec{a}) =
  g_{\lambda [\alpha]}$, where $\alpha := \alpha^{\vec{x}}_{\vec{a}}$, for all
  $\vec{a} \in [n]^k$. Suppose $\Omega (\vec{a}) = g_c$, where $ c \in {0,1}$.
  Then $\lambda$ (and $\theta$) is of the form $(y_1 = y_2)$. But then the
  vocabulary of $\theta$ is empty, contradicting our assumption. Suppose $\Omega
  (\vec{a}) = \Omega(\vec{b})$ for $\vec{a}, \vec{b} \in [n]^k$. Then
  $g_{\lambda[\alpha^{\vec{x}}_{\vec{a}}]} =
  g_{\lambda[\alpha^{\vec{x}}_{\vec{b}}]}$, and so $\alpha^{\vec{x}}_{\vec{a}} =
  \alpha^{\vec{x}}_{\vec{b}}$, and $\vec{a} = \vec{b}$. It follows that $\Omega$
  is injective. Let $T \in \tau$, and let $g_{\psi[\alpha]}, g_{\phi[\beta]} \in
  G$ be such that $\Lambda_T (g_{\psi[\alpha]}) = \Lambda_T (g_{\phi[\beta]})$.
  Then $\psi[\alpha] = \Lambda_T (g_{\psi[\alpha]}) = \Lambda_T
  (g_{\phi[\beta]}) = \phi[\beta]$, and so $g_{\psi[\alpha]} = g_{\phi[\beta]}$.
  It follows that $\Lambda_T$ is injective. We thus have that $C := \langle G,
  \Omega, \Sigma, \Lambda, L \rangle$ is a circuit of order $n$.

\begin{claim}
  The circuit $C$ is symmetric.
\end{claim}
\begin{proof}
  Let $\pi_\sigma : G \rightarrow G$ be a function that fixes constant gates for
  such that for all non-constant gates $g_{\psi[\alpha]} \in G$, $\pi_{\sigma}
  g_{\psi [\alpha]} = g_{\psi[\sigma \alpha]}$. It is easy to see that
  $\pi_\sigma$ is a bijection. We now show that $\pi_{\sigma}$ is an
  automorphism of the circuit extending $\sigma$. Let $g_{\psi[\alpha]} := g \in
  G$. Since the formula is unchanged by $\pi_\sigma$, we have $\Sigma(g) =
  \Sigma (\pi_{\sigma} g)$. Suppose $g$ is an input gate. If $g$ is a constant
  gate then $\pi_\sigma g = g$. If $g$ is a relational gate, then $\psi$ is of
  the form $T(\vec{y})$ for some relation symbol $T$ and $\Lambda_T
  (g_{\psi[\alpha]}) = \alpha \vec{y}$, and $\Lambda (\pi_{\sigma}
  g_{\psi[\alpha]}) = \Lambda g_{\psi[\sigma \alpha]} = \sigma \alpha \vec{y} =
  \sigma \Lambda_T (g_{\psi[\alpha]})$. Suppose $g$ is an internal gate. From
  the construction of the circuit we have that if $g_{\phi[\beta]} \in H_g$ then
  $\phi$ must be an immediate sub-formula of $\psi$ and so $g_{\phi[\sigma
    \beta]}$ is a child of $g_{\psi[\sigma \alpha]} = \pi_{\sigma} g$. It
  follows that $H_{\pi_\sigma g} = \pi_\sigma H_g$. If $g$ is a symmetric gate,
  this is sufficient to conclude that $\pi_{\sigma} L(g)$ is
  isomorphism-equivalent to $L(\pi_{\sigma} g)$. If $g$ not a symmetric gate
  then it is a rank gate, and so $\psi$ is of the form $\rank^r_p \vec{w}
  \vec{z} \phi$. Let $\vec{a} \in A$ and $\vec{b} \in B$, and note that
  $\pi_\sigma L(g) (\vec{a}, \vec{b}) = \pi_{\sigma} g_{\phi[\beta]}$, where
  $\beta \in [n]^{\free {\phi}}$ is compatible with $\alpha$,
  $\alpha^{\vec{w}}_{\vec{a}}$ and $\alpha^{\vec{z}}_{\vec{b}}$. It follows that
  $\sigma \beta$ is compatible with $\sigma \alpha$, $\sigma
  \alpha^{\vec{w}}_{\vec{a}} = \alpha^{\vec{w}}_{\sigma \vec{a}}$ and
  $\alpha^{\vec{z}}_{\vec{b}} = \alpha^{\vec{z}}_{\sigma \vec{b}}$. We thus have
  that $L(\pi_{\sigma} g)(\sigma \vec{a}, \sigma \vec{b}) = g_{\phi[\sigma
    \beta]} = \pi_{\sigma} g_{\phi[\beta]} = \pi_{\sigma} L(g)(\vec{a},
  \vec{b})$. Suppose $g$ is an output gate, and let $\vec{a} \in [n]^k$ and
  $\alpha := \alpha^{\vec{x}}_{\vec{a}}$. We have that $\pi_{\sigma} \Omega
  (\vec{a}) = \pi_\sigma g_{\lambda[\alpha]} = g_{\lambda[\sigma \alpha]} =
  \Omega (\sigma \alpha \vec{x}) = \Omega (\sigma \vec{a})$. This concludes the
  proof of the claim.
\end{proof}

Let $\mathcal{A}$ be a $\tau$-structure with universe $U$ of size $n$ and let
$\gamma \in [n]^{\underline{U}}$. It can be shown that for all $\psi \in
\cl{\lambda}$ and $\alpha \in U^{\free{\psi}}$, $\mathcal{A} \models
\psi[\alpha]$ if, and only if, $C[\gamma \mathcal{A}](g_{\psi[\gamma\alpha]}) =
1$. It follows that $\vec{a} \in U^k$ is an element of $Q(\theta(\vec{x}))$ if,
and only if, $\mathcal{A} \models \lambda[\alpha^{\vec{x}}_{\vec{a}}]$ if, and
only if, $C[\gamma \mathcal{A}](g_{\lambda[\alpha^{\vec{x}}_{\gamma \vec{a}}]})
= 1$. Since $g_{\lambda [\alpha^{\vec{x}}_{\gamma \vec{a}}]} = \Omega (\gamma
\vec{a})$, we have that $C$ computes the query defined by $\theta(\vec{x})$ on
structures of size $n$.

\begin{claim}
  The circuit $C$ is transparent.
\end{claim}
\begin{proof}
  If $C$ has symmetric gates then $C$ is transparent. Suppose $C$ contains a
  non-symmetric gate. Let $g_{\phi[\alpha]}:= g \in G$ be a rank gate and
  suppose $\psi$ is of the form $\rank^r_p \vec{w} \vec{z}. \zeta$, where $\zeta
  = \zeta_1 \land \zeta_2$ and where, $\zeta_1 = ((\forall u. u = u) \land
  \phi)$, and $\zeta_2 = \land (\op{tag}(\vec{w} \cup \vec{z}))$. Let $\beta_1 ,
  \beta_2 \in [n]^{\free{\zeta}}$ both compatible with $\alpha$, and let $h_1 =
  g_{\zeta[\beta_1]}$ and $h_2 = g_{\zeta[\beta_2]}$. For $i \in [2]$ we have
  that $H_{h_i} = \{g_{\zeta_1[\beta_i]}, g_{\zeta_2[\beta_i]}\}$. Since
  $\zeta_1$ contains a sub-formula of the form $(u = u)$, while $\zeta_2$
  contains no sub-formula with equality, we have that a child of a child of
  $g_{\zeta_1[\beta_1]}$ is a constant gate while none of the children of the
  children of $g_{\zeta_2[\beta_2]}$ are constant gates, and thus $g_{\zeta_1
    [\beta_1]} \not\equiv g_{\zeta_2[\beta_2]}$. It follows that $h_1 \equiv
  h_2$ if, and only if, both $g_{\zeta_1 [\beta_1]} \equiv g_{\zeta_1
    [\beta_2]}$ and $g_{\zeta_2[\beta_1]} \equiv g_{\zeta_2[\beta_2]}$.
  Moreover, the fact that $\zeta_2$ contains an ordered sequence of conjunctions
  of tautologies, with each tautology referencing a different variable,
  guarantees that $g_{\zeta_2[\beta_1]} \equiv g_{\zeta_2[\beta_2]}$ if, and
  only if, $\beta_1$ and $\beta_2$ agree on $\free{\zeta_2}$. But we already
  have that $\beta_1$ and $\beta_2$ are compatible with $\alpha$, and so agree
  on $\free{\zeta_1}$. It follows that $g_{\zeta_2[\beta_1]} \equiv
  g_{\zeta_2[\beta_2]}$ if, and only if, $\beta_1 = \beta_2$. This allows us to
  conclude that $g_{\zeta[\beta_1]} \equiv g_{\zeta[\beta_2]}$ if, and only if,
  $\beta_1 = \beta_2$ if, and only if, $g_{\zeta[\beta_1]} =
  g_{\zeta[\beta_2]}$. It follows that $L(g)$ is injective, and no two $h_1, h_2
  \in H_g$ are syntactically-equivalent. We conclude that $g$ has unique labels,
  and so $C$ is transparent.
\end{proof}

We notice that $\vert C \vert \leq n^{\width(\lambda)} \vert \cl{\lambda} \vert
\leq n^{\width(\theta)} \vert \cl{\theta} \vert (5 + \width(\theta))$. It is not
hard to see that $C$ (and $\lambda(\vec{x})$) can be constructed in a number of
steps bounded by a polynomial in the size of $C$. This completes the proof of
the result.

% = C[\gamma \mathcal{A}](\Omega(\gamma \vec{a}))$

% Since $h_{}$Since $\pi_\sigma$ is a bijection, we have that $f$ is a
% bijection. Moreover, we have that $\pi L(g) (a) = L(g) $

% We also have that $\mathcal{A} \models \psi[\gamma \alpha]$ if, and only if,
% $T^{\mathcal{A}} (\gamma \alpha \vec{y})$ if, and only if, $C[\gamma
% \mathcal{A}](g) = 1$

% Suppose $g$ is an internal gate, and we have both the results are true for all
% gates that come before $g$ in the topological ordering of the circuit. Suppose
% $\psi$ is an application of is an application of a logical connective ($land$
% or $\lor$) or negation ($\neg$). We then have that $g$ is a symmetric gate,
% and $f : H_{g} \rightarrow H_{\pi_\sigma g}$ defined by $f(h) = \pi_\sigma h$
% is a bijection.

% $\pi_\sigma g_{\psi[\alpha]} = \Lambda_T ()$, for some formula $\psi$
% containing the relation symbol $T$ and assignment $\alpha$, where $\alpha$
% assigns the free variables in $\psi$ to $\vec{a}$. Then $\pi_\sigma g =
% \Lambda^{-1}_T (\sigma \Lambda_T (g)) = g_{\psi[\sigma \alpha]}$. Suppose $g$
% is an internal gate. If $g = g_{psi[\alpha]}$ is an application of a logical
% connective ( $land$ or $\lor$) or negation ($\neg$) then $\pi_{\sigma}g =
% g_{\psi[\sigma \alpha]}$. If $g$

% It can be shown by induction on the structure of the circuit that all
% $g_{\psi[\alpha]} \in G$

% We should note that $C'$ is not, in general, a circuit. This is because it has
% (potentially) more than three constant gates, and perhaps more than one
% relational gate for each assignment and relation symbol. Let $C$ be formed
% from $C'$ by identifying all gates $g$ and $g'$ such that (i) $g$ and $g'$ are
% constant gates and $\Sigma' (g) = \Sigma'(g')$, or (ii) $g$ and $g'$ are
% relational gates, $T := \Sigma' (g) = \Sigma' (g') $, and $\Lambda_T' (g) =
% \Lambda_T'(g')$. We now have $C := \langle G, \Omega, \Sigma, \Lambda, L
% \rangle$. If the domain of $\Omega'$ includes a constant gate, then
% $\theta(\vec{x}) = \Lambda (\vec{x}) = (x_1 = x_2)$

% Notice that, formally, $C'$ is not a circuit, in that it contains too many
% relational gates.

% We now show that $C$ translates $\theta(\vec{x})$ at $n$. It is obvious that
% $C$ has order $n$. For $\sigma \in \sym_n$ let $\pi_\sigma : G \rightarrow G$
% be a defined such that $\pi_{\sigma} (g) = g$ if $g \in G_c$ and
% $\pi_\sigma(g_{\psi, \alpha}) = g_{\psi, \sigma \alpha}$ for all $\psi \in
% \cl{\lambda}$ and $\alpha \in n^{\free{\psi}}$. It can be shown that
% $\pi_\sigma$ is an automorphism of $C$ extending $\sigma$. It follows that $C$
% is symmetric.

% Let $\mathcal{A}$ be a $\tau$-structure with a universe $U$ of size $n$ and
% let $\gamma U^{\underline{[n]}}$. Let $\psi (\vec{y}) \in \cl{\lambda}$ and
% let $\alpha \in [n]^{\vec{y}}$. It can be shown by induction on the structure
% of the circuit that for all $\psi (\vec{y})$ and $\alpha \in [n]^{\vec{y}}$,
% $C[\gamma \mathcal{A}](g_{\psi, \alpha}) = 1$ if, and only if, $\mathcal{A}
% \models \psi [\gamma (\alpha \vec{y})]$. It follows that

% Let $\vec{a} \in [n]^k$, and $\alpha \in [n]^{\vec{x}}$ be defined such that
% $\alpha (x_i) = a_i$ for all $i \in [k]$. Then $C[\Omega]$

% It is obvious that $C$ has order $n$. Let $\sigma \in \sym_n$. Let $\pi_\sigma
% : G \rightarrow G$ be a defined such that $\pi_{\sigma} (g) = g$ if $g \in
% G_c$ and $\pi_\sigma(g_{\psi, \alpha}) = g_{\psi, \sigma \alpha}$ for all
% $\psi \in \cl{\lambda}$ and $\alpha \in n^{\free{\psi}}$. It can be shown that
% $\pi_\sigma$ is an automorphism of $C$ extending $\sigma$. It follows that $C$
% is symmetric.

% Moreover, let $\alpha \in [n]^{\free{\theta}}$ be an assignment that maps the
% sequence of variables $\vec{x}$ to the tuple $\vec{a} \in [n]^k$. It follows
% that $\Omega (\vec{a}) = g_{\lambda, \alpha}$ evaluates to $1$ If $C$ does not
% contain a rank gate then it is transparent. Suppose $C$ does contain a rank
% gate. Let $g := g_{\psi, \alpha} \in G$ be a rank gate and suppose $\psi
% (\vec{y}) := \rank^r_p \vec{w} \vec{z}. \zeta$, where $\zeta = \zeta_1 \land
% \zeta_2$ and $\zeta_1 = ((\forall u. u = u) \land \phi)$ and $\zeta_2 = \land
% (\op{tag}(\vec{w} \cup \vec{z}))$. Suppose $h_1, h_2 \in H_g$ are
% syntactically-equivalent. It follows that there exist assignments $\beta_1,
% \beta_2 \in [n]^{\vec{w} \cup \vec{z} \cup {y}}$ such that $\alpha \subseteq
% \beta_1$ and $\alpha \subseteq \beta_2$ and $h_1 = g_{\zeta, \beta_1}$ and
% $h_2 = g_{\zeta, \beta_2}$. Then $H_{h_i} = \{g_{\zeta_1, \beta_i},
% g_{\zeta_2, \beta_i}\}$ for $i \in [2]$. Note that a child of a child of
% $g_{\zeta_1, \beta_1}$ is a constant gate while none of the children of the
% children of $g_{\zeta_2, \beta_2}$ are constant gates, and thus $g_{\zeta_1,
% \beta_1} \not\equiv g_{\zeta_2, \beta_2}$ are not syntactically-equivalent. It
% follows that $h_1 \equiv h_2$ if, and only if, both $g_{\zeta_1, \beta_1}
% \equiv g_{\zeta_1, \beta_2}$ and $g_{\zeta_2, \beta_1} \equiv g_{\zeta_2,
% \beta_2}$. However, the construction of $\zeta_2$ (and the fact that $\beta_1$
% and $\beta_2$ agree on the variables in $\vec{y}$) guarantees that
% $g_{\zeta_2, \beta_1} \equiv g_{\zeta_2, \beta_2}$ if, and only if, $\beta_1 =
% \beta_2$ are equal. It follows that $h_1 = g_{\zeta, \beta_1} = g_{\zeta,
% \beta_2} = h_2$. The construction of the circuit guarantees that $g$ has
% injective labels. It follows that $g$ has unique labels, and the circuit $C$
% is transparent.

% and $h := g_{\phi, \beta}$ and $h' := g_{\phi, \beta'}$ are
% syntactically-equivalent. From the construction of $\theta'$ we have that
% $\phi$ must be of a particular form. Suppose $\phi = \eta \land
% \op{add-tag}(\vec{w} \cup \vec{z})$. Let $\gamma$ and $\gamma'$ be the
% restrictions of $\beta$ and $\beta'$ to $\vec{w} \cup \vec{z}$, respectively,
% and let $\zeta(\vec{w}\cup\vec{z}) := \op{add-tag}(\vec{w} \cup \vec{z})$. The
% $\op{add-tag} (\vec{w} \cup \vec{z})$

% We notice that $\vert C \vert \leq n^{\width(\lambda)} \vert \cl{\lambda}
% \vert \leq n^{\width(\theta)} \vert \cl{theta} \vert (5 + \width(\theta))$. It
% can be shown that $C$ (and $\lambda(\vec{x})$ can be constructed in a number
% of steps bounded by a polynomial in the size of $C$.

% If $C'$ contains no rank gates then output $C := C'$. Suppose $C'$ contains at
% least one rank-gate. We note add to $C'$ a gadget that ensures that no two
% input gates to a rank gate can be syntactically-equivalent. Let $\psi \in
% \cl{\theta}$, and suppose $\psi (\vec{y}) = \rank^r_p \vec{w} \vec{z}. \phi$.
% We may think of the assignments to the variables $\vec{w}$ and $\vec{z}$ as
% being the row and column indexes of the matrix. We

% Then for $\alpha \in [n]^{\free (\psi)}$, $H_{g_{\psi, \alpha}}$ consists of
% gates of the form $g_{\phi, \beta}$, where $\beta \in [n]^{\vec{w} \cup
% \vec{z} \cup \vec{y}}$ and $\alpha subseteq \beta$.

% Let $T_{\psi, \alpha} : H_{g_{\psi, \alpha}} \rightarrow [n]^{\vert \vec{w}
% \vert + \vert \vec{z} \vert}$ be defined by $T_{\psi, \alpha} (g_{\phi,
% \beta}) = \beta (\vec{z} \cup \vec{w})$ for all $g_{\phi, \beta} \in H_g$.
% What $T_{\psi, \alpha}$ does is assign to each child of $g_{\psi, \alpha}$ an
% ordered sequence formed Let $T$ be the union of all $T_{\psi}$ and $H$ the
% union of all $G_{\psi}$ for all $\psi \in \cl{\theta}$ an application of the
% rank operator. Let $C := \op{tag-all}(C', R, H, T)$. We note that
% $\op{tag-all}$ may be computed in time polynomial in the size of the input
% circuit.

% Let $a \in [n]$ and let $\op{no-op} (a) := R(a, a) \lor (\neg (R(a, a))$. For
% $\psi (\vec{y})$

% % Let $(b_1, \ldots, b_m) := \beta (\vec{w} \cup \vec{z})$. The $\op{tag-all}
% % function replaces each input $g_{\phi, \beta}$ of the rank gate $g_{\psi,
% % \alpha}$ with the circuit $g_{\phi, \beta} \land \op{no-op}(b_1) \land
% % (\op{no-op}(b_2) \land ( \op{no-op}(b_2) \land (\ldots \land
% % (\op{no-op}(b_m))
% % \ldots)))$. This construction does not effect the evaluation of any rank
% % gate
% % in the cirucit and ensures that no two to $g_{\psi,\alpha}$ in $C$ are
% % syntactically-equivilent.

% $of the form $$R(a, a) \lor (\neg R(a, a))$ that encode the assignment to
% $\vec{w}\cup \vec{z}$.

% We have that $C$ is transparent, has order $n$, and $C$ and $C'$ decide the
% same query. It remains to show that $C$ is still symmetric.

% However, the $\op{tag-all}$ operation adds a tower of tautologies of the form
% $R(a, \ldots, a) \lor (\neg R (a, \ldots, a))$

% Let $g := g_{\psi, \alpha} \in G$ be a rank gate and let $h, h' \in H_g$.
% Suppose $\psi (\vec{y}) := \rank^r_p \vec{w} \vec{z}. \phi$ and $h := g_{\phi,
% \beta}$ and $h' := g_{\phi, \beta'}$ are syntactically-equivalent. From the
% construction of $\theta'$ we have that $\phi$ must be of a particular form.
% Suppose $\phi = \eta \land \op{add-tag}(\vec{w} \cup \vec{z})$. Let $\gamma$
% and $\gamma'$ be the restrictions of $\beta$ and $\beta'$ to $\vec{w} \cup
% \vec{z}$, respectively, and let $\zeta(\vec{w}\cup\vec{z}) :=
% \op{add-tag}(\vec{w} \cup \vec{z})$. The $\op{add-tag} (\vec{w} \cup \vec{z})$

% The structure of this formula gives us that $h$ is syntactically-equivalent to
% $h'$ if, and only if, $g_{\zeta , \beta'}$

% We also note that $C$ has order $n$.
\end{proof}

We now use Lemma~\cite{lem:translating-FOrk} to prove that for every $\FPR$
formula there is a corresponding $\PT$-uniform family of transparent symmetric
rank-circuits that decides the same query.

\begin{thm}
  For each $\FPR$-formula $\theta(\vec{x})$ there exists a family of circuits
  $(C_n)_{n \in \nats}$ that translates $\theta(\vec{x})$.
  \label{thm:translating-formulas-to-circuits}
\end{thm}
\begin{proof}
  Let $\theta(\vec{x}) \in \FPR[\tau]$ and $t := \width (\theta)$. We begin by
  showing that there exists a $P$-uniform family of formulas
  $(\theta(\vec{x}))_{n \in \nats}$ such that for each $n \in \nats$, we have
  $\theta_n(\vec{x}) \in \FOrk[\tau]$, $\width(\theta_n) \leq 4 t$, and
  $\theta_n(\vec{x})$ is equivalent to $\theta(\vec{x})$ on structures of size
  $n$. This result follows from two observations.

  First, using the translation in~\cite{Dawar09logicswith}, we have that for any
  $\phi(\vec{x}) \in \FPR[\tau]$ of the form $\lfp{V, \vec{y}}{\psi(\vec{x},
    \vec{y}, V)}(\vec{x})$ , there is a mapping $n \mapsto \phi_n$, computable
  in time polynomial in both $n$ and the size of $\phi(\vec{x})$ such that for
  all $n \in \nats$, we have $\phi_n(\vec{x}) \in \FOR[\tau]$, $\width (\phi_n)
  \leq 2 \cdot \width (\phi)$, and $\phi_n(\vec{x})$ and $\phi(\vec{x})$ are
  equivalent on structures of size $n$.

  Second, using the translation in~\cite{libkin2004elements}, we have that if
  $\phi (\vec{x}) \in \FOR[\tau]$ is of the form $\rank (\vec{y}\vec{\mu} \leq
  \vec{t}, \vec{z}\vec{\nu} \leq \vec{s}, \pi \leq r) . \psi (\vec{y}\vec{\mu},
  \vec{z} \vec{\nu})$, then there is a mapping $n \mapsto \phi_n$ computable in
  time polynomial in both $n$ and the size of $\phi(\vec{x})$ such that for all
  $n \in \nats$, $\phi_n(\vec{x}) \in \FOrk[\tau]$, $\width(\phi_n) \leq 2 \cdot
  \width(\phi)$, and $\phi_n(\vec{x})$ and $\phi(\vec{x})$ are equivalent on
  structures of size $n$.

  We thus derive the required $\PT$-uniform family $(\theta_n(\vec{x}))_{n \in
    \nats}$ by applying the first translation and then the second. We have from
  Lemma~\ref{} that there exists an algorithm that takes as input $n$ and
  $\theta_n$, runs in time polynomial in $n^{\width(\theta_n)} \vert
  \cl{\theta_n} \vert$, and outputs a circuit $C_n = T(\theta_n, n)$ that
  translates $\theta_n$ for $n$. We note that $\width{\theta_n} \leq 4t$ and
  $\cl{\theta_n}$ must be bounded by a polynomial in $n$ (since $\theta_n$ is a
  member of a $P$-uniform family of circuits). It follows that the function $n
  \mapsto C_n$ is computable in time polynomial in $n$, and so $(C_n)_{n \in
    \nats}$ translates $\theta(\vec{x})$.
\end{proof}

% \begin{proof}
%   We say that a circuit $C$ \emph{translates} a formula $\psi(\vec{x})$ for $n
%   \in \nats$ if $C$ is a transparent symmetric rank-circuit of order $n$ and
%   $C$ computes the query defined by $\psi(\vec{x})$ on structures of size $n$.
%   We say a family of circuits $(C_n)_{n \in \nats}$ \emph{translates} a
%   formula $\psi (\vec{x})$ if $(C_n)_{n \in \nats}$ is $\PT$-uniform and for
%   all $n \in \nats$, $C_n$ translates $\psi(\vec{x})$ for $n$.

%   %   We have from Anderson and Dawar~\cite{AndersonD17} that for every
%   %   formula
%   %   $\phi(\vec{x}) \in \FPC[\tau]$ there is a $\PT$-uniform family of
%   %   symmetric
%   %   circuits $(C^{\phi}_n)_{n \in \nats}$ with unique labels and symmetric
%   %   gates
%   %   that translates $\phi(\vec{x})$.

%   Let $\theta(\vec{x}) \in \FPR[\tau]$ and $t := \width (\theta)$. We begin by
%   showing that there exists a $P$-uniform family of formulas
%   $(\theta(\vec{x}))_{n \in \nats}$ such that for each $n \in \nats$, we have
%   $\theta_n(\vec{x}) \in \FOrk[\tau]$, $\width(\theta_n) \leq 4 t$, and
%   $\theta_n(\vec{x})$ is equivalent to $\theta(\vec{x})$ on structures of size
%   $n$. This result follows from two observations.

%   First, using the translation in~\cite{Dawar09logicswith}, we have that for
%   any $\phi(\vec{x}) \in \FPR[\tau]$ of the form $\lfp{V,
%   \vec{y}}{\psi(\vec{x}, \vec{y}, V)}(\vec{x})$ , there is a mapping $n
%   \mapsto \phi_n$, computable in time polynomial in both $n$ and the size of
%   $\phi(\vec{x})$ such that for all $n \in \nats$, we have $\phi_n(\vec{x})
%   \in \FOR[\tau]$, $\width (\phi_n) \leq 2 \cdot \width (\phi)$, and
%   $\phi_n(\vec{x})$ and $\phi(\vec{x})$ are equivalent on structures of size
%   $n$.

%   Second, using the translation in~\cite{libkin2004elements}, we have that if
%   $\phi (\vec{x}) \in \FOR[\tau]$ is of the form $\rank (\vec{y}\vec{\mu} \leq
%   \vec{t}, \vec{z}\vec{\nu} \leq \vec{s}, \pi \leq r) . \psi
%   (\vec{y}\vec{\mu}, \vec{z} \vec{\nu})$, then there is a mapping $n \mapsto
%   \phi_n$ computable in time polynomial in both $n$ and the size of
%   $\phi(\vec{x})$ such that for all $n \in \nats$, $\phi_n(\vec{x}) \in
%   \FOrk[\tau]$, $\width(\phi_n) \leq 2 \cdot \width(\phi)$, and
%   $\phi_n(\vec{x})$ and $\phi(\vec{x})$ are equivalent on structures of size
%   $n$.

%   We thus derive the required $\PT$-uniform family $(\theta_n(\vec{x}))_{n \in
%   \nats}$ by applying the first translation and then the second. Let $x$ be a
%   variable and $R \in \tau$. Let $\op{no-op}(x) := R(x, x) \land (\neg R(x,
%   x))$. Let $\vec{x} := (x_1, \ldots, x_k)$ and $\op{add-tag} (\vec{x}) :=
%   \op{no-op}(x_1) \land (\op{no-op}(x_2) \land ( \ldots \land
%   (\op{no-op}(x_k)) \ldots)$. For each $\theta_n (\vec{x})$ we construct a new
%   formula $\theta_n' (\vec{x})$ by replacing each sub-formula $\phi(\vec{y})$
%   of $\theta_n(\vec{x})$ of the form $\rank^r_p \vec{w} \vec{z} . \psi$ by
%   $\rank^r_p \vec{w}\vec{z} . (\psi \land \op{add-tag}(\vec{w} \cup
%   \vec{z}))$. It is easy to see that we can construct $\theta_n'$ from
%   $\theta_n$ in time polynomial in $\theta_n$. Moreover, $\theta_n'$ and
%   $\theta_n$ are equivalent and have the same width.

%   Fix $n \in \nats$. We now construct a circuit $C_n$ that translates
%   $\theta_n'$ (and hence $\theta_n$ and $\theta$) at $n$. For each $\psi \in
%   \cl{\theta_n}$ let $G_\psi := \{g_{\psi, \alpha} : \alpha \in
%   [n]^{\free{\psi}} \}$. Let $G := \bigcup_{\psi \in \cl{\theta_n}} G_{\psi}$.
%   Let $g_{\psi, \alpha} \in G$. We define $\Sigma$, $\Lambda$ and $L$ as
%   follows.
%   \begin{itemizens}
%   \item If $\psi (\vec{y}) = R(\vec{y})$, for some $R \in \tau$, then $\Sigma
%     (g_{\psi, \alpha}) = R$ and $\Lambda_R (g_{\psi, \alpha}) = \alpha
%     \vec{y})$.
%   \item If $\psi (y_1, y_2) = (y_1 = y_2)$, then if $\alpha (y_1) = \alpha
%     (y_2)$, $\Sigma (g_{\psi, \alpha}) = 1$, else $\Sigma (g_{\psi, \alpha}) =
%     0$.
%   \item if $\psi (\vec{y}) = \rank^r_p \vec{w} \vec{z}. \phi$, then $\Sigma
%     ()$
%   \end{itemizens}

%   We first consider the base case. Suppose $\psi(\vec{y})$ is an atomic
%   formula. Then $\psi(\vec{y}) \in \FPC[\tau]$, and so, using the translation
%   defined by Anderson and Dawar~\cite{AndersonD17}, there exists a family of
%   circuits $(C^{\psi}_n)_{n \in \nats}$ that translates $\psi(\vec{y})$. Let
%   $T_n(\psi(\vec{y})) := C^{\psi}_n$.

%   We now consider the conjunctive case. Suppose $\psi(\vec{y})$ is a
%   sub-formula of $\theta_n(\vec{x})$ such that $\psi (\vec{y}) =
%   \phi_1(\vec{y}) \land \phi_2(\vec{y})$. Let $R_{\phi_1}$ and $R_{\phi_2}$ be
%   two $\vert \vec{y} \vert$-ary relation symbols not in $\tau$ and let $\psi'
%   (\vec{y}) := R_{\phi_1}(\vec{y}) \land R_{\phi_2}(\vec{y})$. Then
%   $\psi'(\vec{y}) \in \FPC[\tau]$ and so, using the translation defined by
%   Anderson and Dawar, there exists a circuit family $(C^{\psi'}_n)_{n \in
%   \nats}$ that translates $\psi'(\vec{y})$. Let $T_n(\psi (\vec{y})) :=
%   \op{compose}(C^{\psi'}_n, ( T_n(\phi_1(\vec{y})), T_n(\phi_2(\vec{y}))),
%   (R_{\phi_1}, R_{\phi_2}))$. We note that $T_n(\psi(\vec{y}))$ translates
%   $\psi (\vec{y})$ at $n$ if $T_n(\phi_1 (\vec{y}))$ and
%   $T_n(\phi_2(\vec{y}))$ translate $\phi_1(\vec{y})$ and $\phi_2(\vec{y})$ at
%   $n$, respectively.

%   The approach used in the conjunctive case can be summarised as follows.
%   Given a sub-formula $\psi (\vec{y})$, we construct an auxiliary formula
%   $\psi'(\vec{y})$ in which all the sub-formulas of $\psi(\vec{y})$ are
%   replaced with relation symbols. We then construct a circuit $C^{\psi'}$ that
%   translates $\psi'(\vec{y})$ at $n$ using the translation defined by Anderson
%   and Dawar. Finally, we recursively construct circuits that translate the
%   sub-formulas of $\psi(\vec{y})$, and compose these circuits with
%   $C^{\psi'}$.

%   This approach can be used in order to handle every case with the exception
%   of rank. In other words, we can use a similar approach to define
%   $T_n(\psi(\vec{y}))$ where $\psi(\vec{y})$ is a sub-formula of
%   $\theta_n(\vec{x})$ of the form $\phi_1(\vec{y}) \lor \phi_2(\vec{y})$,
%   $\neg \phi(\vec{y})$, $\forall z. \phi (\vec{y}, z)$, or $\exists z. \phi
%   (\vec{y}, z)$.

%   We now handle the rank case. The approach is quite similar, except in this
%   case we cannot use the translation defined by Anderson and Dawar in order to
%   translate the auxiliary formula, and so must define this translation
%   explicitly. Moreover, rather than just composing the circuits, we also add a
%   small gadget in order to ensure the resultant circuit is transparent.

%   Suppose $\psi(\vec{z}, \vec{w}, \vec{y})$ is a sub-formula of
%   $\theta_n(\vec{x})$ such that $\psi (\vec{y}) = \rank^{\leq r}_p \vec{z},
%   \vec{w} .\phi(\vec{z}, \vec{w}, \vec{y})$, for some $p, r \in \nats$ and $p$
%   prime. Let $k:= \vert \vec{y} \vert$, $k_1 : = \vert \vec{z} \vert$ and $k_2
%   := \vert \vec{w} \vert$. Let $R_\phi$ be a $(k_1 + k_2 + k)$-ary relation
%   symbol not in $\tau$. Let $\psi'(\vec{y}) := \rank^{\leq r}_p\vec{z},
%   \vec{w}. R_{\phi} (\vec{z}, \vec{w}, \vec{y})$. Let $G_{\text{out}} :=
%   \{g_{\vec{c}} : \vec{c} \in [n]^k\}$ and $G_{\text{in}} : = \{ g_{\vec{a},
%   \vec{b}, \vec{c}} : \vec{a} \in [n]^{k_1}, \vec{b} \in [n]^{k_2}, \vec{c}
%   \in [n]^{k}\}$. Let $C^{\psi'}_n:= \langle G^{\psi'}, \Omega^{\psi'},
%   \Sigma^{\psi'}, \Lambda^{\psi'}, L^{\psi'}\rangle$ be a
%   $(\mathbb{B}_{\rank}, \{ R_{\phi} \})$-circuit such that
%   \begin{itemizens}
%   \item $G^{\psi'} = G_{\text{out}} \cup G_{\text{in}}$,
%   \item $\Omega^{\psi'}$ be an injection from $[n]^k$ to $G^{\psi'}$ defined
%     by $\Omega^{\psi'}(\vec{c}) = g_{\vec{c}}$ for all $\vec{c} \in [n]^k$,
%   \item $\Sigma^{\psi'} (g) = \rank^r_p$ if $g \in G_{\text{out}}$ and
%     $\Sigma^{\psi'}(g) = R_\phi$ if $g \in G_{\text{in}}$,
%   \item for each $(\vec{a}, \vec{b}, \vec{c}) \in [n]^{k_1 +k_2 + k}$,
%     $\Lambda^{\psi'} (\vec{a}, \vec{b}, \vec{c}) = g_{\vec{a}, \vec{b},
%     \vec{c}}$, and
%   \item for each gate $g_{\vec{c}} \in G_{\text{out}}$, $L(g_{\vec{c}}):
%     [n]^{k_1} \times [n]^{k_2} \rightarrow G^{\psi'}$ is defined by
%     $L(g_{\vec{c}})(\vec{a}, \vec{b}) := g_{\vec{a}, \vec{b}, \vec{c}}$ for
%     each $\vec{a} \in [n]^{k_1}$ and $\vec{b} \in [n]^{k_2}$.
%   \end{itemizens}

%   We note that $C^{\psi'}_n$ translates $\psi (\vec{y})$ at $n$. Let $R$ be a
%   relation symbol in $\tau$. Let
%   \begin{align*} T_n(\psi(\vec{y})) := \op{compose-and-tag}(C^{\psi'}_n,
%     (T_n(\phi(\vec{z}, \vec{w}, \vec{y}))), (R_\phi), R).
%   \end{align*}

%   We recall that $\op{compose-and-tag}$ combines circuits in a similar way to
%   $\op{compose}$ except that it also adds a set of gadgets to the circuit
%   ensuring that no two child gates of an output gate are
%   syntactically-equivalent. We thus have that $T_n(\psi (\vec{y}))$ is
%   transparent if $T_n(\phi (\vec{z}, \vec{w},\vec{y}))$ is transparent, and
%   $T_n(\psi(\vec{y}))$ translates $\psi (\vec{y})$ at $n$ if $T_n(\phi
%   (\vec{z}, \vec{w},\vec{y}))$ translates $\phi(\vec{z}, \vec{w}, \vec{y})$ at
%   $n$.

%   We have completed our definition of $T_n$ and shown, by induction on the
%   structure of $\theta_n(\vec{x})$, that for each $\psi (\vec{y})$ a
%   sub-formula of $\theta_n(\vec{x})$, $T_n(\psi(\vec{y}))$ translates
%   $\psi(\vec{y})$ at $n$.

%   The above definition can be used to define a recursive algorithm computing
%   the function $n \mapsto T_n (\theta_n(\vec{x}))$. We note that, since
%   $\width(\theta_n) \leq 4 t$ for all $n \in \nats$, there are only a finite
%   number of cases that need to be considered at each recursive step in this
%   algorithm. This fact can be used to show that there exists a polynomial
%   bound on the running time of the algorithm that depends only on $n$ and the
%   structure of $\theta(\vec{x})$. As such, we have that $(T_n (\theta_n))_{n
%   \in \nats}$ translates $\theta(\vec{x})$, and the result follows.
% \end{proof}

\bibliography{../references.bib}
% \ifcsdef{mainfile}{}{\bibliography{../references.bib}}
\end{document}