### 2018-02-28 sections/transparency-and-unique-extensions.tex

\documentclass[../paper.tex]{subfiles}
\begin{document}

In the statement of the main theorem of this paper, rather than working with
general symmetric rank-circuits, we instead restrict ourselves to families of
\emph{transparent} circuits. In this section we justify this restriction,
showing first that the a great number of very useful circuit properties (e.g.
syntactic-equivalence) are polynomial-time decidable for transparent circuits
and, second, that deciding most of these circuit properties for general circuits
(with non-symmetric gates) is at least as hard as the graph-isomorphism
problem. Since the proof of the main theorem makes use of the polynomial-time
decidability of many of these properties, these results together justify the
inclusion of the transparency restriction.

\subsection{The Usefulness of Transparency}
In the this subsection we show that a number of important circuit properties are
polynomial-time decidable for transparent circuits (e.g. symmetry, the
syntactic-equivalence relation, etc.). We also provide polynomial-time
algorithms for computing the action of a permutation on a symmetric circuit, as
well as for computing the orbits and supports of gates, and elements of
universes of gates, in a circuit.

We use the polynomial-time decidability of many of these properties throughout
this paper, and particularly for our translation of circuits into formulas. For
example, the polynomial-time decidability of the syntactic-equivalence relation
enables us to transform transparent circuits into circuits with unique labels
(and so unique extensions) in polynomial-time. This result allows us, when
working with $\PT$-uniform families of transparent circuits, to assume without a
loss of generality that these circuits have unique labels and unique extensions.
This is a crucial step in the proof of our main result, as the unique labels
requirement is needed to apply the support theorem, and the unique extensions
property is necessary for applying (and even defining) a great many technical
tools (e.g. supports and supporting partitions) used extensively throughout this
paper.

We now show that the syntactic-equivalence relation can be computed in
polynomial time for transparent circuits.

\begin{lem}
  There is an algorithm that takes as input a transparent circuit $C$ and
  outputs the syntactic-equivalence relation on the gates of $C$. The algorithm
  runs in time polynomial in the size of $C$.
  \label{lem:transparent-syntactic-equiv}
\end{lem}
\begin{proof}
  Let $C := \langle G, \Omega, \Sigma, \Lambda, L \rangle$ be a transparent
  $(\mathbb{B}, \rho)$-circuit of order $n$. We define the syntactic-equivalence
  relation on $G$ by induction on the depth of a gate We note that this
  definition by induction can be implemented as a construction using a dynamic program.

  Suppose $g$ and $h$ are gates in $C$ such that $\Sigma (g) = \Sigma(h)$ and
  either both $g$ and $h$ are output gates and $\Omega^{-1}(g) = \Omega^{-1}(h)$
  or neither are output gates. Let $g$ and $h$ be input gates, then set $g
  \equiv h$ if, and only if, both $g$ and $h$ are constant gates or both are
  relational gates and $\Lambda(g) = \Lambda (h)$. Let $g$ and $h$ be internal
  gates and suppose we have defined the syntactic-equivalence relation for all
  gates of depth less than the depth of either $g$ or $h$. Then $g \equiv h$
  if, and only if, $L(g)$ and $L(h)$ are isomorphism equivalent up to $\equiv$.

  The above algorithm can be implemented by a dynamic program with the number of
  iterations bounded by the number of pairs of gates in $C$, i.e. $\vert C
  \vert$ choose $2$. It remains to show that each iterative step can be
  implemented so as to run in time polynomial in the size of $C$. The only
  condition that is not obviously polynomial-time decidable is checking if
  $L(g)$ is isomorphism-equivalent to $L(h)$ up to $\equiv$. We sketch an
  algorithm that decides this condition in time polynomial in the size of the
  circuit.

  Let $g$ and $h$ be two internal gates in $C$ and suppose we have defined the
  syntactic-equivalence relation for all gates of depth less than the depth of
  either $g$ or $h$. We test if for each gate in $g' \in H_g$, there exists
  $h_{g'} \in H_h$ such that $g' \equiv h_{g'}$ and $\vert \{g'' \in H_g : g''
  \equiv g'\} \vert = \vert \{h'' \in H_h : h'' \equiv h'\}$. If this test
  fails, we set $g \not\equiv h$.

  We then check if both $g$ and $h$ are symmetric gates or if both are
  non-symmetric gates. Suppose both $g$ and $h$ are symmetric gates. We have
  that $L(g)$ is isomorphism-equivalent to $L(h)$ up to $\equiv$ if, and only
  if, there exists a bijection $f : H_g \rightarrow H_h$ such that $f(g') \equiv
  g'$, for all $g' \in H_g$ if, and only if, for each gate in $g' \in H_g$,
  there exists $h' \in H_h$ such that $g' \equiv h'$ and $\vert \{g'' \in H_g :
  g'' \equiv g'\} \vert = \vert \{h'' \in H_h : h'' \equiv h'\}$. Since we have
  already checked this condition we set $g \equiv h$.

  Suppose $g$ and $h$ are non-symmetric gates. Then, since $C$ is transparent,
  both $g$ and $h$ have unique labels. We thus have that for each $g' \in H_g$
  there is exactly one $h_{g'} \in H_h$ such that $g' \equiv h_{g'}$, and so we
  can define the bijection $f : H_g \rightarrow H_h$ by $f(g') := h_{g'}$ for
  all $g' \in H_g$. It can be shown that $L(g)$ is isomorphism-equivalent to
  $L(h)$ up to $\equiv$ if, and only if, $f \circ L(g)$ is
  isomorphism-equivalent to $L(h)$ if, and only if, $L(h)^{-1} f L(g)$ acts on
  $\ind (g)$ like an element of $\aut(g)$. We have from
  Lemma~\ref{lem:aut-partial} that $\sigma := L(h)^{-1} f L(g)$ acts like an
  element of $\aut(g)$ if, and only if, for all $\vec{a}_R, \vec{b}_{R'} \in
  \ind(g)$, $\sigma$ acts like a partial automorphism on $(\vec{a}_R,
  \vec{b}_{R'})$. We thus conduct the test by iterating through every pair
  $(\vec{a}_R, \vec{b}_{R'}) \in \ind(g)$, and set $g \equiv h$ if, and only if,
  $\sigma$ acts like a partial automorphism on every such pair.
\end{proof}

It is sometimes useful to define one circuit from another circuit by
\emph{quotienting} a circuit by the syntactic equivalence relation. Intuitively,
a quotient circuit is defined by `merging' each syntactic-equivalence class into
a single gate and including a wire between two gates in the quotient circuit if,
and only if, there is a wire between an element of each equivalence class.

\begin{definition}
  Let $C := \langle G, \Omega, \Sigma, \Lambda, L \rangle$ be a $(\BB,
  \rho)$-circuit. A \emph{quotient of $C$} is a $(\BB, \rho)$-circuit $C_\equiv
  := \langle G_\equiv , \Omega_\equiv, \Sigma_\equiv , \Lambda_\equiv, L_\equiv
  \rangle$, where $G_\equiv = G /_\equiv$, $\Omega_\equiv = \Omega /_\equiv$,
  $\Sigma = \Sigma /_\equiv$, $(\Lambda_\equiv)_R = \Lambda_R /_\equiv$ for all
  $R \in \rho$, and for all $[g] \in G_\equiv$ there exists $g' \in [g]$ such
  that, $L_\equiv$ associates with $[g]$ a function $L_\equiv([g]) : \ind(g')
  \rightarrow G_\equiv$ where $L_\equiv([g]) = L(g')/_\equiv$.
\end{definition}

We now show that taking the quotient of a circuit preserves important
properties, including the function computed by the circuit, whether the circuit
is symmetric, and has unique unique labels. We also show that the quotient of a
circuit has trivial syntactic-equivalence classes, and hence unique extensions.

\begin{lem}
  Let $C := \langle G, \Omega, \Sigma, \Lambda, L\rangle$ be a $(\BB,
  \rho)$-circuit of order $n$ and $C_{\equiv} = \langle G_\equiv ,
  \Omega_\equiv, \Sigma_\equiv , \Lambda_\equiv, L_\equiv \rangle$ be a quotient
  of $C$. Then $C$ and $C_{\equiv}$ compute the same function and each
  syntacitic-equivalence class in $C_\equiv$ is a singleton. Moreover, if $C$ is
  symmetric then $C_{\equiv}$ is symmetric, and for all $g \in G$, $g$ has
  unique labels in $C$ if, and only if, $[g]$ has unique labels in $C_{\equiv}$.
  \label{lem:quotient-circuits-preserve}
\end{lem}
\begin{proof}
  We first prove that $C_\equiv$ and $C$ compute the same function. Let
  $\mathcal{A}$ be a $\rho$-structure of size $n$ and let $\gamma$ be a
  bijection from $U$ to $[n]$. We now show that for all $g \in G$, $C_\equiv
  [\gamma \mathcal{A}]([g]) = C[\gamma \mathcal{A}](g)$. We do this by induction
  on the depth of a gate. Suppose $g \in G$ has depth $0$. In this case $g$ is
  an input gate and the result follows trivially. Suppose $g$ is an internal
  gate and suppose for all $h$ of depth less than $g$ we have that
  $C_\equiv[\gamma \mathcal{A}]([h]) = C[\gamma \mathcal{A}](h)$. We have that
  there exists $g' \in [g]$ such that $L_\equiv ([g]) = L(g') /_\equiv$ and
  $\Sigma_\equiv([g]) = \Sigma (g) = \Sigma (g')$. We have from Lemma~\ref{} and
  the inductive hypothesis that there exists $\lambda \in \aut(g)$ such that
  $L_\equiv ([g]) = L(g') = L(g) \lambda$. It follows from the fact that
  $\Sigma(g)$ is a structured function that $C_\equiv[\gamma \mathcal{A}]([g]) =
  \Sigma_\equiv ([g])(L^{\gamma \mathcal{A}}_\equiv([g])) = \Sigma (g')
  (L^{\gamma \mathcal{A}}(g')) = \Sigma (g)(L^{\gamma \mathcal{A}}(g) \lambda) =
  \Sigma (g) (L^{\gamma \mathcal{A}}(g)) = C[\gamma \mathcal{A}](g)$.

  We now show that each syntactic-equivalence class in $C_\equiv$ is a
  singleton. Suppose $[g], [h] \in G_\equiv$ and suppose $[g] \equiv [h]$. If
  $[g]$ and $[h]$ are both input or output gates then $[g] = [h]$. Suppose $[g]$
  and $[h]$ are internal gates that are not output gates. Then $\Sigma (g) =
  \Sigma_\equiv ([g]) = \Sigma_\equiv([h]) = \Sigma(h)$. There exists $g', h'
  \in G$ such that $g' \equiv g$ and $h' \equiv h$, and $L_\equiv ([g]) = L(g')
  /_\equiv$ and $L_\equiv ([h]) = L(h') /_\equiv$. It follows from $[g] \equiv
  [h]$ that $L(g') /_\equiv$ is isomorphism-equivalent to $L(h') /_\equiv$. It
  follows that $g' \equiv h'$, and so $[g] = [h]$.

  We now show that if $C$ is symmetric then $C_\equiv$ is symmetric. Suppose $C$
  is symmetric. Let $\sigma \in \sym_n$. Then there exists $\pi \in \aut(C)$
  that extends $\sigma$. Let $\pi_\equiv = pi /_\equiv$. We now show that
  $\pi_\equiv$ is an automorphism of $C_\equiv$ extending $\sigma$. It is easy
  to see that $\pi_\equiv$ is a bijection from $G_\equiv$ to $G_\equiv$ that
  preserves syntactic-equivalence, and is thus is a well-defined function. Let
  $[g] \in G_\equiv$. We have that $\Sigma_\equiv (\pi_\equiv [g]) =
  \Sigma_\equiv ([\pi g]) = \Sigma (\pi g) = \Sigma (g) = \Sigma_\equiv ([g])$.
  It is easy to check the automorphism conditions for input gates. Suppose $[g]$
  is an internal gate. Let $g' \in [g]$ be such that $L_\equiv([g]) = L(g')
  /_\equiv$ and $h \in [g]$ be such that $L_\equiv(\pi_\equiv[g]) = L_\equiv
  ([\pi h]) = L(\pi h) /_\equiv$. We have that $\pi L(g')$ is
  isomorphism-equivalent to $L(\pi g')$, and it follows that $(\pi
  L(g'))/_\equiv$ is isomorphism-equivalent to $L(\pi g') /_\equiv$. We then
  have $(\pi L(g'))/_\equiv = \pi_\equiv (L(g') /_\equiv) = \pi_\equiv L_\equiv
  ([g])$ and, since $g' \equiv h$ and so $\pi g' \equiv \pi h$, we have that
  $L(\pi g')/_\equiv$ is isomorphism-equivalent to $L(\pi h) /_\equiv =
  L_\equiv(\pi_\equiv[g])$. It follows that $\pi_\equiv L_\equiv ([g])$ is
  isomorphism-equivalent to $L_\equiv(\pi_\equiv [g]) /_\equiv$. Suppose $[g]$
  is an output gate. Then for $\vec{a} \in \dom (\Omega)$, $\pi_\equiv
  \Omega_\equiv (\vec{a}) = [\pi \Omega (\vec{a})] = [\Omega (\sigma \vec{a})] =
  \Omega_\equiv (\sigma \vec{a})$.

  Let $g \in G$. Suppose $[g]$ has unique labels. There exists $h \in [g]$ such
  that $L_\equiv([g]) = L(h) /_\equiv$. Since $L_\equiv([g])$ is injective,
  $L(h) /_\equiv$ must be injective and so $L(h)$ must be injective and no two
  child gates of $L(h)$ must be syntactically-equivalent. It follows that $h$
  has unique labels. Since $h \equiv g$ and $h$ has unique labels, it follows
  that $g$ has unique labels. Suppose $g$ has unique labels. Let $h \in [g]$ be
  such that $L_\equiv([g]) = L(h) /_\equiv$. Since $h \equiv g$ and $g$ has
  unique labels $h$ has unique labels and so $L_\equiv([g])$ has injective
  labels. Since each equivilence class in $C_\equiv$ is a singlton it follows
  that $[g]$ has unique labels.

  % If $[g]$ is an input gate then $\pi_\equiv [g] = [\pi g]$, and

  % Let $\mathcal{A}$ be a $\rho$-structure of size $n$ and let $\gamma$ be a
  % bijection from $U$ to $[n]$. We have from Lemma~\ref{} that for all $g, g'
  % \in
  % G$, if $g \equiv g'$ then $C[\gamma \mathcal{A}](g) = C[\gamma
  % \mathcal{A}](g')$. Moreover, we should note that for each internal gate

  % We show that for all $g \in G$, $C_\equiv [\gamma \mathcal{A}]([g]) =
  % C[\gamma
  % \mathcal{A}](g)$ by induction on the structure of the circuit. Suppose $g
  % \in
  % G$ is an input gate then, since every input gate is part of a singleton
  % equivalence class, we have that $C /_\equiv [\gamma \mathcal{A}]([g]) =
  % C[\gamma \mathcal{A}](g)$. Suppose $g \in G$ is an internal gate. Let $h \in
  % [g]$ such that $L /_\equiv ([g]) = L(h)/_\equiv$. Then $(L/_\equiv)^{\gamma
  % \mathcal{A}}([g])(x) = C /_\equiv[\gamma \mathcal{A}](L /_\equiv ([g])(x)) =
  % C [\gamma \mathcal{A}](L(h)(x)) = L^{\gamma \mathcal{A}}(h)(x) \equiv
  % L^{\gamma \mathcal{A}}(g) \circ \lambda (x)$ for some $\lambda \in \aut(g)$.
  % But then, since $\Sigma (g)$ is a structured function, we have that
  % $\Sigma(g)((L^{\gamma \mathcal{A}}(g) \lambda)) = \Sigma(g)(\L^{\gamma
  % \mathcal{A}})$. It follows that $C /_\equiv [\gamma\mathcal{A}]([g]) =
  % \Sigma /_\equiv ((L/_\equiv)^{\gamma mathcal{A}}([g])) = \Sigma (L^{\gamma
  % \mathcal{A}}(g)\lambda) = \Sigma (L^{\gamma \mathcal{A}}(g)) = C[\gamma
  % \mathcal{A}](g)$. It follows that $C/_\equiv [\gamma \mathcal{A}]
  % (\Omega/_\equiv (\vec{a})) = C [\gamma \mathcal{A}](\Omega (\vec{a}))$ for
  % all
  % $a \in [n]^q$. It follows that $C_{\equiv}$ and $C$ compute the same
  % function.

  % Suppose $C$ is symmetric. Suppose $C$ is symmetric. We have from For $\sigma
  % \in \sym_n$ let $\pi \in \aut(C)$ be an extension of $\sigma$. It can be
  % shown
  % that $\pi$ is invariant under $\equiv$, and hence we have that $\pi
  % /_{\equiv}$ is a function. Since each gate that is either an input or output
  % gate is a member of a singleton equivalence class we only need to check the
  % symmetry condition for internal gates. Suppose $g$ is an internal gate then
  % $\pi /_{\equiv} L /_{\equiv}([g]) (x) = [\pi L(g) (x)] = [L(\pi g) (\lambda
  % x)] = L([\pi g]) (\lambda x) = L(\pi /_{\equiv} [g]) (\lambda x)$ for some
  % $\lambda \in \aut(g)$. It follows that $\pi /_{\equiv}$ is an automorphism
  % of
  % $C_{\equiv}$ extending $\sigma$, and so $C_{\equiv}$ is symmetric.

  % Let $[g], [g'] \in G/_{\equiv}$. Suppose $[g] \equiv [g']$. If $g$ is an
  % input
  % or output gate we have that $g = g'$. Suppose $g$ is an internal gate. Then
  % there exists $h \in [g]$ and $h' \in [g']$ such that $L /_\equiv ([g]) =
  % L(h)
  % /_\equiv$ and $L /_\equiv ([g']) = L(h') /_\equiv$. We have that
  % $L_{\equiv}([g])$ is isomorphism-equivilent to $L_{\equiv}$

  % Let $g \in G$. Since every gate in $G /_{\equiv}$ is part of a singleton
  % equivalence class.

  % Then for each gate $g,
  % g' \in G$ such that $g \equiv g'$ we have that $C[\gamma\mathcal{A}](g) =
  % C[\gamma \mathcal{A}](g')$. If $g$ is an input gate we have that
  % $C[\gamma\mathcal{A}](g) = C /_\eqiv [\gamma \mathcal{A}]([g])$. Suppose
  % $[g] \in G /_{\equiv}$ is an internal gate. Let $L^{\gamma}([g]) : \ind(g)
  % \rightarrow \{0,1\}$ be defined by $L^{\gamma}([g]) (x) = C[\gamma
  % \mathcal{A}]([g])$
\end{proof}

It is worth noting that if $C$ is a transparent circuit we can compute a
quotient of $C$ in time polynomial in the size of $C$. Since the circuit is
transparent we can compute the syntactic-equivalence classes of $C$ in
polynomial time (see Lemma~\ref{lem:transparent-syntactic-equiv}). Picking a
representative of each equivalence class and wiring up the circuit in the
obvious way then yields the desired quotient circuit.

We now show that transparent circuits can be transformed in polynomial time into
equivalent circuit with unique labels, and hence unique extensions (see
Lemma~\ref{}), and that this transformation preserves important properties of
the circuit.

\begin{lem}
  Let $C := \langle G, \Omega, \Sigma, \Lambda, L \rangle$ be a transparent
  $(\mathbb{B}, \rho)$-circuit. There is an algorithm that that takes in such a
  circuit and outputs a $(\mathbb{B} \cup \mathbb{B}_{\std}, \rho)$-circuit $C'$
  such that $C$ and $C'$ compute the same function, $C'$ has unique labels, and
  if $C$ is symmetric then $C'$ is symmetric. Moreover, this algorithm runs in
  time polynomial in the size of the input circuit.
  \label{lem:transparent-unique}
\end{lem}
\begin{proof}
  Let $C_0 := \langle G_0, \Omega_0, \Sigma_0, \Lambda_0, L_0 \rangle$ be the
  quotient of $C$. If $C_0$ does not contain the constant gates $g_0$ and $g_1$
  such that $\Sigma(g_0) = 0$ and $\Sigma (g_1) = 1$ we construct a new circuit
  from $C_0$ by adding these gates. We abuse notation and also call this new
  circuit $C_0$. The addition of these constant gates to the circuit does not
  alter the function computed by the circuit, nor does it effect the symmetry of
  the circuit or weather it has unique labels.

  The proof proceeds by first defining a circuit $C_1$ from $C_0$ and then
  defining $C'$ from $C_1$. We then show that each of these constructions
  preserves the relevant circuit properties and that $C'$ has unique labels.

  We will define $C_1$ from $C_0$ by adding a designated gate that always
  evaluates to $1$ and which is part of a singleton syntactic-equivalence class,
  and then adding a wire from that gate to each $\AND$-gate in the circuit. This
  process ensures that each $\AND$-gates in $C_1$ has a fan-in of at least two.
  The construction of $C'$ from $C_1$ involves adding in a number of
  single-input $\AND$-gates, and so the construction of $C_1$ will allow us to
  distinguish these new gates from those in $C_1$.

  We now define the circuit $C_1$ formally. Let $C_1 := \langle G_1, \Omega_1,
  \Sigma_1 , \Lambda_1, L_1 \rangle$ be defined as follows. Let $G_1 = G_0
  \uplus \{ g_\lor\}$. Let $\Omega_1 = \Omega_0$ and $\Lambda_1 = \Lambda_0$.
  Let $g \in G_1$. If $g = g_\lor$ then $\Sigma_1 (g) = \lor[2]$, $L_1(g)(1) =
  g_0$, and $L_1(g)(2) = g_1$. If $g \in G_0$, $\Sigma(g) = \lor[2]$ and $H_g =
  \{g_0, g_1\}$ then $\Sigma_1 (g) = \lor[3]$, $L_1(g)(1) = L_0(g)(1)$,
  $L_1(g)(2) = L_0(g)(2)$, and $L_1(g)(3) = g_\lor$. If $g \in G_0$ and $g =
  \land[k]$ for some $k \in \nats$ then $\Sigma_1(g) = \land[k+1]$ and
  $L_1(g)(i) = L_0(g)(i)$ for all $i \in [k]$ and $L_1(g)(k+1) = g_\lor$.
  Otherwise let $\Sigma_1(g) = \Sigma_0(g)$ and $L_1(g) = L_0(g)$.

  More informally, we define $C_1$ from $C_0$ by adding in an $\OR$-gate
  $g_\lor$ that always evaluates to one, and then adding a wire from that gate
  to all $\AND$-gates in the circuit, and also a wire from $g_\lor$ to another
  $\OR$-gate in order to ensure that $g_\lor$ is part of a singleton
  syntactic-equivalence class in $C_1$. The important point to note is that each
  $\AND$-gate in $C_1$ has fan-in at least two, a property we will use to
  construct $C'$. We now show that the construction of $C_1$ preserves the
  relevant circuit properties.

  % We have defined $C_1$ from $C_0$ by in a gate $g_\lor$ that always evaluates
  % to $1$ and then adding a wire from that gate to each $\AND$-gate in the
  % circuit. This process ensures that each $\AND$-gates in $C_1$ has a fan-in
  % of at least two. The construction of $C'$ from $C_1$ involves adding in a
  % number of single-input $\AND$-gates, and so the construction of $C_1$ will
  % allow us to distinguish these new gates from those in $C_1$.

  % We have constructed $C_1$ from $C_0$ by adding in a gate $g_\lor$ and then
  % adding a wire from $g_\lor$ to each $\AND$-gate in the circuit and each gate
  % two-input $\OR$-gate with the same children as $g_\lor$. Since $g_\lor$
  % always evaluates to $1$, adding as an input to each $\AND$-gate (or the
  % $\OR$-gate) has no effect on the evaluation of the circuit.

  It is easy to see that if $C$ has order $n$ then $C_0$ has order $n$ and so
  $C_1$ has order $n$. Let $\mathcal{A}$ be a $\rho$-structure of size $n$ and
  let $\gamma$ be a bijection from the universe of $\mathcal{A}$ to $[n]$. We
  have that $C_1[\gamma \mathcal{A}](g_\lor) = 1$. We constructed $C_1$ from
  $C_0$ by adding a single wire from $g_\lor$ to each $\AND$-gate and each
  two-input $\OR$-gate with only the two constant gates as children. Notice that
  if $g$ is a two-input $\OR$ gate in $C_0$ with the two constant gates as
  children, then since $g$ has $g_1$ as a child and $g$ is an $\OR$-gate
  $C_0[\gamma \mathcal{A}](g) = C_1[\gamma \mathcal{A}](g) = 1$. It can be shown
  by induction that if $g \in G_1 \setminus \{g_\lor\}$ then $C_0 [\gamma
  \mathcal{A}](g) = C_1[\gamma \mathcal{A}](g)$. Since $\Omega_1 = \Omega_0$, it
  follows that $C_0$ and $C_1$ compute the same function. We thus have from
  Lemma~\ref{lem:quotient-circuits-preserve} that $C$ and $C_1$ compute the same
  function.

  Suppose $C$ is symmetric. From Lemma~\ref{lem:quotient-circuits-preserve} it
  follows that $C_0$ is symmetric. Let $\sigma \in sym_n$ and let $\pi_0 \in
  \aut(C_0)$ be an extension of $\sigma$. We define a bijection $\pi_1 : G_1
  \rightarrow G_1$ such that $\pi_1 (g) = \pi_0(g)$ for all $g \in G_0$ and
  $\pi_1 (g_\lor) = g_\lor$. It is easy to see that $\pi_1$ is an automorphism
  of $C_1$ extending $\sigma$. It follows that $C_1$ is symmetric.

  We have from Lemma~\ref{lem:quotient-circuits-preserve} that $C_0$ has trivial
  syntactic-equivalence classes. Since $g_\lor$ is the only two input $\OR$-gate
  with exactly the two constant gates as children, it follows that $g_\lor$ is
  contained in a singleton syntactic-equivalence class. It can be shown by
  induction that if $g, g' \in G_0$ are syntactically equivalent in $C_1$ then
  they must be syntactically-equivalent in $C_0$. It follows that if $g$ and
  $g'$ are syntactically-equivalent in $C_1$ then $g = g'$. We thus have that
  $C_1$ has trivial syntactic-equivalence classes.

  Let $C' := \{G', \Omega', \Sigma' \Lambda' L'\}$ be defined as follows. For
  each $h \in G_1$ let $c^h_g := \vert L_1(g)^{-1}(h) \vert$ if $h \in H_g$ and
  $c^h_g := 0$ otherwise. For each $h \in G_1$ let $c^h$ be the maximum of
  $c^h_g$ for $g \in G_1$. For each $h \in G_1$ if $c^h > 1$ let $G_h :=
  \{g^{h}_{1}, ldots, g^h_{c_h-1}\}$ otherwise let $G_h := \emptyset$. Let
  $G_{\land} := \biguplus_{h \in G_1} G_h$ and $G' = G_{\land} \uplus G_1$. Let
  $\Omega' = \Omega_1$ and $\Lambda' = \Lambda_1$. For $g \in G'$ if $g \in G_1$
  let $\Sigma'(g) = \Sigma_1(g)$ and otherwise let $\Sigma'(g) = \land[1]$. For
  each $g \in G_1$ and $h \in H_g$ let $\{x^{h,g}_0, x^{h,g}_1 ,\ldots,
  x^{h,g}_{c^h_g - 1}\} := L_1(g)^{-1}(h)$ be a fixed ordering of the set
  $L_1(g)^{-1}(h)$. Let $g \in G'$. If $g \in G_1$ and $x \in \ind(g)$ then
  there exists $h \in H_g$ and $i \in \{0, \ldots , c^h_g -1\}$ such that $x =
  x^{h,g}_i$, and we let $L'(g)(x) = h$ if $i = 0$ and $L'(g)(x^{h,g}_i) =
  g^h_i$ otherwise. If $g \in G_{\land}$ then $g = g^h_i$ for some $h \in G_1$
  and $i \in [c^h_g]$, and we let $L'(g)(1) = h$ if $i = 1$ and $L'(g)(1) =
  g^h_{i-1}$ otherwise.

  The construction ensures that $C'$ has injective labels. Let $g \in G_1$. In
  order to avoid confusion we let $H_g$ be the set of children of $g$ in $C_1$
  and $H_g'$ be the set of children of $g$ in $C'$. If $g' \in G_1$ we let $g
  \equiv g'$ denote syntactic-equivalence in $C_1$ and $g \equiv' g'$ denote
  syntactic-equivalence in $C'$.

  \begin{claim}
    $C'$ has trivial syntactic-equivalence classes.
  \end{claim}
  \begin{proof}
    We prove this result by induction on depth. Let $g \in G'$. If $g$ has depth
    $0$ then $g$ is an input gate and so $[g] = \{g\}$. Let $g \in G'$ be an
    internal gate in $C'$ and suppose for each gate $h$ of depth less than $g$
    we have that $[h] = \{ h \}$. Let $g' \in G'$ and suppose $g \equiv' g'$. We
    now show that $g = g'$, and so prove the claim. Since $g \equiv' g'$ it
    follows that $\Sigma (g) = \Sigma(g')$ and both $g$ and $g'$ have the same
    depth. Moreover, from the inductive hypothesis, we have that there exists
    $\lambda \in \aut(g)$ such that for all $x \in \ind(g)$, $L'(g) (x) =
    L'(g')(x)$. It follows that $H_{g}' = H_{g'}'$. Moreover, from the inductive
    hypothesis the children of $g$ and $g'$ must be contained in singleton
    syntactic-equivalence classes, and so , since $C'$ has injective labels, it
    follows that $g$ and $g'$ have unique labels in $C'$.

    % Suppose $g = g_\lor$. Then $H_g = H_g' = \{g_0, g_1\} = H_{g'}'$ and
    % $\Sigma' (g) = \Sigma'(g') = \lor[2]$. But from the construction the only
    % two-input $\OR$-gate in $C_1$, and so $C'$, with children $\{g_0 , g_1\}$
    % is
    % $g$. It follows that $g = g'$.

    Suppose $g \in G_{\land}$. Then there exists $h \in G_1$ and $i \in [c^h -
    1]$ such that $g = g^h_i$. Suppose $g' \in G_1$. Then $\Sigma'(g) =
    \land[1]$. But, from the construction of $C_1$, there are no single-input
    $\AND$-gates in $G_1$. It follows $\Sigma'(g') \neq \Sigma'(g)$ and so $g'
    \not\equiv' g$, a contradiction, and so we must have $g' \in G_{\land}$. If
    $i = 1$ then $\{ h \} = H_g' = H_{g'}'$. Since the only gate in $G_\land$
    that has $h$ as a child gate is $g^h_1$, it follows that $g = g^h_1 = g'$.
    If $i > 1$ then $\{g^{h}_{i-1}\} = H_g' = H_{g'}'$. Since $g^h_i$ is the
    only gate in $G_\land$ that has $g^h_{i-1}$ as a child gate, we have $g =
    g^h_i = g'$. It follows that if $g \in G_\land$ then $g = g'$.

    Suppose $g \in G_1$. We have already shown that if $g' \in G_\land G_1$ then
    $g \not\equiv' g'$, a contradiction, and so we must have $g' \in G_1$. Since
    $g \equiv' g'$ there exists $\lambda \in \aut(g)$ such that for all $x \in
    \ind(g)$, $L'(g)(\lambda x) = L'(g')(x)$. From the construction, we have
    that for all $x \in \ind(g)$ and $h \in G_1$, $L'(g)(x) \in \{h, g^h_1 ,
    \ldots, g^h_{c^h_g -1}\}$ if, and only if, $L_1(g)(x) = h$. Suppose $g$ is
    not a symmetric gate. Since $C$ is transparent, $C_0$ is transparent and so
    $C_1$ is transparent. It follows that for all $h \in H_g$, $H_g = H_{g'}$,
    and $c^h_g = 1 = c^h_{g'}$. It follows that for all $x \in \ind(g)$, there
    exists a unique $h \in G_1$ such that $L_1(g)(\lambda x) = L'(g)(\lambda x)
    = L'(g')(x) = L_1 (g')(x) = h$. It follows that $g \equiv g'$ and so $g =
    g'$. Suppose $g$ is a symmetric gate. Let $x \ind(g)$ and $h = L_1(g)(x)$.
    If $L'(g)(\lambda x) = L'(g')(x)= h$ then $L_1(g)(\lambda x) = h$ if, and
    only if, $L'(g)(\lambda x ) = h* = L'(g')(x)$, where $h* = h$ or $h* =
    g^{h}_i$, for some $i \in [c^h_g -1]$, if, and only if, $L_1(g) (x) = h$. It
    follows that for all $x \in \ind(g)$, $L_1 (g) (\lambda x) = L_1 (g)(x)$,
    and so $g \equiv g'$ and so $g = g'$.

    % It follows that for all $g \in G'$, $[g] = \{g\}$. The result follows.

    % Suppose $g \in G_0$. If $g$ is a non-symmetric gate then $g$ has unique
    % labels in $C_0$ and so has unique labels in $C_1$. It follows that for all
    % $h \in H_g$, $c^h_g = 1$, $g$ has injective labels in $C'$ and so, using
    % the
    % inductive hypothesis, $g$ has unique labels. Since $g \equiv' g'$ we also
    % have that $g'$ has unique labels.

    % It is easy to show that for all $h, g \in G_1$, if $W_t(h,g)$ in $C_1$
    % then
    % $W_t(h,g)$ in $C'$. We prove this result by induction on depth. Suppose
    % $g$
    % and $g'$ have depth $0$ in $C'$ then $g$ and $g'$ are input gates, and so
    % $g
    % \equiv' g'$ if, and only if, $g = g'$ if, and only if, $g \equiv g'$.
    % Suppose $g, g' \in G_1$ are internal symmetric gates and for all $h , h'
    % \in
    % G_1$ of depth less than $g$ or $g'$ in $C'$ $h \equiv h'$ if, and only if,
    % $h \equiv' h'$. Suppose $g \equiv g'$. Then there exists $\lambda \in
    % \aut(g)$ such that for all $x \in \ind(g)$, $L_1(g)(x) \equiv
    % L_1(g')(\lambda x)$. If $H_g'\cup \H_{g'}' \subseteq G_1$ then the result
    % follows obviously from the inductive hypothesis. For each $h \in H_g$ and
    % $x
    % \in L(g)^{-1}(h)$ if $c^h_g = 1$ then let $\lambda' x = x$ and if $c^h_g >
    % 1$ and if $x = x^{h,g}_i$ for some $i \in [c^h_g - 1]$ then $\lambda' x =
    % x^{h', g'}_i$ where $h' = L_1(g')(\lambda x)$. It can be shown that
    % $\lambda'$ is a bijection on $\ind(g)$. Since $g$ and $g'$ are symmetric
    % gates it follows that $\lambda' \in \aut(g)$...
  \end{proof}

  Since $C'$ has injective labels and trivial syntactic-equivalence classes it
  follows $C'$ has unique labels.

  We have already shown that if $C$ is symmetric then $C_1$ is symmetric. We now
  show that if $C_1$ is symmetric then $C'$ is symmetric. Suppose $C_1$ is
  symmetric. Let $\sigma \in \sym_n$ and let $\pi_1 \in \aut(C_1)$ be an
  extension of $\sigma$. We define the function $\pi' : G' \rightarrow G'$ as
  follows. Let $g \in G'$. If $g \in G_1$ then $\pi'(g) = \pi_1(g)$, otherwise
  $g \in G_\land$, and so there exists $h \in G_1$ and $i \in [c^h - 1]$ such
  that $g = g^h_i$, let $\pi'(g) = g^{\pi_1 h}_i$. It is easy to show that
  $\pi'$ is an automorphism of $C'$ extending $\sigma$.

  \begin{claim}
    Let $g \in G_1$, $\mathcal{A}$ be a $\rho$-structure of size $n$, and
    $\gamma$ be a bijection from the universe of $\mathcal{A}$ to $[n]$. Then
    $C'[\gamma \mathcal{A}](g) = C_1[\gamma \mathcal{A}](g)$.
  \end{claim}
  \begin{proof}
    It is easy to see that for all $h \in G_1$ and $i \in [c^h-1]$ we have that
    $C'[\gamma \mathcal{A}](g^h_i) = C'[\gamma \mathcal{A}](h)$. Suppose $g \in
    G_1$. We now prove the result by induction on the depth of a gate. Suppose
    $g$ has depth $0$, then $g$ is an input gate, and the case follows
    trivially. Suppose $g$ is an internal gate, and for all $h$ of depth less
    than $g$ we have that if $h \in G_1$ then $C'[\gamma \mathcal{A}](h) =
    C_1[\gamma \mathcal{A}](h)$. Let $x \in \ind(g)$. From the construction of
    $C'$ we have that if $L'(g)(x) \in G_1$ then ${L'}^{\gamma
      \mathcal{A}}(g)(x) = L^{\gamma \mathcal{A}}_1(g)(x)$. If $L'(g)(x) \not\in
    G_1$ then $L'(g)(x) \in G_{\land}$ and so $L'(g)(x) = g^h_i$, where $h =
    L_1(g)(x)$ and some $i \in [c^h_g -1]$. But then ${L'}^{\gamma
      \mathcal{A}}(g)(x) = C'[\gamma \mathcal{A}](g^h_i) = C'[\gamma
    \mathcal{A}](h) = C_1[\gamma \mathcal{A}](h) = L^{\gamma
      \mathcal{A}}_1(g)(x)$. The penultimate equality follow from from the
    inductive hypothesis. The final equality follows from the fact that $h = L_1
    (g)(x)$. We thus have $C'[\gamma \mathcal{A}](g) = \Sigma'({L'}^{\gamma
      \mathcal{A}}(g)) = \Sigma_1 (L^{\gamma \mathcal{A}}_1(g)) = C_1[\gamma
    \mathcal{A}](g)$.
  \end{proof}

  Since $\Omega' = \Omega_1$, we have that $C'$ and $C_1$ compute the same
  function. It follows that, since $C_1$ and $C$ compute the same function, $C$
  and $C'$ compute the same function. Since $C$ is transparent, we may construct
  the quotient circuit $C_0$ in time polynomial in $\vert C \vert$. Since $C_1$
  is constructed by adding in a single gate and then adding at most $\vert C_0
  \vert$ wires, we may construct $C_1$ from $C_0$ in time polynomial in $\vert C
  \vert$. It is is easy to see that $C'$ can be constructed in time polynomial
  in $\vert C_1 \vert$ and hence polynomial in $\vert C \vert$. The result
  follows.
  %
  % It is easy to see that $C'$ has injective labels. Let $g$ be an internal
  % gate
  % in $C'$ and suppose each gate of depth less than $g$ has unique labels. Let
  % $h, h' \in H_g$ such that $h \equiv h'$. If $h, h' \in G_1$ it follows $h =
  % h'$. Suppose $h, h' \in G_{\land}$. Then there exists $d, d' \in G_1$ and
  % $i,
  % j$ such that $g^d_i \equiv g^{d'}_j$. If $j = i$ then $d \equiv d'$ and so
  % $d
  % = d'$ and $h = h'$. Suppose, without a loss of generality, that $i > j$. It
  % follows that $g^d_{i-j} \equiv d'$. But then $d'$ is a single-input
  % $\AND$-gate, which is impossible, as from the construction $d'$ must have at
  % least two inputs. Suppose

  % Thus $g \equiv g'$ if, and only if, $g = g'$
  % Let $C'':= \langle G'', \Omega'', \Sigma'', \Lambda'', L'' \rangle$ be a
  % quotient of $C$. We have from Lemma~\ref{lem:quotient-circuits-preserve}
  % that
  % $C''$ is transparent, i.e. all the non-symmetric gates have unique labels,
  % and
  % that no two input gates to any one gate are syntactically-equivalent. We now
  % construct a circuit $C'$ from $C''$ that leaves the non-symmetric gates
  % unchanged but such that each symmetric gate in $C'$ has unique labels. We
  % construct $C'$ by taking each symmetric gate $g$ in $C''$ and each gate $h
  % \in
  % H_g$ and adding in a sequence of single-input $\AND$-gates, one for each
  % wire
  % from $h$ to $g$, such that the first $\AND$-gate in this sequence has $h$ as
  % its input and each successive $\AND$-gate has the previous one as an input.
  % We
  % then remove all but one of the wires from $h$ to $g$ and add a wire from
  % each
  % $\AND$-gate in this sequence to $g$. We now construct $C'$ formally.

  % Let $C_i = \langle G_i, \Omega_i, \Sigma_i, \Lambda_i, L_i \rangle$ be a
  % transparent circuit of order $n$. We construct $C_{i+1} := \langle G_{i+1},
  % \Omega_{i+1}, \Sigma_{i+1}, \Lambda_{i+1}, L_{i+1} \rangle$ as follows. For
  % each $g \in G_i$ and $h \in H_g$ let $s^g_h = \{x \in \ind(g) : L(g)(x)
  % \equiv
  % h\}$ and let $c_h$ be the maximum $\vert c^g_h \vert$ for all $g \in G_i$.
  % Let
  % $G_{\land} := \{g^h_\land : h \in G_i, \, c_h > 1\}$. Let $G_{i}' := G_i
  % \cup
  % G_{\land}$. Let $\Omega_{i}' := \Omega_i$. For $g \in G_i$ let $\Sigma_{i}'
  % (g) := \Sigma_i (g)$ if $g \in G_i$ and $\Sigma_{i}'(g) := \land[1]$
  % otherwise. Let $\Lambda_{i}' := \Lambda_i$. We now define $L_{i}'(g)$ for
  % each
  % $g \in G_{i}'$. If $g \in G_{i}$ and $g$ is not a symmetric gate then
  % $L_{i}'(g) = L_i(g)$. If $g \in G_{\land}$ and $g = g^h_\land$ for some $h
  % \in
  % G_i$ let $L(g)(1) = h$. Suppose $g \in G_{i}$ and $g$ is a symmetric gate.
  % For
  % each $h \in H_g$ select one $x_h \in \ind(g)$. For $x \in \ind(g)$ let
  % $L_{i+1}(g) (x) = h$ if $x = x_h$ and $L_{i}'(g)(x) = g^h_\land$ otherwise.
  % Let $F(C_i) := C_{i+1} := \langle G_{i+1}, \Omega_{i+1}, \Sigma_{i+1},
  % \Lambda_{i+1}, L_{i+1} \rangle$ be a quotient of $C_i'$.

  % Let $C_0' := C$ and $C_1$ be a quotient of $C_0$. For each natural number
  % $i>
  % 0$ we define a circuit $C_{i+1}$ from $C_{i} := \langle G_{i}, \Omega_{i},
  % \Sigma_{i}, \Lambda_{i}, L_{i} \rangle$ as follows. We first define $C_i' =
  % \langle G_i', \Omega_i', \Sigma_i', \Lambda_i', L_i' \rangle$ as follows.
  % For
  % each $g, h \in G_i$ let $c^g_h = \vert L_i(g)^{-1}(h) \vert$ and let $c_h$
  % be
  % the maximum $c^g_h$ for all $g \in G_i$. Let $G_{\land} := \{g^h_\land : h
  % \in
  % G_i, \, c_h > 1\}$. Let $G_{i}' := G_i \cup G_{\land}$. Let $\Omega_{i}' :=
  % \Omega_i$. For $g \in G_i$ let $\Sigma_{i}' (g) := \Sigma_i (g)$ if $g \in
  % G_i$ and $\Sigma_{i}'(g) := \land[1]$ otherwise. Let $\Lambda_{i}' :=
  % \Lambda_i$. We now define $L_{i}'(g)$ for each $g \in G_{i}'$. If $g \in
  % G_{i}$ and $g$ is not a symmetric gate then $L_{i}'(g) = L_i(g)$. If $g \in
  % G_{\land}$ and $g = g^h_\land$ for some $h \in G_i$ let $L(g)(1) = h$.
  % Suppose
  % $g \in G_{i}$ and $g$ is a symmetric gate. For each $h \in H_g$ select one
  % $x_h \in \ind(g)$. For $x \in \ind(g)$ let $L_{i}'(g) (x) = h$ if $x = x_h$
  % and $L_{i}'(g)(x) = g^h_\land$ otherwise. Let $C_{i+1} = F(C_i) := C_{i+1}
  % :=
  % \langle G_{i+1}, \Omega_{i+1}, \Sigma_{i+1}, \Lambda_{i+1}, L_{i+1} \rangle$
  % be a quotient of $C_i'$.

  % Let $i \in \nats$. For $g \in G_i$, in order to avoid ambiguity we let $H_g$
  % denote the set of children of $g$ in $C_i$ and $H_g'$ denote the set of
  % children of $g$ in $C_i'$. We let $\equiv$ be the syntactic-equivalence
  % relation in $C_i$ and $\equiv'$ be the syntactic-equivalence relation in
  % $C_i'$.

  % If $C_i$ has order $n$ then $C_i'$ is has order $n$ and so $C_{i+1}$ has
  % order
  % $n$. We first show that $C_{i+1}$ and $C_i$ compute the same function. Let
  % $\mathcal{A}$ be a $\rho$-structure of size $n$ and let $\gamma$ be a
  % bijection
  % from the universe of $\mathcal{A}$ to $[n]$. The construction of $C_i'$ from
  % $C_i$ by identifying those gates $h, g \in G_i$ such that $h \in H_g$ and
  % $\vert L_i(g)^{-1}(h) \vert > 1$ and then replacing all but one of the wires
  % from $h$ to $g$ with a wire from $g^h_\land$ to $g$. Since $g^h_\land$ is a
  % single-input $\AND$-gate with input $h$, we have that $C_i[\gamma
  % \mathcal{A}](h) = C_{i}'[\gamma \mathcal{A}](h) = C_i'[\gamma
  % \mathcal{A}](g^h_\land)$. It follows that $C_i [\gamma \mathcal{A}](g) =
  % C_i'
  % [\gamma \mathcal{A}](g)$. We thus have that $C_i$ and $C_i'$ compute the
  % same
  % function and so, using Lemma~\ref{lem:quotient-circuits-preserve}, $C_{i+1}$
  % computes the same function as $C_i$.

  % Suppose $C_i$ is symmetric. Let $\sigma \in \sym_n$ and $\pi \in \aut(C)$ be
  % an extension of $\sigma$. It can be shown that $c_h = c_{\pi h}$. As such,
  % we
  % can extend $\pi$ to $G_i'$, defining $\pi' : G_i' \rightarrow G_i'$ by $\pi'
  % g
  % = \pi g$ if $g \in G_i$ and letting $\pi' g^h_\land = g^{\pi h}_\land$ if
  % $g^h_\land \in G_{\land}$. It can be shown that $\pi'$ is an automorphism of
  % $G_i'$ extending $\sigma$, and so $C_i'$ is symmetric. It follows from
  % Lemma~\ref{lem:quotient-circuits-preserve} that $C_{i+1}$ is symmetric.

  % It can be shown that for all $g, g' \in G_i$, $g \equiv g'$ if, and only if,
  % $g \equiv' g'$.

  % % It follows that for $g^h_\land, g^{h'}_\land \in G_\land$ if
  % % $g^h_\land \equiv' g^{h'}_\land$ then $h \equiv' h'$ and so $h \equiv h'$.

  % Suppose $C_i$ has unique labels. It follows that $c_h = 1$ for all $h \in
  % G_i$
  % and so $C_i = C_{i}'$. We thus have, using
  % Lemma~\ref{lem:quotient-circuits-preserve}, that $C_{i+1}$ has unique
  % labels.
  % Suppose $C_i$ does not have unique labels. For $g \in G_{i}$ and $h \in H_g$
  % let $s^h_g := \vert \{x \in \ind(g): L_i(g)(x) \equiv h\} \vert$. In other
  % words $s^h_g$ is the number of wires to $g$ from either $h$ or a gate
  % syntactically-equivalent to $h$. For $g \in G_i$ let $d_g$ be the minimum
  % depth of a gate $h \in H_g$ in $C_i$ such that $s^h_g > 1$ and let $d_g
  % :=\infty$ if no such gate exists. In other words $d_g$ is the depth of the
  % minimum-depth gate witnessing that $g$ does not have unique labels. For $g
  % \in
  % G_i'$ we define $d_g'$ similarly for $C_i'$. For $g \in G_i$ let $S^i_g :=
  % \{h
  % \in H_g : \depth(C_i, h) < d_g \}$ be a set of gates in $C_i$. For $g \in
  % G_i'$ we define $S^i_g'$ similarly for $C_i'$.

  % Since $C_i$ does not have unique labels, there exists $g \in G_i$ such that
  % $d^i_g \neq \infty$. Let $H^d_g := \{h \in H_g : \depth (C_i, h) = d_g\}$ be
  % a
  % set of gates in $C_i$. We now prove that $H^d_g \cup S^i_g \subseteq
  % S^i_g'$.
  % It can be shown that for all $h, h' \in H_g$, if $\depth(C_i, h) <
  % \depth(C_i,
  % h')$ then $\depth(C_i', h) < \depth(C_i', h')$. Since the construction
  % preserves the `child of' relation, we have for all $h \in H_g$ that
  % $\depth(C_i, h) \leq \depth(C_i', h)$. Let $h \in S^i_g$. Since $s^h_g = c_h
  % =
  % 1$ we have that $\vert L_{i}'(g)^{-1}(g) \vert = 1$. From
  % Lemma~\ref{lem:quotient-circuits-preserve} we have that $h \not\equiv h'$
  % for
  % all $h' \in H_g$. So suppose $h \equiv' g^{h'}_{\land}$ for some $g^h'_\land
  % \in G_\land$. It follows that $h'$ must be syntactically-equivalent to the
  % single child of $h$ in $C_i$ and so $\depth(C_i',h') < \depth(C_i', h)$

  % It follows that $h' \in H_{h}'$ (since $C_i$ is a quotient) Suppose there
  % exists a gate $h' \in G_i'$ such that $h \equiv h'$ in $C_i'$. Since $C_i$
  % has
  % trivial symmetric-equivalence classes, it follows that $h' = g^{h''}_\land$
  % for

  % Let $d$ be the maximum depth in $C_i'$ of a gate in $H^d_g \cup S^i_g$.
  % Suppose $h \in G_i'$ such that $\depth (C_i', h) \leq d$ and $s^h_g' > 1$.

  % We have from the construction that $\vert L_i'(g)^{-1}(h) \vert = 1$. It
  % follows that there exists $h' \in H_g'$, $h' \neq h$ such that $h' \equiv' h
  % $. Moreover, since $G_i$ has singleton-equivalence classes we have that $h'
  % \not\in G_i$. So there exists $h'' \in H_g$ such that $c^g_{h''} > 1$ and
  % $h'
  % = g^{h''}_i \equiv' h'$. Since $C_i$ is a quotient we have that $h''$ is the
  % only child of $h'$ in $C_i$, and so $\depth(C_i, h'') < \depth(C_i, h')$.
  % Moreover, since $\depth(C_i, h') \leq \depth(C_i', h') \leq d \leq d_g$ We
  % thus have that $d_g \leq \depth(C_i, h'') < \depth(C_i, h') \leq \depth(C_i,
  % h) \leq d_g$

  % If $d < d_g'$ we are done. Suppose $d \geq d_g'$. Then there exists $h \in
  % H^d_g \cup S^i_g$ such that $d_g' \leq d(C_i', h) \leq d$

  % and suppose there exists $h \in H_g'$ such that $\depth(C_i, h) < d$

  % Since $C_i$ does not have unique labels there exists $g \in G_i$ and $h \in
  % H_g$ such that $d(C_i, g) \neq \infty$. If $h \in s(C_i, g)$ and $c_h = 1$
  % then $s(C_i', g)$.

  % Let $H^{d}_g \subseteq H_g$ be the set of children of $g$ of depth
  % $d(C_i,g)$.

  % We have for each $h \in G_i$ such that $c_h > 1$ that $g^h_i$ has a depth in
  % $C_i'$ strictly greater than $h$ ensures that $d_g \leq d_g'$

  % For $g \in G_{i}$ and $h \in H_g$ let $s^h_g := \vert \{x \in \ind(g):
  % L_i(g)(x) \equiv h\} \vert$. Let $d_g$ be the minimum depth of a gate $h \in
  % H_g$ in $C_i$ such that $s^h_g > 1$ and let $d_g :=\infty$ if no such gate
  % exists. It follows that $d_g = \infty$ for all $g \in G_i$ if, and only if,
  % $C_i$ has unique labels.

  % Suppose $C_i$ does not have unique labels. Let $g \in G_i$ such that $d_g
  % \neq
  % \infy$ and let $H^{d}_g \subseteq H_g$ be the set of children of $g$ of
  % depth
  % $d_g$. We notice that for all $h, h' \in H^d_g$, $h'$ and $g^h_\land$ have
  % different depths in $C_i'$ and so $h' \no\equiv g^h_\land$ in $C_i'$.
  % Moreover,

  % For $g \in G_i$ let $s_g := s^h_g$ where $h$ is the minmimal-depth where $h
  % \in H_g$ is the

  % be the minimum depth the is minimum depth child such that $\vert H_g
  % /_\equiv
  % \vert > 1$, and $s_g = 0$ if no such gate exists. Let $S(C_i)$ be the
  % maximum
  % of $s_0$.

  % Notice that $s_i = 0$ if, and only if, $C_i$ has unique labels. Suppose
  % $S(C_i) \neq 0$. Let $H^s_g = \{ h \in H_g : c^h_g > 1\}i$. Let $h$ be the
  % gate of minimal depth in $H^s_g$. Then $h$ is not syntactically-equivalent
  % to
  % $g^h_\land$

  % Suppose $C_i$ does not have unique labels. Then

  % Suppose $C_i$ is symmetric.

  % % Let $H \subseteq G''$ be the set of all symmetric gates. For each $h \in
  % % G''$
  % % let $c_h$ be the maximum of $\vert L(g)^{-1}(h) \vert$ for $g \in H$ such
  % % that
  % % $h \in H_g$. For each $h \in G''$ let $G_{h, \land} := \{g^h_1, \ldots,
  % % g^h_{c_h - 1}\}$. Let $G_\land := \bigcup_{h \in G''} G_{h, \land}$. Let
  % % $G'
  % % := G'' \cup G_{\land}$. Let $\Omega' := \Omega''$. For $g \in G''$ let
  % % $\Sigma' (g) := \Sigma'' (g)$ if $g \in G$ and $\Sigma'(g) := \land[1]$
  % % otherwise. Let $\Lambda' := \Lambda''$. We now define $L'(g)$ for each $g
  % % \in
  % % G'$. If $g \in G' \setminus (H \cup G_\land)$ let $L'(g) = L''(g)$. If $g
  % % \in
  % % G_\land$ and $g = g^h_i$ for some $h \in G''$ and $i \in [c_h]$, if $i =
  % % 1$
  % % let $L(g)(1) = h$ otherwise let $L(g)(1) = g^h_{i-1}$. Suppose $g \in H$.
  % % For
  % % each $h \in H_g$ let $\{x^h, x^h_1, \ldots, x^h_k\} := L(g)^{-1}(h)$, and
  % % notice that $\vert L(g)^{-1}(h) \vert \leq c_h$. For each $x \in \ind(g)$
  % % let
  % % $h = L(g)(x)$ and let $L'(g) (x) = h$ if $x = x^h$ and $L'(g)(x) = g^h_i$
  % % if
  % % $x = x^h_i$ for some $i \in [\vert L(g)^{-1} \vert - 1]$.

  % e have from Lemma~\ref{lem:quotient-circuits-preserve} that $C''$ has unique
  % labels if $C$ has unique labels, $C''$ is symmetric if $C$ is symmetric and
  % $C''$ and $C$ compute the same function. Since $C''$ is constructed from
  % $C'$
  % by replacing some of the wires from $h$ to the symmetric gate $g$ with
  % $\AND$-gates that evaluate to the same value as $h$, it is not hard to see
  % that $C'$ computes the same function as $C''$, and hence $C'$. Moreover, the
  % construction also ensures that each symmetric gate (and so each internal
  % gate)
  % in $C'$ has injective labels. It is easy to see that for a fixed $h \in G''$
  % no two gates in $G_{h,\land} \uplus \{h\}$ are syntactically-equivalent.
  % Suppose $h, h' \in G''$ and $g^h_i \equiv g^{h'}_j$ for some $i \in [c_h -
  % 1]
  % $ and $j \in [c_{h'} -1]$. If $j = i$ then it can be shown that $h \equiv
  % h'$
  % and so $h = h'$. If $i > j$ then it can be shown that $g^h_{i-j} \equiv h'$,
  % but then $W_T(h', h)$ in $C''$

  % Let $C/_{\equiv} := \langle G /_{\equiv}, \Omega/_{\equiv},
  % \Sigma/_{\equiv},
  % \Lambda/_{\equiv}, L/_{\equiv}$ be the quotient of $C$ by
  % syntactic-equivalence. Since each gate that is an input or output gates is a
  % member of a singleton equivalence class, we have that $C/_{\equiv}$ is
  % indeed
  % a circuit.

  % We have that no two gates in $G /_{\equiv}$ are syntactically-equivalent.
  % Suppose $C$ is symmetric. We have from For $\sigma \in \sym_n$ let $\pi \in
  % \aut(C)$ be an extension of $\sigma$. It can be shown that $\pi$ preserves
  % $\equiv$, and hence we can define $\pi /_{\equiv}$. We note that each gate
  % that is either an input or output gate is a member of a singleton
  % equivalence
  % class. Moreover, if $g$ is an internal gate then $\pi /_{\equiv} L
  % /_{\equiv}([g]) (x) = [\pi L(g) (x)] = [L(\pi g) (\lambda x)] = L([\pi g])
  % (\lambda x) = L(\pi /_{\equiv} [g]) (\lambda x)$ for some $\lambda \in
  % \aut(g)$. It follows that $\pi /_{\equiv}$ is an automorphism of
  % $C_{\equiv}$
  % extending $\sigma$, and so $C_{\equiv}$ is symmetric.

  % are each members of singleton equivalence classes $\pi /_{\equiv}$ is an
  % automorphism of $C /_{\equiv}$ extending We first define a circuit $C''$ by
  % quotienting $C$ by the syntactic-equivalence relation. Let $H_1, \ldots,
  % H_q$
  % be the syntactic-equivalence classes of $C$, and let $H_i = \{h^i_1, \ldots
  % h^i_{\vert H_i \vert}\}$ for all $i \in [q]$. We now define a circuit $C''$.
  % Let $G'' := G \setminus (\bigcup_{i \in [q]} (H_i \setminus \{h^i_1\}))$. We
  % notice that each input and output gate in $C$ is in a singleton
  % syntactic-equivalence class. Let $\Omega'' := \Omega$ and $\Lambda'' :=
  % \Lambda$. Let $\Sigma'' (g) = \restr{\Sigma}{G''}$. For all $g \in G''$ and
  % $x
  % \in \ind(g)$, if $L(g)(x) \in H_i$ for some $i \in [q]$ then let $L''(g) =
  % h^i_1$. Let $C'' := \langle G'', \Omega'', \Sigma'', \Lambda'', L''
  % \rangle$.
  % Notice that $C''$ is just the process

  % Let $q : G \rightarrow G''$ be defined for $g \in G$ by $q(g) = g'$ if, and
  % only if, $g'$ and $g$ are syntactical-equivalent in $C$.

  % For $\sigma \in \sym_n$ let $\pi_\sigma\in \aut(C)$ be an extension of
  % $\sigma$ and let $\pi_\sigma'': G'' \rightarrow G''$ be defined by
  % $\pi_{\sigma}'' (g) = q(\pi_{\sigma}g))$. It can be shown that
  % $\pi_{\sigma}''$ is an automorphism of $C''$ that extends

  % Given a circuit $g$ Let $C' := \op{make-injective-all}(\op{merge-all} (C,
  % G))$
  % (see Definition~\ref{def:make-injective} and~\ref{def:merge-all} for the
  % definitions of these functions). It follows that if $C$ is symmetric then
  % $C'$
  % is symmetric. We have from Lemma~\ref{lem:unique-labels-syntactic-equiv}
  % that
  % the syntactic-equivalence relation can be computed in polynomial time for
  % transparent circuits. It follows that the construction of $C'$ from $C$ can
  % be
  % implemented so as to run in time polynomial in the size of $C$. Clearly
  % every
  % gate $g$ in $C'$ belongs to a singleton syntactic-equivalence equivalence
  % class, and as such $C'$ has unique labels.
\end{proof}
% GW: read this when fresh

We now show that it is possible to compute the action of $\sym_n$ on the gates
of a circuit of order $n$ with unique labels in polynomial time. Recall that
from Proposition~\ref{prop:transparent-unique} we may, without a loss of
generality, restrict our attention from transparent circuits to circuits with
unique labels. The proof technique used here is similar to the one used by
Anderson and Dawar~\cite{AndersonD17}.

\begin{lem}
  There is an algorithm takes as input a $(\mathbb{B}, \rho)$-circuit $C$ of
  order $n$ with unique labels and $\sigma \in \sym_n$ and outputs for each gate
  $g$ its image under the unique automorphism extending $\sigma$ (if it exists).
  This algorithm runs in time polynomial in the combined size of the input
  circuit and the encoding of the permutation.
  \label{lem:compute-automorphisms}
\end{lem}
\begin{proof}
  Let $C := \langle G, \Omega, \Sigma, \Lambda, L \rangle$. We recursively build
  up the mapping $\pi$ extending $\sigma$. If at some point in the recursive
  construction we arrive at a point where no mapping for $g$ can be found we
  halt at that point and return that no automorphism exists.

  Let $h$ be any gate in the circuit. Suppose $h$ is an input gate. If $h$ is a
  constant gate then let $\pi (h) = h'$. If $h$ is a relational gate such that
  $R := \Sigma(h)$, then check if there exists $h'$ such that $\Sigma (h') = R$,
  $\sigma \Lambda_R(h) = \Lambda_R(h')$, and either both $h$ and $h'$ are output
  gates or neither are. If no such $h'$ exists then halt and output that no
  automorphism exists. If both $h$ and $h'$ are not output gates then set $\pi
  (h) = h'$. If both are output gates then check if $\sigma \Omega^{-1}(h) =
  \Omega^{-1}(h')$. If equality holds set $\pi(h) = h'$, otherwise halt and
  output that no automorphism exists. We have from Lemma~\ref{} that there is at
  most one gate $h'$ that meets the above criteria, and so $\pi(h)$ is well
  defined.

  Let $h$ be an internal gate in the circuit and assume we have defined $\pi
  (g)$ for every gate $g$ of depth less than $h$. Let $h'$ be a gate in the
  circuit such that $\Sigma(h) = \Sigma (h')$, $\pi L(h)$ is
  isomorphism-equivalent to $L(h')$ and, if $h$ is an output gate then $h'$ is
  an output gate such that $\sigma \Omega^{-1}(h) = \Omega^{-1}(h')$. Since $C$
  has unique labels (and so unique extensions) we have that $\pi L(h)$ is
  isomorphism-equivalent to $L(h')$ if, and only if, $L(h')^{-1}\pi L(h)$ acts
  on $\ind(h)$ like an automorphism of $\str{h}$. This can be determined easily.
  If no such $h'$ exists halt and output that there is no automorphism extending
  $\sigma$. Again, we have that if $C$ has unique labels then there is at most
  one such $h'$.

  This recursive approach can be implemented as an algorithm that runs in time
  polynomial in the combined size of the inputs and outputs the required
  automorphism if it exists.
\end{proof}

We now that, using the algorithm from Lemma~\ref{lem:compute-automorphisms}, we
can similarly compute the action of $\sym_n$ on the elements of the universes of
the gates of a circuit of order $n$ in polynomial time.

\begin{lem}
  There is an algorithm takes as input a $(\mathbb{B}, \rho)$-circuit $C$ of
  order $n$, $\sigma \in \sym_n$, $g$ a gate in $C$, and $a \in \universe{g}$
  and, if there exists an automorphism of $C$ extending $\sigma$ such that
  $\sigma in \stab(g)$, outputs image of $a$ under $\sigma$. The algorithm runs
  in time polynomial in the size of $C$ and the encoding of $\sigma$.
  \label{lem:compute-automorphisms-labels}
\end{lem}
\begin{proof}
  Let $C = \langle G, \Omega, \Sigma, \Lambda, L \rangle$. Use the algorithm
  from Lemma \ref{lem:compute-automorphisms} to check if $\sigma$ extends to an
  automorphism on $C$ and $\sigma \in \stab(g)$. If not, return that no such
  automorphism exists. We thus have that $C$ is symmetric. Let $h \in H_g$ and
  $\vec{a} := L(g)^{-1}(h)$ be such that $a \in \vec{a}$, and let $i$ be the
  index of $a$ in $\vec{a}$. Output $\sigma \cdot a := L(g)^{-1}(\sigma h)(i)$.
\end{proof}

We now show that it is possible to compute the orbits and canonical supporting
partitions of gates in circuits with unique labels in polynomial-time. We also
show that it is possible to compute the orbits and canonical supporting
partitions of elements on the universes of gates in circuits with unique labels
in polynomial-time. In order to prove both of these results we first prove a
more general result for group actions.

\begin{lem}
  \label{lem:computing-support-orbit}
  There is an algorithm that takes in a subset $S \subseteq [n]$ and $x \in X$,
  where $X$ is some finite set such that there is a group action of $\stab(S)$
  on $X$ computable in time polynomial in both $n$ and $\vert X \vert$ and
  outputs $\orb_{\stab(S)} (x)$ and $\SP_{\stab(S)}(x)$. This algorithm runs in
  time polynomial in both $\vert X \vert$ and $n$.
\end{lem}
\begin{proof}
  Let $(u, v) \in \sym_{[n] \setminus S}$ be a transposition. We note that
  $(u,v) \in \stab(S)$ and there are ${n - \vert S \vert}\choose{2}$ many such
  transpositions. For a transposition $(u, v) \in \sym_{[n] \setminus S}$, let
  \begin{align*}
    \mathcal{P}_{(u,v)} := \{ \{u,v\}\} \bigcup_{w \in [n] \setminus \{ u,v \}} \{ \{ w \} \}
  \end{align*}
  be a partition of $[n]$. Then $\mathcal{P}_{(u,v)}$ supports
  $\stab_{\stab(S)}(x)$ if, and only if, $(u,v) \cdot x = x$.

  Let $\mathcal{P}$ be the partition formed by applying $\mathcal{E}$ to all
  $\mathcal{P}_{(u,v)}$ such that $(u,v) \cdot x = x$ (which can be checked in
  time polynomial in $\vert X \vert$ and $n$). From
  Proposition~\ref{prop:combining-supporting-patitions} we have that
  $\mathcal{P}$ supports $\stab_{\stab(S)}(x)$. Suppose that $\mathcal{P}$ is
  not the coarsest supporting partition of $\stab_{\stab(S)}(x)$. Then there
  exists a partition $\mathcal{P}'$ supporting $\stab_{\stab(S)}(x)$ such that
  $\mathcal{P}' \preceq P$ and $\mathcal{P}' \neq \mathcal{P}$. And so there
  exists $P \in \mathcal{P}$ and $P' \in \mathcal{P}'$ such that $P \subsetneq
  P'$. But then there exists $a , b \in P'$ such that $a \not\in P$. Note that
  $(a,b)$ fixes $\mathcal{P}'$ and, since $\mathcal{P}'$ supports
  $\stab_{\stab(S)}(x)$, it follows that $(a,b) \in \stab_{\stab(S)}(x)$ and $a
  \not\in S$ and $b \not\in S$. But then we have that $(a,b) \cdot x = x$, and
  so $\mathcal{P}_{(a,b)}$ supports $\stab_{\stab(S))}(x)$ and thus, from the
  construction of $\mathcal{P}$, $\mathcal{P}$ is fixed by $(a,b)$. But we
  selected $a$ and $b$ such that $\mathcal{P}$ is not fixed by $(a,b)$, and so
  we have a contradiction. We thus have that $\mathcal{P}$ is the coarsest
  supporting partition of $\stab_{\stab(S)}(x)$.

  It remains to compute $\orb_{\stab(S)}(x)$. Let $M_0 := \{x\}$ and for each $i
  \geq 0$ let $M_{i+1} := M_i\cup \bigcup_{(u,v) \in \sym_{[n] \setminus S}}
  ((u,v) \cdot M_i)$. Let $M$ be the least fixed point of this sequence. It is
  easy to see that $M \subseteq \orb_{\stab(S)}(x)$ as every element of $M$ is
  equal to the action of some finite sequence of transpositions acting on $x$.
  Moreover, if $y \in \orb_{\stab(S)}(x)$, then there exists $\pi \in \stab(S)$
  such that $y = \pi \cdot x$. But then, since $\sym_{[n] \setminus S}$, is
  generated by the set of all transpositions in $\sym_{[n] \setminus S}$, it
  follows that $\pi$ can be written as a sequence of $t$ transpositions (for
  some $t \in \nats$). Thus $y \in M_t \subseteq M$, and hence
  $\orb_{\stab(S)}(x) \subseteq M$, and so $\orb_{\stab(S)}(x) = M$.

  Note that the set of all transpositions in $\sym_{[n] \setminus S}$ can be
  computed in time $\mathcal{O}(n^{2})$ and whether a transposition fixes $x$
  can be computed in time polynomial in both $\vert X \vert$ and $n$. Moreover,
  since it is easy to show that $\mathcal{E}$ can be computed in time polynomial
  in $n$, it follows that $\mathcal{P}$ (defined above) can be computed in
  $\mathcal{O} (n^{2} q(\vert X \vert, n))$, for some polynomial $q$.

  Furthermore, since, when computing the orbit, the least fixed point must be
  reached in fewer than $\vert X \vert$ iterations, and each iteration requires
  at most $\mathcal{O} (n^{2})$ applications of the group action, it follows
  that the procedure terminates in time $\mathcal{O}(\vert X \vert n^{2} q'
  (\vert X \vert, n))$, for some polynomial $q'$. The result follows.
\end{proof}

We now use Lemma~\ref{lem:computing-support-orbit} to prove the result for the
orbits and canonical supporting partitions of gates in the circuit.

\begin{lem}
  There is an algorithm that takes in a circuit $C$ of order $n$ with unique
  labels and outputs if the circuit is symmetric. If it is symmetric then it
  outputs the orbit and coarsest supporting partition of every gate. This
  algorithm runs in time polynomial in the size of the circuit.
  \label{lem:computing-support-orbit-gate}
\end{lem}

\begin{proof}
  For each transposition $(u,v) \in \sym_n$ and each gate $g$ in $C$ use
  Lemma~\ref{lem:compute-automorphisms} to check if the image of $g$ under
  $(u,v)$ exists. If for any transposition and gate the algorithm returns that
  no image exists then output that the circuit is not symmetric.

  We note that if all gates have an image under all transpositions then, since
  $\sym_n$ is generated by the set of transpositions, we have that $C$ is
  symmetric.

  We have from Lemma~\ref{lem:compute-automorphisms} that computing the action
  of $\sym_n$ on the gates of $C$ can be done in time polynomial in the size of
  the circuit. For each gate $g$ in $C$ we run the algorithm from
  Lemma~\ref{lem:computing-support-orbit} with $S := \emptyset$, $X := G$ (the
  set of gates in C), and $x := g$, and output the results.

  Since there are ${{n}\choose{2}} \leq n^2$ transpositions in $\sym_n$, the
  initial symmetry check can be completed in time polynomial in the size of the
  circuit. Moreover, from Lemmas~\ref{lem:compute-automorphisms} and
  \ref{lem:computing-support-orbit}, and the fact that $\vert C \vert \geq n$
  and $\vert C \vert \geq \vert \universe{g} \vert$, we have that the rest of
  the algorithm also runs in time polynomial in the size of the circuit.
\end{proof}

We now use Lemma~\ref{lem:computing-support-orbit} to prove the result for the
orbits and canonical supporting partitions of elements of the universes of gates
in the circuit.

\begin{lem}
  There is an algorithm that takes in a circuit $C$ of order $n$ with unique
  labels, a gate $g$ in $C$ with small support, and $a \in \universe{g}$, and
  outputs if the circuit is symmetric. If $C$ is symmetric it outputs the orbit
  $\orb_{\consp(g)} (a)$ and coarsest supporting partition $\SP_{\consp(g)}(a)$.
  This algorithm runs in time polynomial in the size of the circuit.
  \label{lem:computing-support-orbit-index}
\end{lem}
\begin{proof}
  Use the algorithm from Lemma~\ref{lem:computing-support-orbit-gate} in order
  to compute the canonical support of $g$. If the algorithm returns that $C$ is
  not symmetric, output that $C$ is not symmetric.

  Use the algorithm from Lemma~\ref{lem:computing-support-orbit}, with $S :=
  \consp(g)$, $X := \universe{g}$ and $x:= a$, and output the results.

  Notice that, since $\vert \universe{g} \vert \leq \vert C \vert$ and $n \leq
  \vert C \vert$, this algorithm runs in polynomial time.
\end{proof}

Anderson and Dawar~\cite{AndersonD17}, in their study of circuits with symmetric
gates, show that such circuits may be transformed in polynomial time into
equivalent circuits with unique extensions. They do this by showing that a
circuit with symmetric gates may be transformed into an equivalent \emph{rigid}
circuit in polynomial time, and that all rigid circuits have unique extensions.
Moreover, they also prove the existence of polynomial-time algorithms computing
the action of a permutation on a rigid circuit, and the orbits and coarsest
supporting partitions of gates in a rigid circuits. We have shown that
transparent circuits can be translated into equivalent circuits with unique
labels in polynomial time, and that circuits with unique labels have unique
extensions (see Proposition~\ref{}). Moreover, we have similarly shown that, for
transparent circuits, the action of a permutation, as well as the coarsest
supporting partition and orbit of a gate (or element of the universe of a gate),
can be computed in polynomial-time. Since all circuits with symmetric gates are
transparent, this is direct generalisation of the results of Anderson an Dawar.

\subsection{The Necessity of Transparency}
% The reader will notice that while we have included a specific requirement that
% each non-symmetric internal gate in a circuit have unique labels (i.e. that
% the circuit be transparent) no similar condition In this section we discuss
% the transparency condition on circuits. It is worth noting that, since the
% symmetric circuits of Anderson and Dawar~\cite{AndersonD17} have no
% non-symmetric internal gates, these circuits are transparent.

% we provide evidence f (which from lemma~\ref{} may be assumed to have unique
% extensions)or necessity of the transparency condition, presenting reductions
% from the graph-isomorphism problem to the problem of deciding some of this
% important properties (e.g. symmetry, syntactic-equivalence, computing the
% orbit of a gate, etc.) over general, not necessarily transparent, circuits. we
% note that in all of these cases the decision problem restricted to transparent
% circuits is polynomial-time decidable.

% we further argue that, since most natural algorithms for transforming an
% arbitrary symmetric circuit into an equivalent transparent circuit make use of
% the polynomial-time decidability of these properties, we should consider these
% hardness results evidence that computing this transformation may be as hard as
% the graph-isomorphism problem.

We have from the first subsection that a great number of useful properties are
polynomial-time computable on transparent circuits. However, it is still
reasonable to ask if the explicit inclusion of the transparency restriction is
strictly necessary for proving these results. In this subsection we present a
number of results suggesting the importance of the transparency, and unique
labels, requirements. In particular, we present reductions from the
graph-isomorphism problem to many important decision problems addressed in the
first subsection (e.g. deciding if a circuit is symmetric, if a gate has unique
labels, if two gates are syntactically-equivalent, of if two gates are in the
same orbit).

These results together strongly suggest the necessity of the transparency
condition for enabling the polynomial-time decidability of many basic properties
of circuits with non-symmetric gates. Moreover, while we do not show that there
is no polynomial-time computable transformation from a general circuit to an
equivalent transparent circuit, the difficulty associated with computing basic
circuit properties that seem essential for defining such a transformation
provide evidence in favour of this conjecture. (Indeed, the translation defined
by Anderson and Dawar~\cite{AndersonD17} from circuits with symmetric gates to
equivalent circuits with unique extensions makes explicit use of the
polynomial-time computability of syntactic-equivalence on such circuits).

As a final point, we note that while we present evidence against the ability to
transform arbitrary circuits into transparent circuits in polynomial-time, we do
show that the class of transparent circuits is polynomial-time decidable. In
contrast, we show that deciding the class of circuits with unique extensions is
at least as hard as the graph-isomorphism problem.

% We particularly note the hardness results for syntactic-equivalence note The
% difficulty of computing the orbit for non-transparent circuits is particularly
% important to note, as we make use of the existence of the polynomial-time
% decidability of the orbit in the proof of our main result.

% In particular, we show that the graph-isomorphism problem reduces to the
% problem of computing the syntactic-equivalence relation for general circuits.

% As we noted, Anderson and Dawar~\cite{AndersonD17} show that circuits (with
% symmetric gates) can be transformed in polynomial time into equivalent rigid
% circuits (in fact, their algorithm transforms these circuits into equivalent
% circuits with unique labels, a stronger condition than rigidity). Their proof
% of this result makes explicit use of the fact that the syntactic-equivalence
% relation on the gates of these circuits can be computed in polynomial-time

% These results together suggest that the a polynomial-time definable

% use of to compute the orbit is a particularly important limitation, as we make
% use of a polynomial-time algorithm for deciding the orbit of a gate (or
% element of a universe of a gate) in our proof of the main result.

% Moreover, we also show that

% However, if we allow gates that compute non-symmetric functions, it is not
% clear that the syntactic-equivalence relation remains polynomial-time
% decidable.

% . prove results analogous to Lemma~\cite{} The reader will notice that, while
% we require that the circuit families in our main result consist of transparent
% circuits, no similar explicit condition is introduced in Anderson and
% Dawar~\cite{}.

% When Anderson and Dawar~\cite{} prove these results for symmetric circuits
% (with symmetric gates) they explicitly use the fact that the syntactic
% equivalence relation can be computed in polynomial time.

% The proofs of these results by Anderson and Dawar~\cite{} can prove all of
% these results

% We now turn to the question of the importance of the transparency condition.
% It could be that, just as for the case of circuits with symmetric gates, a
% general circuit (not necessarily transparent) can be transformed in polynomial
% time into an equivalent transparent circuit. If so, from the argument above,
% we could assume without a loss of generality that $P$-uniform families of
% symmetric circuit have unique extensions, eliminating the need for the
% explicit transparency requirement present in the main theorem.

% While we do not show that no such transformation exists, we do show that many
% related problems are as hard as the graph isomorphism problem for a general,
% not-necessarily transparent, circuit). In particular, we show that the related
% problems of computing the syntactic equivalence relation for a circuit, or
% whether a gate in the circuit has unique labels, or unique extensions (in the
% sense defined below), are all also at least as hard as the graph isomorphism
% problem. We also show that computing other basic properties of a symmetric
% circuit, such as the orbit of a gate, is also at least as hard as the graph
% isomorphism problem. This is a particularly important limitation, as we make
% use of a polynomial-time algorithm for deciding the orbit of a gate (or
% element of a universe of a gate) in our proof of the main result.

% ,; address two related decision problems in Proposition \ref{} and Corollaries
% \ref{} and \ref{}. In partic

% As we will show in Proposition \ref{}, formulas of $\FPR$ can be translated
% into. For this transparent circuits. However, the transformation is
% complicated by the requirement that the resultant circuit be transparent.
% However, it is worth asking if the transparency condition is required, i.e. if
% it can be shown that

% \emph{rigid} circuit. They show that all circuits such circuits can be
% transformed into equivalent rigid circuits . They show that if a circuit is
% rigid then it has unique extensions and that every circuit with injective
% labels and symmetric gates can be transformed into an equivalent rigid circuit
% in polynomial time. We have similarly shown that a transparent circuit can be
% transformed into an equivalent one with unique labels in polynomial time, and
% that circuits with unique labels have unique extensions. Moreover, since all
% circuits with symmetric gates and injective labels are transparent, this is a
% strict generalisation of their result.

% and they show that a symmetric gate

% Using the language developed in this paper, a circuit $C_n$ is \emph{rigid} if
% it has injective labels and there are no two gates $g$ and $h$ in $C_n$ such
% that $g$ and $h$ are syntactically equivalent. Since their discussion is
% limited to circuits with symmetric gates and injective labels this is equ

% circuit, a condition analogous to the rigidity condition defined in this
% paper. The following

% It is worth noting that all symmetric circuits are transparent, and

% prove that it's a sufficient condition for establishing the uniqueness of the
% induced automorphism on the circuit. Moreover, they prove that any circuit
% (noting that they limit themselves to a discussion of circuits over a
% symmetric basis) may be transformed into an equivalent rigid circuit in
% polynomial time. As such, they may assume $P$-uniform symmetric circuit
% families are rigid without a loss of generality, and hence be may assume
% without a loss of generality that each permutation in $\sym_n$ uniquely
% extends to an automorphism of the circuit.

% However, in this subsection we show that deciding the rigidity of circuits
% over bases which may include non-symmetric functions is at least as hard as
% the graph isomorphism problem. However, we shoudl that on classes

% Although the rigidity condition we present in this paper is similarly defined,
% the existence of a polynomial-time computable transformation of a circuit into
% an equivalent rigid circuit is not as straight forward. Indeed, it is possible
% to show that the problem of determining if a given $(\mathbb{B},
% \tau)$-circuit is rigid is as least as hard as the graph isomorphism problem.
% This is a direct consequence of the following result.

% \begin{definition}
%   The \emph{syntactic equivalence class problem} is the decision problem for
%   the set of all triples $(C, g_1, g_2)$, where $C$ is a $(\mathbb{B},
%   \tau)$-circuit, and $g_1$ and $g_2$ are two syntactically equivalent gates
%   in $C$.
% \end{definition}

% \begin{definition}
%   The \emph{rigidity problem} is the decision problem for the set of all
%   circuits $C$ where $C$ is a rigid circuit.
% \end{definition}

% \begin{definition}
%   The \emph{unique labels problem} is the decision problem for the set of all
%   circuits $C$ where $C$ has unique labels.
% \end{definition}

% \begin{prop}
%   Let $\tau$ be any non-empty relational vocabulary and let $n, r, p \in
%   \nats$ with $p$ prime. There is a a polynomial time reduction from the graph
%   isomorphism problem to the problem of deciding if a given pair of gates in a
%   given symmetric rank-circuit for structures of size $n$ that (i) has
%   injective labels, (ii) contains no constant gates, and (iii) contains at
%   most two rank gates with bound $r$ and prime $p$, are syntactically
%   equivalent.
%   \label{prop:syntactic-graph-iso}
% \end{prop}
% \begin{proof}
%   We have from \cite{} that there is a polynomial time reduction from the
%   graph isomorphism problem to the bipartite isomorphism problem. The
%   bipartite isomorphism problem is the decision problem for the set of all
%   pairs of partitioned bipartite graphs $B_1 := (U_1, V_1, E_1)$ and $B_2 :=
%   (U_2, V_2, E_2)$ such that there exists a graph isomorphism $\pi : B_1
%   \rightarrow B_2$ with $\restr{\pi}{U_1} = U_2$ and $\restr{\pi}{V_1} = V_2$.

%   Suppose we are given two partitioned bipartite graphs $B_1 := (U_1, V_1,
%   E_1)$ and $B_2 := (U_2, V_2, E_2)$. We assume, without a loss of generality,
%   that there exists $a, b \in \nats$ such that $U_1 = U_2 = [a]$ and $V_1 =
%   V_2 = [b]$.

%   Let $R$ be a relation symbol in $\tau$ and let $k := \arty{R}$. Let $G_{R}
%   := \{g_{R, \vec{c}} : \vec{c} \in [n]^k\}$, $G_{\text{mid}} := \{\land_0,
%   \land_1, \neg_0, \land_{\text{out}} \}$, $G_{\rank} : = \{ g_{\rank, 1},
%   g_{\rank, 2}\}$, and $G_{\land} := \{g_{i, (u,v)} : i \in [2] \, (u,v) \in
%   E_i \}$. Let $C_n = \langle G, \Omega, \Sigma, \Lambda, L \rangle$ be a
%   $(\{\rank^{1}_2\}, \tau)$-circuit defined as follows. Let $G = G_R \cup
%   G_{\text{mid}} \cup G_{\rank} \cup G_{\land}$ and $\Omega$ be the $0$-ary
%   function $\land_{\text{out}}$. For each $\vec{c} \in [n]^{\arty{R}}$ let
%   $\Lambda(\vec{c}) = g_{R, \vec{c}}$. Define $\Sigma$ as follows. For each $g
%   \in G$,
%   \begin{itemize}
%     \setlength\itemsep{0mm}
%   \item if $g \in G_R$ then $\Sigma(g) = R$,
%   \item if $g$ equals $\land_0$ or $\land_1$ then $\Sigma(g) = \land[n^k]$
%   \item if $g$ equals $\land_{\text{out}}$ then $\Sigma(g) = \land[2]$,
%   \item if $g \in G_\rank$ let $\Sigma(g) = \rank^r_p [a,b]$, and
%   \item if $g \in G_{\text{in}}$ and $g$ then $\Sigma(g) = \land[1]$.
%   \end{itemize}
%   Define $L$ as follows. For each $g \in G$,
%   \begin{itemize}
%     \setlength\itemsep{0mm}
%   \item if $g \in G$ and $g = g_{\text{out}}$ then for each $i \in [2]$,
%     $L(g)(i) := g_{\rank, i}$,
%   \item if $g \in G_\rank$ and $g = g_{\rank, i}$ then $L(g)(p,q) = g_{i,
%     (p,q)}$,
%   \item if $g \in G_{\land}$ and $g = g_{i, (p,q)}$ then $L(g)(1) = \land_1$
%     if $(p,q) \in E_i$ and $L(g)(1) = \neg_0$ otherwise,
%   \item if $g = \neg_0$ then $L(g)(1) = \land_0$, and
%   \item if $g = \land_0$ or $g = \land_1$ then for all $p \in [n^k]$ we have
%     that $L(g)(p)$ equals $\Lambda^{-1}(\vec{c}_p)$, where $\vec{c}_p$ is the
%     $p$th element of $[n]^k$ in the lexicographical ordering on $[n]^k$.
%   \end{itemize}

%   We note that for $i \in [2]$, $L(g_{\rank, i})(p,q) = \land_1$ if $(p,q)$ is
%   an edge in in $B_i$ and $L(g_\rank, i)(p,q) = \neg_0$ if $(p,q)$ is not an
%   edge in $B_i$. We thus have that $B_1$ and $B_2$ are bipartite isomorphic
%   (in the sense defined above) if, and only if, $g_{\rank, 1}$ is
%   syntactically equivalent to $g_{\rank, 2}$.

%   Since the construction of $C_n$ can be implemented in time polynomial in the
%   combined sizes of the input graphs, the mapping of $(B_1, B_2)$ to the tuple
%   $(C_n, (g_{\rank, 1}, g_{\rank, 2}))$ witnesses the reduction. The result
%   follows.
% \end{proof}

We now present a reduction from the graph-isomorphism problem to the problem of
deciding if two gates in a circuit are syntactically-equivalent. In fact, we
prove a stronger result, presenting a reduction from the graph-isomorphism
problem to the problem of computing the syntactic-equivalence relation over an
ostensibly simpler class of circuits.

\begin{prop}
  Let $\rho$ be any non-empty relational vocabulary and let $n, r, p \in \nats$,
  with $p$ prime. There is a a polynomial time reduction from the
  graph-isomorphism problem to the problem of deciding if a given pair of gates
  in a given symmetric rank-circuit for structures of size $n$ that (i) has
  injective labels, (ii) contains no constant gates, and (iii) contains at most
  two rank gates with bound $r$ and prime $p$, are syntactically equivalent.
  \label{prop:syntactic-graph-iso}
\end{prop}
\begin{proof}
  We have from \cite{} that there is a polynomial time reduction from the graph
  isomorphism problem to the bipartite isomorphism problem. The bipartite
  isomorphism problem is the decision problem for the set of all pairs of
  partitioned bipartite graphs $B_1 := (U_1, V_1, E_1)$ and $B_2 := (U_2, V_2,
  E_2)$ such that there exists a graph isomorphism $\pi : B_1 \rightarrow B_2$
  with $\restr{\pi}{U_1} = U_2$ and $\restr{\pi}{V_1} = V_2$.

  Suppose we are given two partitioned bipartite graphs $B_1 := (U_1, V_1, E_1)$
  and $B_2 := (U_2, V_2, E_2)$. We assume, without a loss of generality, that
  there exists $a, b \in \nats$ such that $U_1 = U_2 = [a]$ and $V_1 = V_2 =
  [b]$.

  Let $R$ be a relation symbol in $\rho$ and let $k := \arty(R) > 0$. Let $G_{R}
  := \{g_{R, \vec{c}} : \vec{c} \in [n]^k\}$, $G_{\text{mid}} := \{g_{\lor},
  g_{\land}, g_{\text{out}} \}$, $G_{\rank} : = \{ g^1_{\rank}, g^2_{\rank}\}$,
  and $G_{\text{nodes}} := \{g_{i, (u,v)} : i \in [2] \, (u,v) \in [a] \times
  [b] \}$. Let $C = \langle G, \Omega, \Sigma, \Lambda, L \rangle$ be a
  $(\{\rank^{1}_2\}, \rho)$-circuit of order $n$ defined as follows. Let $G =
  G_R \cup G_{\text{mid}} \cup G_{\rank} \cup G_{\text{nodes}}$ and $\Omega$ be
  the $0$-ary function $\land_{\text{out}}$. For each $\vec{c} \in
  [n]^{\arty(R)}$ let $\Lambda_R(\vec{c}) = g_{R, \vec{c}}$. Define $\Sigma$ as
  follows. For each $g \in G$,
  \begin{itemize}
    \setlength\itemsep{0mm}
  \item if $g = \land_{\text{out}}$ then $\Sigma(g) = \land[2]$,
  \item if $g \in G_\rank$ then $\Sigma(g) = \rank^r_p [a,b]$,
  \item if $g \in G_{\text{nodes}}$ then $\Sigma(g) = \land[1]$,
  \item if $g = \land[n^k]$, $\Sigma (g) = \land[n^k]$ and if $g = g_{\lor}$
    then $\Sigma(g) = \lor[n^k]$, and
  \item if $g \in G_R$ then $\Sigma(g) = R$.
  \end{itemize}
  Define $L$ as follows. For each $g \in G$,
  \begin{itemize}
    \setlength\itemsep{0mm}
  \item if $g = \land_{\text{out}}$ then for each $i \in [2]$, $L(g)(i) :=
    g^i_{\rank}$,
  \item if $g \in G_\rank$ and $g = g^i_{\rank}$ then for all $(p,q) \in
    \ind(g)$, $L(g)(p,q) = g_{i, (p,q)}$,
  \item if $g \in G_{\text{nodes}}$ and $g = g_{i, (p,q)}$ then if $(p, q \in
    E_i)$ then $L(g)(1) = g_{\land}$, otherwise $L(g)(1) = g_\lor$, and
  \item if $g = g_\land$ or $g = g_\lor$ then for all $p \in [n^k]$ we have that
    $L(g)(p)$ equals $\Lambda^{-1}(\vec{c}_p)$, where $\vec{c}_p$ is the $p$th
    element of $[n]^k$ in the lexicographical ordering on $[n]^k$.
  \end{itemize}

  We note that for $i \in [2]$, the child of $L(g^i_{\rank})(p,q)$ equals
  $g_\land$ if, and only if, $(p,q)$ is an edge in in $B_i$ and the child of
  $L(g^i_\rank)(p,q) $ equals $g_\lor$ if, and only if, $(p,q)$ is not an edge
  in $B_i$. We thus have that $B_1$ and $B_2$ are bipartite-isomorphic (in the
  sense defined above) if, and only if, there exists $\lambda \in \sym_{a}
  \times \sym_{b} = \aut(g^1_{\rank}) = \aut(g^2_{\rank})$ such that for all
  $(u,v) \in [a] \times [b] = \ind(g^1_{\rank})$, $L(g^1_{\rank})((u,v)) = g_{1,
    (u,v)} \equiv g_{2, \lambda (u,v)} = L(g^2_{\rank})(\lambda (u,v))$. It
  follows that $B_1$ and $B_2$ are bipartite-isomorphic if, and only if,
  $g^1_{\rank} \equiv g^2_{\rank}$.

  Since the construction of $C$ can be implemented in time polynomial in the
  combined sizes of the input graphs, the mapping of $(B_1, B_2)$ to the tuple
  $(C, (g^1_{\rank}, g^2_{\rank}))$ is a reduction, and the result follows.
\end{proof}

Since there is a trivial reduction from the problem of computing the
syntactic-equivalence relation over a restricted class of circuits to the
problem over the general class of circuits, we have the following corollary.

\begin{cor}
  There is a polynomial-time reduction from the graph isomorphism problem to the
  problem of determining if a given pair of gates in a given circuit are
  syntactically-equivalent.
  \label{lem:syntactically-equivilent-general-hard}
\end{cor}

Proposition~\ref{prop:syntactic-graph-iso} gives us that computing the the
syntactic-equivalence for a given circuit remains hard even if we restrict
ourselves to circuits with injective labels. In fact, every gate but the two
rank gates (and, possibly, the output gate) have unique labels, indicating that
even if we restrict ourselves to circuits with injective labels and such that
each non-symmetric gates has the property that all of its input gates have
unique labels, the syntactic-equivalence relation remains hard to compute.

We now show that the syntactic-equivalence relation remains hard to compute if
we restrict ourselves to class of circuits such that each non-symmetric gate has
unique children.

\begin{lem}
  There is a polynomial-time reduction from the graph-isomorphism problem to the
  problem of deciding if two gates in a given circuit with the property that
  each non-symmetric gate has unique children are syntactically-equivalent.
  \label{lem:syntactic-equivalence-unique-gates-hard}
\end{lem}
\begin{proof}
  We use a similar approach as in Proposition~\ref{prop:syntactic-graph-iso},
  again using the reduction from graph-isomorphism to the bipartite
  graph-isomorphism problem.

  Suppose we are given two partitioned bipartite graphs $B_1 := (U_1, V_1, E_1)$
  and $B_2 := (U_2, V_2, E_2)$. We assume, without a loss of generality, that
  there exists $a, b \in \nats$ such that $U_1 = U_2 = [a]$ and $V_1 = V_2 =
  [b]$.

  Let $C = \langle G, , \Omega, \Sigma, \Lambda, L \rangle$ be the circuit
  defined in the proof of Proposition~\ref{prop:syntactic-graph-iso}. Let $C' =
  \langle G', \Omega', \Sigma', \Lambda', L'\rangle$ be defined as follows. Let
  $G' = G \setminus G_{\text{nodes}}$, $\Omega' = \Omega$, $\Lambda' = \Lambda$,
  and $\Sigma' = \restr{\Sigma}{G'}$. For all $g \in G' \setminus G_{\rank}$,
  $L'(g) = L(g)$. For $i \in [2]$ and $(p,q) \in [a] \times [b]$,
  $L(g^i_{\rank})(p,q) = g_{\land}$ if $(p,q) \in E_i$ and $L(g^i_{\rank})(p,q)
  = g_{\lor}$ otherwise.

  We have defined $C'$ from $C$ by deleting the gates in $G_{\text{nodes}}$ and
  for each $g \in G_{\text{nodes}}$ adding a wire directly from the child of $g$
  in to the parent of $g$. Using an argument similar to that of
  Proposition~\ref{prop:syntactic-graph-iso} we may then prove that $B_1$ and
  $B_2$ are bipartite-isomorphic if, and only if, $g^1_{\rank} \equiv
  g^2_{\rank}$. Since the construction of $C$, and so $C'$, can be implemented
  in time polynomial in the combined sizes of the input graphs, the mapping of
  $(B_1, B_2)$ to the tuple $(C', (g^1_{\rank}, g^2_{\rank}))$ is a reduction,
  and the result follows.
\end{proof}

We recall that a circuit $C$ is transparent if, and only if, every non-symmetric
gate in $C$ has injective labels and unique children. We have shown that
computing the syntactic-equivalence relation is at least as hard as the
graph-isomorphism problem if we restrict ourselves to either the class of
circuits in which non-symmetric gates have injective labels
(Proposition~\ref{prop:syntactic-graph-iso}) or the class of circuits in which
each non-symmetric gate has unique children
(Lemma~\ref{lem:syntactic-equivalence-equivalence-unique-gates-hard}).

We now present a reduction from the problem of computing the
syntactic-equivalence relation to the problem of deciding if a given gate in a
circuit has unique labels. From then transitivity of polynomial-time many-one
reductions and Proposition~\ref{prop:syntactic-graph-iso}, this gives us a
reduction from the graph-isomorphism problem to the problem of deciding if a
gate has unique labels.

\begin{lem}
  There are a polynomial-time reduction from the problem of determining if a
  given pair of gates in a given circuit are syntactically equivalent to the
  problems of determining if a given gate in a given circuit has unique labels.
  \label{lem:syntactically-equivalent-unique-labels}
\end{lem}

\begin{proof}
  Let $C := \langle G, \Omega, \Sigma, \Lambda, L \rangle$ be a circuit of order
  $n$ and let $g_1, g_2 \in G$. Let $D$ be the circuit formed from $C$ by
  removing every gate $g \in G \setminus\{g_1, g_2\}$ such that $\neg W_t(g,
  g_1) \land \neg W_t(g, g_2)$, where $W_t$ is the transitive closure of the $W$
  relation (i.e. we remove all those gates in the circuit such that there is no
  path from the gate to either $g_1$ or $g_2$). Let $C'$ be the circuit formed
  from $D$ by adding in a single two-input $\AND$-gate $g'$ and connecting the
  outputs of $g_1$ and $g_2$ to the inputs of $g'$. Moreover, we let this $g'$
  be the single output gate of $C'$.

  It follows that $g_1$ and $g_2$ are syntactically-equivalent in $C$ if, and
  only if, $g'$ has unique labels in $C'$. Since the construction of $C'$ from
  $C$ can be completed in polynomial time, the mapping of $(C, (g_1, g_2))$ to
  $(C', g')$ is a reduction.
\end{proof}

We can construct a similar argument for reducing the problem of deciding if two
gates are syntactically equivalent in a circuit with injective labels to the
problem of deciding if a given gate $g$ in a circuit $C$ with injective labels
does not have \emph{unique extensions}. We say a gate $g$ has unique extensions
if there is no permutation such that two automorphisms of the circuit extend the
permutation and disagree with each other on $g$ (i.e. $g$ is not a
counterexample to $C$ having unique extensions). From then transitivity of
polynomial-time many-one reductions and
Proposition~\ref{prop:syntactic-graph-iso}, this gives us a reduction from the
graph-isomorphism problem to the problem of deciding if a gate has unique
extensions.

\begin{lem}
  There is a polynomial-time reduction from the problem of determining if a
  given pair of gates gates in a given $(\mathbb{B}, \rho)$-circuit with
  injective labels are syntactically-equivalent to the problem of determining if
  for a given pair $(C, g)$, where $C$ is a circuit of order $n$ with injective
  labels, $g$ is a gate in $C$, that there exists $\sigma \in \sym_n$ and
  automorphisms $\pi, \pi' \in \aut(C)$ extending $\sigma$ such that $\pi (g)
  \neq \pi' (g)$.
  \label{lem:syntactically-equivilent-unique-extensions}
\end{lem}
\begin{proof}
  Let $C$ be a circuit of order $n$ and let $g_1$ and $g_2$ be two gates in $C$.
  Note that for any gate $g$ in $\sigma \in sym_n$, if $\pi, \pi' \in \aut(C)$
  extend $\sigma$ and $\pi_e := \pi'\pi{-1}$ then $\pi(g) \neq \pi'(g)$ if, and
  only if, $\pi_e (g) \neq g$. It follows that there exists $\sigma \in sym_n$
  such that there are automorphisms $\pi, \pi' \in \aut(C)$ extending $\sigma$
  such that $\pi (g) \neq \pi'(g)$ if, and only if, there exists $\pi_e$
  extending the trivial permutation such that $\pi (g) \neq g$.

  Let $C'$ be the circuit constructed from $C$ as in the proof of
  Lemma~\ref{lem:syntactically-equivalent-unique-labels}. We now prove that the
  mapping $(C, g_1, g_2)$ to $(C', g_1)$ is a reduction. Let $\pi_e$ be
  bijection from the gates of $C'$ to the gates of $C'$ that swaps $g_1$ and
  $g_2$ and fixes all other gates. It follows that if $g_1$ and $g_2$ are
  syntactically-equivalent in $C$, then they are syntactically-equivalent in
  $C'$, and so $\pi_e$ is a non-trivial automorphism extending the trivial
  permutation, and $g_1$ does not have unique extensions in $C'$. In the other
  direction, suppose $g_1$ does not have unique extensions in $C'$. Then there
  exists an automorphism $\pi_e \in \aut(C')$ extending the trivial permutation
  and such that $\pi_e(g_1) \neq g_1$. But $g_1$ is a child of the single output
  gate $g'$ (which must be fixed by any automorphism), and the only other child
  of $g'$ is $g_2$. It follows $\pi_e$ swaps $g_1$ and $g_2$, and so $g_1$ and
  $g_2$ are syntactically-equivalent in $C'$. The result follows.
\end{proof}

We now show that there is a reduction from the graph-isomorphism problem to the
problem of deciding if a given circuit is symmetric. In fact, we prove a
stronger result, showing that this reduction holds even if we restrict ourselves
to the class of circuits with trivial syntactic-equivalence classes and in
which, all but two gates in the circuit have unique labels.

\begin{prop}
  The graph-isomorphism problem is polynomial-time reducible to the problem of
  deciding if a circuit with trivial syntactic-equivalence classes and in which
  all but two gates in the circuit have unique labels is symmetric.
  \label{prop:graph-iso-symmetric}
\end{prop}
\begin{proof}
  We use a similar approach as in Proposition~\ref{prop:syntactic-graph-iso},
  and so proving a reduction from the bipartite-graph-isomorphism to the problem
  of deciding if a circuit is symmetric.

  Suppose we are given two partitioned bipartite graphs $B_1 := (U_1, V_1, E_1)$
  and $B_2 := (U_2, V_2, E_2)$. We assume, without a loss of generality, that
  there exists $a, b \in \nats$ such that $U_1 = U_2 = [a]$ and $V_1 = V_2 =
  [b]$.

  Let $\rho:= \{R\}$ be a relational vocabulary, where $R$ is a unary relational
  symbol. Fix any number $r$ and prime $p$. We define a $(\rho, \RB)$-circuit $C
  := \langle G, \Omega, \Sigma, \Lambda, L \rangle$ of order two as follows. Let
  $G_{R} := \{g^1_{R}, g^1_{R}\}$ and $G_{\text{mid}} := \{g^1_{\land},
  g^2_\land, g^1_\lor, g^2_\lor , g_{\text{out}}\}$, and $G_{\rank} : = \{
  g^1_{\rank}, g^2_{\rank}\}$. Let $G = G_R \cup G_{\text{mid}} \cup G_{\rank}$
  and $\Omega$ be the $0$-ary function $g_{\text{out}}$. Let $\Lambda(1) :=
  g^1_{R}$ and $\Lambda(2) := g^2_{R}$. Define $\Sigma$ as follows. For each $g
  \in G$,
  \begin{itemizens}
  \item if $g$ equals $\land_{\text{out}}$ then $\Sigma(g) = \land[2]$,
  \item if $g \in G_\rank$ let $\Sigma(g) = \rank^r_p [a,b]$,
  \item if $g \in G_{\text{mid}}$ and $g = g^i_s$ for $ \in [2]$ and symbol $s$
    then $\Sigma(g) = s[1]$, and
  \item if $g \in G_R$ then $\Sigma(g) = R$.
  \end{itemizens}
  Define $L$ as follows. For each $g \in G$,
  \begin{itemizens}
  \item if $g = g_{\text{out}}$ then for each $i \in [2]$, $L(g)(i) :=
    g^i_{\rank}$,
  \item if $g \in G_\rank$ and $g = g^i_{\rank}$ for some $i \in [2]$ then for
    $(p,q) \in [a]\times [b]$, $L(g)(p,q) = g^i_\land$ if $(p,q) \in E_i$ and
    $L(g)(p,q) = g^i_\lor$ otherwise, and
  \item if $g = g^i_\land$ or $g = g^i_{\lor}$ for some $i \in [2]$, then
    $L(g)(1) = \Lambda^{-1}(i)$.
  \end{itemizens}

  We note that for $i \in [2]$, $L(g^i_{\rank})(p,q)= g^i_\land$ if, and only
  if, $(p,q)$ is an edge in $B_i$ and $L(g^i_{\rank})(p,q) g^i_\lor$ if, and
  only if, $(p,q)$ is not an edge in $B_i$. Let $\pi_{(1,2)} : G \rightarrow G$
  that fixes the output gate, and swaps each gate $g^1_s$, for some symbol $s$,
  with the gate $g^1_s$. It is easy to see that $\pi_{(1,2)}$ is a bijection.
  Moreover, $C$ is symmetric if, and only if, $\pi_{(1,2)}$ is an automorphism
  of the circuit extending the transposition $(1,2)$, if and only if, $(1,2)
  (L(g^1_{\rank}))$ is isomorphism-equivalent to $L(g^2_{\rank})$ if, and only
  if, $B_1$ and $B_2$ are bipartite-isomorphic.

  % Let
  % $G_{R} := \{g_{R, 1}, g_{R_2}\}$ and $G_{\text{mid}} := \{g^1_{\land},
  % g^2_\land, g^1_\lor, g^2_\lor , \land_{\text{out}}\}$, $G_{\rank} : = \{
  % g_{\rank, 1}, g_{\rank, 2}\}$, and $G_{\text{nodes}} := \{g_{i, (u,v)} : i
  % \in
  % [2] \, (u,v) \in [a] times [b] \}$. Let $G = G_R \cup G_{\text{mid}} \cup
  % G_{\rank} \cup G_{\text{nodes}}$ and $\Omega$ be the $0$-ary function
  % $\land_{\text{out}}$. Let $\Lambda(1) := g_{R, 1}$ and $\Lambda(2) :=
  % g_{R,2}$. Define $\Sigma$ as follows. For each $g \in G$,
  % \begin{itemize}
  %   \setlength\itemsep{0mm}
  % \item if $g$ equals $\land_{\text{out}}$ then $\Sigma(g) = \land[2]$,
  % \item if $g \in G_\rank$ let $\Sigma(g) = \rank^r_p [a,b]$,
  % \item if $g \in G_{\text{nodes}}$ then $\Sigma(g) = \land[1]$,
  % \item if $g \in G_{\text{mid}}$ and $g = g^i_s$ for $ \in [2]$ and symbol
  %   $s$
  %   then $\Sigma(g) = s[1]$, and
  % \item if $g \in G_R$ then $\Sigma(g) = R$.
  % \end{itemize}
  % Define $L$ as follows. For each $g \in G$,
  % \begin{itemize}
  %   \setlength\itemsep{0mm}
  % \item if $g = g_{\text{out}}$ then for each $i \in [2]$, $L(g)(i) :=
  %   g_{\rank,
  %   i}$,
  % \item if $g \in G_\rank$ and $g = g_{\rank, i}$ then $L(g)(p,q) = g_{i,
  %   (p,q)}$,
  % \item if $g \in G_{\text{nodes}}$ and $g = g_{i, (p,q)}$ for some $i \in
  %   [2]$,
  %   then $L(g)(1) = g^i_\land$ if $(p,q) \in E_i$ and $L(g)(1) = g^i_\lor$
  %   otherwise , and
  % \item if $g = g^i_\land$ or $g = g^i_{\lor}$ for some $i \in [2]$, then
  %   $L(g)(1) = \Lambda^{-1}(i)$.
  % \end{itemize}

  % We note that for $i \in [2]$, the child of $L(g_{\rank, i})(p,q)$ equals
  % $g^i_\land$ if, and only if, $(p,q)$ is an edge in $B_i$ and the child of
  % $L(g_{\rank, i})(p,q)$ equals $g^i_\lor$ if, and only if, $(p,q)$ is not an
  % edge in $B_i$. It can be shown that $C$ is symmetric if, and only if, there
  % exists an extension of the transposition $(1,2)$ to an automorphism of the
  % circuit, which holds if, and only if, there exists an extension of $(1,2)$
  % to
  % an automorphism of the circuit such that $(1,2) (L(g_{\rank, 1}))$ is
  % isomorphism-equivalent to $L(g_{\rank, 2})$, which holds if, and only if,
  % $B_1$ and $B_2$ are bipartite-isomorphic.

  % Suppose $C$ is not symmetric. If $\vert L(g_{\rank,1})^{-1}(\land^1_1) \vert
  % \neq L_(g_{\rank,2})^{-1}(\land^2_1) \vert$ then clear $B_1$ and $B_2$ are
  % not biparetite-isomorphic as they have different numbers of edges. Suppose
  % $\vert L(g_{\rank,1})^{-1}(\land^1_1) \vert =
  % L_(g_{\rank,2})^{-1}(\land^2_1) \vert$. Then the only reason $C$ would not
  % be symmetric is that there is no extension of $\sigma$ to the circuit t

  % If $B_1$ and $B_2$ are

  % We thus have that if $B_1$ and $B_2$ are bipartite-isomorphic

  % (in the sense defined in the proof of
  % Proposition~\ref{prop:syntactic-graph-iso}) if, and only if, $g_{\rank, 1}$
  % is syntactically-equivalent to $g_{\rank, 2}$.
  It is easy to see that all gates in the circuit are part of a singleton
  syntactic-equivalence class and that all gates have unique labels except for
  the two rank gates. Since the construction of $C$ can be implemented in time
  polynomial in the combined sizes of the input graphs, the mapping of $(B_1,
  B_2)$ to $C$ is a reduction, and the result follows.
\end{proof}

Note that, in Proposition~\ref{prop:graph-iso-symmetric} we again have a
hardness result for deciding a basic circuit property even if we restrict
ourselves to circuits with trivial syntactic-equivalence classes (and so ones
where each gate has unique children). It is possible to alter this reduction,
using a construction analogous to the one used in the proof of
Proposition~\ref{prop:syntactic-graph-iso}, to prove that deciding symmetry is
also hard for circuits with injective labels. In contrast, we have from
Proposition~\ref{lem:computing-support-orbit-index} that we can decide symmetry
for transparent circuits (i.e. circuits where every non-symmetric gate has
unique children and injective labels) in polynomial-time. These observations
again suggest both the robustness of this hardness result, and the importance of
the transparency condition.

We now show that computing the orbit of a gate for a general circuit is as hard
as the graph isomorphism problem. We make use of the ability to compute the
orbit of a gate in polynomial-time in
Section~\ref{sec:translating-formulas-to-FPR}, when translating families of
circuits into equivalent formulas.

\begin{lem}
  There is a polynomial-time reduction from the graph isomorphism problem to the
  problem of computing if two gates in a given circuit are in the same orbit.
  \label{lem:graph-iso-to-orbit}
\end{lem}
\begin{proof}
  Let $C$ be the circuit constructed in
  Proposition~\ref{prop:syntactic-graph-iso}. We again make use of the reduction
  from the graph isomorphism problem to the problem of deciding if two bipartite
  graphs are bipartite-isomorphic. Let $B_1$ and $B_2$ be two bipartite graphs.
  The mapping $(B_1, B_2)$ of to $(C, g^1_{\rank}, g^2_{\rank})$ is a reduction.
\end{proof}

We have from Propositions~\ref{prop:syntactic-graph-iso}
and~\ref{prop:graph-iso-symmetric}, and
Lemmas~\ref{lem:syntactically-equivalent-unique-labels},
\ref{lem:graph-iso-to-orbit}
and~\ref{lem:syntactically-equivilent-unique-extensions}, that a number of basic
circuit properties are at least as hard for general circuits as the
graph-isomorphism problem. In contrast, all of these properties are known to be
polynomial-time decidable for transparent circuits (as follows from
Lemmas~\ref{lem:unique-labels-syntactic-equiv}
and~\ref{lem:computing-support-orbit-gate}).

The hardness results for the problems of computing syntactic-equivalence
relation and deciding if a gate has unique labels are particularly worth noting
as many natural algorithms for translating a circuit into an equivalent
transparent circuit (or one with unique labels) make use of polynomial-time
decidability of these properties. Indeed, the translation defined by Anderson
and Dawar~\cite{AndersonD17} from circuits with symmetric gates to equivalent
circuits with unique extensions makes explicit use of the polynomial-time
computability of the syntactic-equivalence relation on such circuits. As such,
these results, along with the hardness results for many other basic circuit
properties (e.g. symmetry, unique extensions, or the orbits of a gate) provides
evidence against the existence of an easily-definable polynomial-time
translation from circuits to transparent circuits, or circuits with unique
labels.

% We note that in all of the above casesgive us that many circuit properties of
% interest are harder to compute than the graph-isomorphism problem. In
% particular, determining

% We have shown that basic properties of a circuit, including the ability to
% compute the syntactic equivalence relation or whether a gate has unique
% labels, are as hard to decide as the graph-isomorphism problem. Moreover,

% incluidng that various properties of the circuit that seem necessary for
% defining a translation of a circuit into an equivalent transparent circuit are
% as hard to compute as the graph-isomorphism problem. We note that all of these
% properties are polynomial-time computable for transparent circuits (as follows
% from Lemma~\ref{} and Lemma~\ref{}).

While we have provided evidence against the ability to efficiently transform a
circuit into an equivalent transparent circuit, we now show that, in contrast,
there is a polynomial-time algorithm for checking if a circuit is transparent.
This has a certain significance, as we now have transparency is not some
unverifiable condition on circuits, further motivating the usefulness of this
definition.

\begin{prop}
  There is an algorithm that takes as input a circuit, decides if that circuit
  is transparent and runs in time polynomial in the size of the circuit.
  \label{prop:transparent-polynomial-time}
\end{prop}
\begin{proof}
  Let $C = \langle G, \Omega, \Sigma, \Lambda, L \rangle$ be a $(\mathbb{B},
  \rho)$-circuit. We first check that, for each non-symmetric gate $g \in G$,
  $L(g)$ is an injection. If not, we return that $C$ is not transparent.

  For each $p \in \nats$ let $G^p \subseteq G$ be the set of all
  gates of depth $p$ and let $G^{\leq p} = \bigcup_{0 \leq i \leq p}G^i$. Note
  that each $g \in G^1$, $g$ has unique labels. Thus, from
  Lemma~\ref{lem:transparent-syntactic-equiv}, we can compute the syntactic
  equivalence relation between gates in $G^{\leq 1}$ efficiently. We construct a
  loop as follows. Let $p$ be initialised to $1$. We have that every
  non-symmetric gate $g \in G^{\leq p}$ has unique gates. We compute the
  syntactic-equivalence relation on $G^{\leq p}$ and check if each non-symmetric
  gate in $g \in G^{i + 1}$ has unique labels. If not, halt and output that the
  circuit is not transparent. If $p = \depth(C) -1$ then halt and output that
  the circuit is transparent. If neither halt state is reached, then increment
  $p$ by one and continue to the next iteration of the loop.

  We note that in every iteration of the loop we have that every non-symmetric
  gate in $G^{\leq p}$ has unique labels. Thus, from Proposition
  \ref{lem:transparent-syntactic-equiv}, we may compute the
  syntactic-equivalence relation on this set of gates efficiently. It follows
  that the above algorithm can be implemented so as to run in polynomial time.
\end{proof}

We similarly have that the unique labels property on circuits can be decided in
time polynomial in the number of gates in the circuit.

\begin{cor}
  There is an algorithm that takes in a circuit and decides if that circuit has
  unique labels and runs in time polynomial in the size of the circuit.
\end{cor}
\begin{proof}
  Let $C$ be the input circuit. From
  Proposition~\ref{prop:transparent-polynomial-time} we may check if $C$ is
  transparent in time polynomial in the size of $C$. If $C$ is not transparent
  halt and output that $C$ does not have unique labels. If $C$ is transparent
  then from Lemma~\ref{lem:transparent-syntactic-equiv} we may compute the
  syntactic-equivalence relation on the gates of $C$ in time polynomial in the
  size of $C$. Note that $C$ has unique labels if, and only if, for each gate
  $g$ in $C$, $\vert \ind(g) \vert = \vert H_g /_\equiv \vert$. We may thus
  check if $C$ has unique labels by iterating through the gates of $C$. This can
  be done in time polynomial in the size of $C$.
\end{proof}

% perhaps related to the problem of translating a circuit into an equivalent
% transparent circuit are at least as hard as the graph isomorphism problem. We
% now show that additionally recognising if a circuit has unique extensions is
% at least as hard as the graph isomorphism problem.

We have restricted ourselves to transparent circuits in large part because they
can be transformed in polynomial-time into circuits with unique extensions and
because the transparency property is polynomial-time verifiable. It may be asked
why we do not just include an explicit restriction to circuits with unique
extensions. We now show that, in contrast with transparency, the problem of
deciding if a circuit has unique-extensions is at least as hard as the
graph-isomorphism problem.

\begin{lem}
  There is a polynomial-time reduction from the graph isomorphism problem to the
  problem of determining if a given circuit does not have unique extensions.
  \label{lem:unique-extensions-hard}
\end{lem}
\begin{proof}
  We use a similar approach as in Proposition~\ref{prop:syntactic-graph-iso},
  again arguing using a reduction from the graph-isomorphism problem to
  bipartite-graph isomorphism problem. Suppose we are given two partitioned
  bipartite graphs $B_1 := (U_1, V_1, E_1)$ and $B_2 := (U_2, V_2, E_2)$. We
  assume, without a loss of generality, that there exists $a, b \in \nats$ such
  that $U_1 = U_2 = [a]$ and $V_1 = V_2 = [b]$.

  Let $C := \langle G, \Omega, \Sigma, \Lambda, L \rangle$ be a $(\RB,
  \{R\})$-circuit of order one defined as follows. Let $G_R := \{g_R\}$, $G_1 :=
  \{ g_{\lor}, g_{\land} \}$, $G_{\rank} := \{ g^1_{\rank}, g^2_{\rank} \}$ and
  $G = G_R \cup G_1 \cup G_1 \cup G_{\rank} \cup \{ g_{\text{out}} \}$. Let
  $\Omega = g_{\text{out}}$ an $\Lambda_R (g_R) = 1$. Let $\Sigma (g_R) = R$,
  $\Sigma (g_\lor) = \lor[1]$, $\Sigma (g_\land) = \land[1]$, $\Sigma
  (g_{\text{out}}) = \land[2]$ and for all $g \in G_{\rank}$, $\Sigma (g) =
  \land[a,b]$. Let $L(g_\lor)(1) L(g_\land)(1) = g_R$, for all $i \in [2]$,
  $L(g_{\text{out}})(i) = g^i_{\rank}$ and for all $(u, v) \in \ind(g)$,
  $L(g^i_{\rank})(u,v) = g_{\land}$ if $(u, v) \in E_i$ and $L(g^i_{\rank} (u,v)
  = g_{\lor}$ otherwise.

  Suppose $B_1$ and $B_2$ are bipartite-isomorphic. Then there exists $\lambda
  \in [a] \times [b]$ such that $(u,v) \in E_1$ if, and only if, $\lambda (u,v)
  \in E_2$. It follows that $L(g^1_{\rank})((u,v)) =
  L(g^2_{\rank})(\lambda(u,v))$. Let $\pi_{\lambda}: G \rightarrow G$ be defined
  such that if $g = g^i_{\rank}$ for $i \in [2]$, $\pi_{\lambda} g =
  g^j_{\rank}$ where $j \in [2] \setminus \{i\}$ and otherwise $\pi_{\lambda} g
  = g$. It is easy to see that $\pi_{\lambda}$ is a bijection. Moveover, for
  $(u,v) \in \sym_a \times \sym_b$ $\pi_{\lambda} L(g^1_{\rank})((u,v)) =
  L(g^2_{\rank})(\lambda (u,v)) = L(\pi_\lambda g^1_{\rank})(\lambda (u,v))$.
  Similarly, $\pi_\lambda L(g^2_{\rank})((u,v)) =
  L(\pi_{\lambda}g^2_{\rank})(\lambda^{-1}(u,v))$. It follows that $\pi_\lambda$
  is a non-trivial automorphism of $C$ extending the trivial permutation, and so
  $C$ does not have unique extensions.

  Suppose $C$ does not have unique extensions. Then there exists an automorphism
  $\pi$ of the circuit extending the trivial permutation such that $\pi
  g^1_{\rank} = g^2_{\rank}$. But then there exists $\lambda \in \sym_a \times
  \sym_b$ such that for all $(u,v) \in [a]\times [b]$, $\pi
  L(g^1_{\rank})((u,v)) = L(g^1_{\rank})((u,v)) = L(g^2_{\rank})(\lambda
  (u,v))$. But then $(u,v) \in E_1$ if, and only if, $L(g^1_{\rank})((u,v)) =
  g_{\land}$, if and only if, $L(g^2_{\rank})(\lambda(u,v)) = g_{\land}$ if, and
  only if, $\lambda (u,v) \in E_2$. It follows $B_1$ and $B_2$ are
  bipartite-isomorphic.
\end{proof}

% \begin{proof}
%   We use a similar argument as for Proposition~\ref{prop:syntactic-graph-iso}.
%   We again argue using a reduction from the bipartite isomorphism problem. Let
%   $B_1 := (U_1, V_1, E_1)$ and $B_2 := (U_2, V_2, E_2)$, and let $C = \langle
%   G, \Omega, \Sigma, \Lambda, L\rangle$ be the associated circuit defined in
%   the proof of Proposition~\ref{prop:syntactic-graph-iso}. We recall that we
%   may assume, without a loss of generality, that there exists $a, b \in \nats$
%   such that $[a] = U_1 = U_2$ and $[b] = V_1 = V_2$ and $B_1$ and $B_2$ are
%   not identical. Recall that $B_1$ and $B_2$ are bipartite-isomorphic if, and
%   only if, there exists $\lambda \in \sym_{a} \times \sym_{b} = \aut(g_{\rank,
%   1}) = \aut(g_{\rank, 2})$ such that for all $(u,v) \in [a] \times [b] =
%   \ind(g_{\rank, 1})$, $L(g_{\rank_,1})((u,v)) = g_{1, (u,v)} \equiv g_{2,
%   \lambda (u,v)} = L(g_{\rank, 2})(\lambda (u,v))$.

%   Suppose $B_1$ and $B_2$ are bipartite-isomorphic. Then there exists $\lambda
%   \in \sym_{a} \times \sym_b$ be such that for all $(u,v) \in [a] \times [b]$,
%   $L(g_{\rank_,1})((u,v)) \equiv L(g_{\rank, 2})(\lambda (u,v))$. Let
%   $\pi_\lambda : G \rightarrow G$ be defined as follows. Let $g \in G$. If $g
%   \not\in G_{\rank} \cup G_{nodes}$ let $\pi_\lambda g = g$. If $g \in
%   G_{nodes}$ then $g = g_{i, (u,v)}$ for some $i \in [2]$ and $(u,v) \in [a]
%   \times [b]$, and let $\pi_{\lambda} g = g_{j, \lambda (u,v)}$, where $\{i,
%   j\} = [2]$. If $g \in G_{\rank}$ then $g = g_{\rank, i}$ for some $i \in
%   [2]$, and let $\pi_\lambda g = g_{\rank, j}$, where $\{i, j\} = [2]$. It is
%   easy to see that $\pi_\lambda$ is a bijection and $\pi_\lambda$ fixes the
%   input gates. Moreover, for $i, j \in [2]$ distinct, $(u,v) \in
%   \ind(g_{\rank, i})$, $\pi_\lambda L(g_{\rank, i}))((u,v)) = \pi_\lambda
%   g_{i, (u,v)} = g_{j, \lambda (u,v)} = L( \pi_\lambda g_{\rank, i})(\lambda
%   (u,v))$. It follows that $\pi_\lambda$ is an automorphism of $C$ extending
%   the trivial permutation. Moreover, if $\lambda$ is not the identity
%   automorphism then $\pi_\lambda$ is not the identity automorphism, and so $C$
%   does not have unique extensions.

%   It $C$ does not have unique extensions then there is no permutation that
%   maps $g_{\rank, 1}$ to $g_{\rank, 2}$. In other words there exists

%   there exists $\lambda \in \sym_{a} \times \sym_{b} = \aut(g_{\rank, 1})$
%   such that for all $(u,v) \in \ind(g_{\rank, 1})$, $g_{1,
%   (u,v)}L(g_{\rank_,1})(x) = L(g_{\rank, 2})(\lambda x) = g_{2, \lambda
%   (u,v)}$.

%   if, and only if, $g_{\rank, 1}$ and $g_{\rank, 2}$ are
%   syntactically-equivalent. If $g_{\rank,1}$ and $g_{\rank, 2}$ are
%   syntactically-equivalent then there exists $\lambda \in \aut(g_{\rank,1})$
%   such that $L(g_{\rank, 1})(\lambda x) \equiv L(g_{\rank, 2})(x)$. We can
%   define an automorphism $\pi : G \rightarrow G$ such that for $g \in G$, $\pi
%   (g) = g$ if $g$ is an input gate, $\pi g_{\rank, 1} = g_{\rank, 2}$ and $\pi
%   g_{\rank, 2} = g_{\rank, 1}$ and

%   Since Both $g_{\rank, 1}$ and $g_{\rank, 2}$ have unique labels.

%   Let $C'$ be a quotient of $C$.

%   which holds if, and only if, there exists $\pi \in \aut(C')$ extending the
%   trivial permutation such that $\pi (g_{\rank, 1}) = g_{\rank, 2}$, which is
%   true if, and only if, $C'$ does not have unique extensions. it follows that
%   the mapping of $(B_1, B_2)$ to $C'$ defines a reduction.
% \end{proof}

\begin{remark}
  It's interesting to note that the circuit used in the reduction does not have
  injective labels. If we instead ask about a reduction from GI to the problem
  of deciding if a circuit has unique extensions and injective labels, then I
  can reduce from to the problem of deciding if two rigid graphs are isomorphic.
  It is unknown if this decision problem is as hard as the full graph
  isomorphism problem. After looking around a bit I discovered that it's a very
  important open problem in quantum computation. Is this worth mentioning?
\end{remark}

% We have that if a circuit is transparent then we may compute in polynomial
% time the syntactic equivalence relation on the gates of that circuit, and
% hence compute an equivalent circuit that has unique extensions. However, we've
% shown that absent the transparency condition, computing the the syntactic
% equivalence relation is at least as hard as the graph isomorphism problem.
% Moreover, other properties that can be computed in polynomial time for
% transparent circuits, e.g. deciding if particular gates have unique labels or
% might act as counterexamples to the unique extensions, have also been shown to
% be at least as the graph isomorphism problem.

% While we have not shown that there is no polynomial time algorithm that
% translates an arbitrary circuit into an equivalent transparent circuit, we
% have provided evidence in favour of the conjecture that such an algorithm
% might not exist.

% \begin{remark}
%   I really don't like the slightly vague language here. I would prefer to say
%   this differently. How does one provide `evidence' that a functional problem
%   is hard, when one has proved results about a few related decision problems?
% \end{remark}

% However, we must then address another question: If we must include an explicit
% requirement on circuits, why not simply require that the circuit have unique
% extensions? But we have from Lemma \ref{lem:unique-extensions-hard} that at
% the least verifying that a circuit have unique extensions is at least as hard
% as the graph isomorphism problem. We show that, in contrast, deciding if a
% circuit is transparent is in $\PT$.

% \begin{prop}
%   There is an algorithm that takes as input a circuit, decides if that circuit
%   is transitive and runs in time polynomial in the size of the circuit.
%   \label{prop:tansitive-polynomial-time}
% \end{prop}
% \begin{proof}
%   Let $C = \langle G, \Omega, \Sigma, \Lambda, L \rangle$ be a $(\mathbb{B},
%   \tau)$-circuit. We first check that, for each non-symmetric gate $g \in G$,
%   $L(g)$ is an injection. If not, we return that $C$ is not transitive.

%   The height of a gate $g$ in the circuit is the distance from $g$ to the set
%   of input gates. For each $p \in nats$ let $G^p \subseteq G$ be the set of
%   all gates of height $p$ and let $G^{\leq p} = \bigcup_{0 \leq i \leq p}G^i$.
%   Note that each $g \in G^1$, $g$ has unique labels. Thus, from Lemma \ref{},
%   we can compute the syntactic equivalence relation between gates in $G^{\leq
%   1}$ efficiently. We construct a loop as follows. Let $p$ be initialised to
%   $1$. We have that every non-symmetric gate $g \in G^{\leq p}$ has unique
%   gates. We compute the syntactic equivalence relation on $G^{\leq p}$ and
%   check if each non-symmetric gate in $g \in G^{i + 1}$ has unique labels. If
%   not, halt and output that the circuit is not transitive. If $p = \depth(C)
%   -1$ then halt and output that the circuit is transitive. If neither halt
%   state is reached, then increment $p$ by one and continue to the next
%   iteration of the loop.

%   We note that in every iteration of the loop we have that every non-symmetric
%   gate in $G^{\leq p}$ has unique labels. Thus, from Proposition
%   \ref{prop:unique-labels-syntactic-equiv}, we may compute the syntactic
%   equivalence relation on this set of gates efficiently. Thus it is easy to
%   see that the above algorithm can be implemented so as to run in polynomial
%   time.
% \end{proof}

% A similar result for circuits with unique labels.

% \begin{cor}
%   There is an algorithm that takes in a circuit and decides if that circuit
%   has unique labels and runs in time polynomial in the size of the circuit.
% \end{cor}
% \begin{proof}
%   Let $C$ be the input circuit. From Proposition
%   \ref{prop:transitive-polynomial-time} we may check if $C$ is transitive in
%   polynomial time. If $C$ is not transitive halt and output that $C$ does not
%   have unique labels. If $C$ is transitive then from Lemma \ref{} we may
%   compute the syntactic equivalence relation on the gates of $C$ efficiently.
%   This then allows us to check if each gate in $C$ has unique labels.
% \end{proof}

% We thus have that we can verify in polynomial time that a family of circuits
% indeed is transparent or has unique labels. The same cannot be said for the
% unique extensions property.

% \begin{lem}
%   There is an algorithm that takes in a symmetric circuit $C$ of order $n$
%   with unique labels, a gate $g$ in $C$ and $a \in \universe{g}$, the orbit
%   $\orb_g (a)$ and coarsest supporting partition $\SP_g (a)$.
% \end{lem}
% \begin{proof}
%   Let $(u,v) \in \sym_n$ be a transposition. We can check if $(u,v) \in
%   \stab(g)$ using Lemma~\ref{lem:compute-automorphisms}. Then $(u,v)$ acts
%   like an isomorphism at $g$, and hence $(u,v) \cdot a = L(g)^{-1}((\sigma
%   L(g)(\vec{x}))) (i)$, where $\vec{x} \in \ind(g)$ such that $\vec{x} (i) =
%   a$ for some $i \in [\vert \vec{x} \vert]$. Thus, since
%   Lemma~\ref{lem:compute-automorphisms} gives us that that we can compute
%   $\sigma L(g)(\vec{x})$ in polynomial time, we can compute $(u,v) \cdot a$ in
%   polynomial-time.

%   We note that

%   and hence we can compute $\sigma \cdot a$ effeciently. and hence compute
%   $\orb_g(a) = \{ \sigma \cdot a : \sigma \in \stab(g)\}$.

% \end{proof}

% Once again we note that computing the orbit of a gate for a general circuit is
% as hard as the graph isomorphism problem. We make use of the ability to
% compute the orbit of a gate in polynomial time in sub-section
% \ref{sec:translating-formulas-to-FPR} when translating families of circuits
% into equivalent formulas. As such the approach in this paper would fail for
% the general circuit.

% \begin{prop}
%   There is a polynomial-time reduction from the graph isomorphism problem to
%   the problem of computing if two gates in a given circuit are in the same
%   orbit.
%   \label{prop:graph-iso-to-orbit}
% \end{prop}
% \begin{proof}
%   Let $C$ be a circuit. Let $g_1$ and $g_2$ be two gates in $C$ such that
%   $H_{g_1} = H_{g_2}$. Then there exists $\pi \in \aut(C)$ such that $\pi
%   (g_1) = g_2$ if, and only if, $g_1$ is syntactically equivalent to $g_2$. As
%   such, we may use the same construction as in
%   Proposition~\ref{prop:syntactic-graph-iso} to prove this result.
% \end{proof}

% \begin{lem}
%   There is a polynomial-time Turing reduction from the problem of determining
%   if a given pair of gates in a given circuit are syntactically equivalent to
%   the problem of deciding if a circuit is transparent.
%   %   If the problem of deciding if a given circuit is transparent can be
%   %   decided
%   %   in
%   %   polynomial time then the graph isomorphism problem is in polynomial
%   %   time.
% \end{lem}
% \begin{proof}
%   Suppose there is an algorithm that takes as input, runs in time polynomial
%   in the size of the circuit, and decides whether or not the circuit is
%   transparent.

%   From Corollary \ref{} it is sufficient to show that there is a
%   polynomial-time algorithm for the problem of deciding if a pair of gates in
%   a given circuit are syntactically equivalent. Let $C_n$ be a circuit and
%   $g_1$ and $g_2$ be two gates in the circuit.

%   %   We simply need a polynomial-time reduction from the problem of computing
%   %   t
%   %   to the problem
%   %   of checking if a given circuit is transparent. Let $C_n =\langle G,
%   %   \Omega,
%   %   \Sigma, \Lambda, L\rangle$ be a circuit and let $g_1$ and $g_2$ be two
%   %   gates
%   %   in $G$. We now construct a circuit $C_n'$. If $g_1 = g_2$ then let
%   %   $C_n'$ be
%   %   any transparent circuit. If $g_1$ and $g_2$ do not satisfy any of
%   %   conditions
%   %   (i), (ii), (iii) or (v) of syntactic equivalence then let $C_n'$ be any
%   %   non-transparent circuit. So suppose all these conditions are met and
%   %   $g_1
%   %   \neq
%   %   g_2$, then let $D_n$ be the sub-circuit of $C_n$ consisting of exactly
%   %   those
%   %   gates $g$ in $C_n$ such that $W_t(g, g_1)$ or $W_t(g, g_2)$. Then let
%   %   $C_n'$
%   %   be $D_n'$ with an additional gate $g_\rank$, such that $g_\rank $ is
%   %   labelled
%   %   by the function $\rank^1_2[2,1]$ and $g_\rank$
% \end{proof}

% \begin{corollary}
% \end{corollary}

% We have from this proposition that even many nice properties Since there is a
% polynomial-time reduction from this problem reduces the general problem of
% deciding the syntatic equivilence relation for an arbitary

% \begin{corollary}
%   There is a polynomial-time reduction from the graph isomorphism problem to
%   the problem of determining if a given $(\mathbb{B}, \tau)$-circuit is rigid.
% \end{corollary}

% \begin{corollary}
%   There is a polynomial-time reduction from the graph isomorphism problem to
%   the problem of determining if a given gate in a given $(\mathbb{B},
%   \tau)$-circuit has unique labels.
% \end{corollary}

% reduction from the graph isomorphism from to the problem of of computing the
% the syntactic equivalence relation on the gates of an arbitrary $(\mathbb{B},
% \tau)$-circuit.

% It is worth noting that if we restrict the input to a class of circuits
% defined over symmetric bases then we can decide the syntactic equivalence
% relation in polynomial time.

% % \begin{prop}
% %   Let $C_n$ be a circuit with symmetric gates. There is an algorithm that
% %   takes in such a circuit, runs in time polynomial in the size of the circuit,
% %   and outputs the syntactic equivalence relation on the gates of $C_n$.
% %   \label{prop:symmetric-syntactic}
% % \end{prop}
% % \begin{proof}
% % \end{proof}

% Moreover, it turns out that being able to compute the syntactic equivalence
% relation in polynomial time for some class of circuits $\mathcal{C}$ implies
% that the existence of a polynomial-time algorithm for converting circuits in
% $C$ into equivalent rigid circuit.

% % \begin{prop}
% %   Let $\mathcal{C}$ be a set of circuits. Suppose the syntactic equivalence
% %   relation can be computed in polynomial time for circuits in $\mathcal{C}$.
% %   It follows that there is an algorithm that runs in polynomial time that
% %   takes in a circuit $C \in \mathcal{C}$ and outputs a rigid circuit $C''$
% %   such that if $C$ is symmetric then $C'$ is symmetric.
% %   \label{prop:syntactic-equivilence-rigid}
% % \end{prop}
% % \begin{proof}
% % \end{proof}

% The following result gives us that the unique labels condition is sufficient.

% % \begin{prop}
% %   Let $\mathcal{C}$ be syntactically transparent class of circuits. It follows
% %   that there exists an algorithm that takes in a $(\mathbb{B}, \tau)$-circuit
% %   $C \in \mathcal{C}$ and outputs a $(\mathbb{B} \cup \mathbb{B}_{\std},
% %   \tau)$-circuit $C'$ such that $C$ and $C'$ compute the same function, $C'$
% %   is rigid, and if $C$ is symmetric then $C'$ is symmetric. Moreover, this
% %   algorithm runs in time polynomial in the size of the input circuit.
% % \end{prop}
% % \begin{proof}
% %   Let $C' = \make-injective-all(\op{merge-all} (C, G)($). We have that $C'$ is
% %   symmetric if $C$ is symmetric. Since the syntactic equivalence relation can
% %   be computed in polynomial time for circuits in $\mathcal{C}$ it follows that
% %   this operation can be implemented so as to run in time polynomial in the
% %   size of $C$. Clearly every gate $g$ in $C'$ belongs to a singleton syntactic
% %   equivalence class.
% % \end{proof}

% \begin{lem}
%   Let $\mathbb{B}$ be a basis of symmetric functions. Then for any class of
%   circuits $\mathcal{C}$ defined over the basis $\mathbb{B}$, $\mathcal{C}$ is
%   a syntactically transparent class of circuits.
% \end{lem}
% In contrast to
% \begin{lem}
%   If the class of all circuits is syntactically transparent then then graph
%   isomorphism problem is in $\PT$.
% \end{lem}

% \begin{lem}
%   Let $\mathcal{C}$ be a syntactically transparent class of circuits. There is
%   an algorithm that takes as input a $(\mathbb{B}, \tau)$-circuit $C \in
%   \mathcal{C}$ and decides if $C$
% \end{lem}

% Since the class of The following result, also proved by Anderson and Dawar
% \cite{AndersonD17}, follows as a corollary.

% \begin{lem}
%   Let $\mathcal{C}$
% \end{lem}

% \begin{lem}
%   Let $C_n = \langle G, \Omega, \Sigma, \Lambda, L\rangle$ be a $(\mathbb{B},
%   \tau)$-circuit computing a $q$-ary query. If $C_n$ has unique labels then it
%   is $C_n$ rigid
%   \label{lem:unique-implies-rigid}
% \end{lem}
% \begin{proof}

%   %   Suppose there exists a gate $g'$ such that $W_t (g, g')$ and $W_t
%   %   (\pi(g),
%   %   g')$ and $\pi (g') = g'$, then $\pi H_{g'} = H_{g'}$, and since $g'$ has
%   %   unique labels it follows that $\pi$ acts trivially on $H_{g'}$ (i.e. for
%   %   all
%   %   $h \in H_{g'}$, $\pi(h) = h$). Moreover, for any pair of gates $h,h' \in
%   %   G$
%   %   such that $W(h, h')$ and $W_t(h', g')$, if $\pi$ acts on trivially on
%   %   the
%   %   children of $h'$ then $\pi (h) = h$ and so $\pi$ must act trivially

%   %   $\pi$ acts trivially on the children of $h$.

%   %   If $\Omega_g \cap \Omega_{\pi(g)} = \emptyset$ then, since
%   %   $\Omega_{\pi(g)}
%   %   =
%   %   \pi \Omega_g$, there exists $g' \in \Omega_g$ such that $\pi g' \neq
%   %   g'$.
%   %   But
%   %   $\pi$ must act trivially on output gates, so this is a contradiction. So
%   %   there
%   %   exists $g' \in \Omega_g \cap \Omega_{\pi(g)}$. But then $W_t (g, g')$
%   %   and
%   %   $W_t
%   %   (\pi (g), g')$.

%   %   We have that $g$ must be a non-output internal gate. So $g$ is
%   %   syntactically
%   %   equivalent to $\pi (g)$.

%   %   and let $h$ be a gate such that $W_t (g, h)$.
% \end{proof}

% \begin{prop}
%   Let $C_n$ be a circuit with unique labels. If
% \end{prop}

% \begin{prop}

% \end{prop}
% \begin{proof}
% \end{proof}

% \begin{prop}
%   Let $\mathcal{C}$ be a class on which the Let $C_n := \langle G, \Omega,
%   \Sigma, \Lambda, L \rangle$ be a $(\mathbb{B}, \tau)$-circuit with unique
%   labels. There is a deterministic algorithm that takes in such a circuit and
%   outputs the syntactic equivalence relation on $G$. Moreover, this algorithm
%   runs in time polynomial in the size of the circuit.
%   \label{prop:unique-labels-syntactic-equiv}
% \end{prop}

% % \begin{prop}
% %   There is an algorithm that takes in a $(\mathbb{B}, \tau)$-circuit $C$ with
% %   unique labels and outputs the syntactic equivalence relation on its gates.
% %   This algorithm runs in time polynomial in the size of $C$.
% %   \label{prop:unique-labels-syntactic-equiv}
% % \end{prop}
% % \begin{proof}
% %   We build of the relation by induction. Let $g$
% % \end{proof}

% Putting together Propositions \ref{prop:symmetric-syntactic},
% \ref{prop:syntactic-equivilence-rigid}, and
% \ref{prop:unique-labels-syntactic-equiv}, we have that there is a
% polynomial-time algorithm that takes as input a circuit with unique labels and
% outputs an equivalent rigid circuit. In fact, there is an equivalence of sorts
% between the unique labels condition and the rigidity condition.

% \begin{prop}
%   Let $\mathcal{C}$ be a family of circuits. There is an algorithm that runs
%   in polynomial-time that takes in a circuit $C \in \mathcal{C}$ and outputs a
%   rigid circuit $C'$ such that if $C$ is symmetric then $C'$ is symmetric if,
%   and only if, there is an algorithm that runs in polynomial-time and takes in
%   a circuit $C \in \mathcal{C}$ and outputs a circuit $C'$ with unique labels
%   such that if $C$ is symmetric then $C'$ is symmetric.
% \end{prop}

% It follows then that for the symmetric circuits defined over bases of
% symmetric functions discussed by Anderson and Dawar~\cite{} we may assume the
% circuits are rigid and/or have unique labels without a loss of generality. For
% the more general circuit discussed in this paper we cannot make such an
% assumption as deciding rigidity is harder then graph isomorphism. However, if
% we restrict ourselves to families of circuits with unique labels then again we
% may assume rigidity without a loss of generality.

% Of course, this restriction to circuits with unique labels is only useful if
% we can show that formulas in rank logic can be translated into $P$-uniform
% families circuits with unique labels. In the next subsection we prove that
% this is indeed the case.

\end{document}