### 2018-03-01 sections/conclusion.tex
A complete draft minus the conclusion. Large changes made to a number of sections.

\documentclass[../paper.tex]{subfiles}
\begin{document}
I have commented out what I had written here and will review it and write it as
soon as I have time.

% In the quest for a logic that captures polynomial time

% At the time of writing there are two candidate logics which 
% Recent work on the problem of capturing $\PT$ has produced two `candidate'
% logics for capturing polynomial-time.
% In this paper we extend Anderson and Dawar's symmetric circuit characterisation
% of $\FPC$, showing that $P$-uniform families of transparent symmetric circuits
% with rank gates have the same expressive power as $\FPR$. Since $\FPR$ remains
% one of the two c
% the candidate logics for capturing $\PT$, along with choiceless polynomial time, 
\end{document}