### 2018-03-10 sections/conclusion.tex
I have been trying to improve the presentation of the translation of circuits into formulas. To this end, I have changed the notation we use to talk about the usual logical operators (e.g. AND, OR, NOT, etc). I have also added tried to simplify the writing, and better explain the approach taken when translating circuits into formulas. I am in the process of writing the conclusion, but there is so much to think about.

\documentclass[../paper.tex]{subfiles}
\begin{document}

The main result of this paper establishes an equivalence between the expressive
power of $\FPR$ and the computational power of $\PT$-uniform transparent
symmetric rank-circuits, generalising a similar result by Anderson and
Dawar~\cite{AndersonD17}. Our motivation for proving this characterisation  The hope is that this characterisation will
contribute to our understanding of one of the two candidate logics for capturing
$\PT$.

% In the quest for a logic that captures polynomial time

% At the time of writing there are two candidate logics which Recent work on the
% problem of capturing $\PT$ has produced two `candidate' logics for capturing
% polynomial-time. In this paper we extend Anderson and Dawar's symmetric
% circuit characterisation of $\FPC$, showing that $P$-uniform families of
% transparent symmetric circuits with rank gates have the same expressive power
% as $\FPR$. Since $\FPR$ remains one of the two c the candidate logics for
% capturing $\PT$, along with choiceless polynomial time,
\end{document}